---
title: "Untitled"
author: "Zavadska Daryna"
date: "2025-05-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(DT)

library("tidyverse")
library("ggplot2")
library("data.table")
library("scales")
library("ggforce")
library("dgof")
library("remotes")
library("smmr")
#library("coin")

#library("rcompanion")

library("agricolae")
library("microshades")
library("gridExtra")

library(cluster)  

library(quantreg)
library(rcompanion) # for Efron's pseudo r-squared

library(ggpubr) #for geom_bracket

```


# !! FUNCTIONS

### preprocessing (outlier removal) function

```{r, include=FALSE}

fluo_preprocess <- function(all_meas, debugplot=FALSE, outlier_method = "IQR") {
all_meas_quart <- data.frame()

for (slide in unique(all_meas$file_prefix)) {
for (sp in unique(all_meas$est_sp)) {
  
  one_sp <- all_meas[all_meas$est_sp == sp, ]
  one_sp <- one_sp[one_sp$file_prefix == slide, ]

  quant_filtered_df <- one_sp
  
  if (outlier_method == "QUART") {
    ##### filtering quartiles
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int >= summary(one_sp$nucl_sum_int)[2]), ]
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int <= quantile(one_sp$nucl_sum_int)[4]), ]
  }
  else if(outlier_method == "IQR") {
      ##### filtering outliers
       qnt <- quantile(quant_filtered_df$nucl_sum_int, probs=c(.25, .75))
   H <- 1.5*IQR(quant_filtered_df$nucl_sum_int)
   quant_filtered_df <- quant_filtered_df[intersect(which(quant_filtered_df$nucl_sum_int < (qnt[2] + H) ) , which(quant_filtered_df$nucl_sum_int > (qnt[1] - H) ) ), ]
 
  }
  else(print("outlier_method should be either IQR or QUART(default), no other values currently allowed!"))

  
  all_meas_quart <- rbind(all_meas_quart, quant_filtered_df)
  
    
  if(debugplot == TRUE){
  print( ggplot(one_sp, aes(est_sp, as.numeric(nucl_sum_int)))+geom_boxplot()+
     geom_jitter(position=position_jitter(height=0, width=0.2, seed = 1234), alpha = 0.7)+ #geom_point(aes(shape = sp, colour = file_prefix), size = 2, alpha = 0.7)+
    theme_classic()+ scale_y_continuous(labels = label_comma())+
    theme(axis.text.x = element_text(angle =90))+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+ggtitle(paste0("After filtering", slide)) )

  print( ggplot(quant_filtered_df, aes(est_sp, as.numeric(nucl_sum_int)))+geom_boxplot()+
     geom_jitter(position=position_jitter(height=0, width=0.2, seed = 1234), alpha = 0.7)+ #geom_point(aes(shape = sp, colour = file_prefix), size = 2, alpha = 0.7)+
    theme_classic()+ scale_y_continuous(labels = label_comma())+
    theme(axis.text.x = element_text(angle =90))+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+ggtitle(paste0("After filtering", slide)) )
  }
  
}
}
return(all_meas_quart)
}
```


```{r}

fluo_preprocess_adjustable <- function(all_meas, debugplot=FALSE, slideORseries = "slide", outlier_method = "IQR") {
all_meas_quart <- data.frame()

if(slideORseries == "slide")  { LIST <- unique(all_meas$file_prefix) } 
  else if(slideORseries == "series") { LIST <- unique(all_meas$exp_series) }
  else{print("slideORseries must be either slide or series - no other options are supported so far.")}
    
  for (UNIT in LIST) {
  
  if(slideORseries == "slide")  {
    one_sp1 <- all_meas[all_meas$file_prefix == UNIT,] } 
    else if(slideORseries == "series") {
    one_sp1 <- all_meas[all_meas$exp_series == UNIT,]
  }



for (sp in unique(one_sp1$est_sp)) {
  
  one_sp <- one_sp1[one_sp1$est_sp == sp, ]

  quant_filtered_df <- one_sp
  
  if (outlier_method == "QUART") {
    ##### filtering quartiles
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int >= summary(one_sp$nucl_sum_int)[2]), ]
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int <= quantile(one_sp$nucl_sum_int)[4]), ]
  }
  else if(outlier_method == "IQR") {
      ##### filtering outliers
       qnt <- quantile(quant_filtered_df$nucl_sum_int, probs=c(.25, .75))
   H <- 1.5*IQR(quant_filtered_df$nucl_sum_int)
   quant_filtered_df <- quant_filtered_df[intersect(which(quant_filtered_df$nucl_sum_int < (qnt[2] + H) ) , which(quant_filtered_df$nucl_sum_int > (qnt[1] - H) ) ), ]
 
  }
  else(print("outlier_method should be either IQR or QUART(default), no other values currently allowed!"))

  
  all_meas_quart <- rbind(all_meas_quart, quant_filtered_df)
  
    
  if(debugplot == TRUE){
  print( ggplot(one_sp, aes(est_sp, as.numeric(nucl_sum_int)))+geom_boxplot()+
     geom_jitter(position=position_jitter(height=0, width=0.2, seed = 1234), alpha = 0.7)+ #geom_point(aes(shape = sp, colour = file_prefix), size = 2, alpha = 0.7)+
    theme_classic()+ scale_y_continuous(labels = label_comma())+
    theme(axis.text.x = element_text(angle =90))+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+ggtitle(paste0("After filtering", UNIT)) )

  print( ggplot(quant_filtered_df, aes(est_sp, as.numeric(nucl_sum_int)))+geom_boxplot()+
     geom_jitter(position=position_jitter(height=0, width=0.2, seed = 1234), alpha = 0.7)+ #geom_point(aes(shape = sp, colour = file_prefix), size = 2, alpha = 0.7)+
    theme_classic()+ scale_y_continuous(labels = label_comma())+
    theme(axis.text.x = element_text(angle =90))+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+ggtitle(paste0("After filtering", UNIT)) )
  }
  
}
}
return(all_meas_quart)
}
```


### k-mean clusterization function

```{r, include=FALSE}

fluo_clustering_combiner <- function(sp_by_sp_quart, combination_factors = c("est_sp", "slidedate",  "exp_series_date", "") ) {
  
column_combinations <- combn(combination_factors, 2)
kmean_comb_out <- data.frame()


for (i in 1:ncol(column_combinations)) {
  #debug
 #print(paste(column_combinations[1, i], column_combinations[2, i]))
  
  for (est_sp in unique(sp_by_sp_quart[, column_combinations[1, i]])) {
    
    #if ("sp" %in% column_combinations[1, i]) { CLUSTER <- 2 }else{ CLUSTER <- 2 }
    
    dd_comb <- data.frame()
    
    if (column_combinations[2, i] %in% c("", NA)) {  # Handle missing column case properly
      local <- sp_by_sp_quart[sp_by_sp_quart[, column_combinations[1, i]] == est_sp,]
      
      if (nrow(local[local[, column_combinations[1, i]] == est_sp, ]) > 2) {
        # Clustering
        km <- kmeans(local$nucl_sum_int[local[, column_combinations[1, i]] == est_sp], 2)
        dd <- cbind(local[local[, column_combinations[1, i]] == est_sp, ], cluster = km$cluster)
        
        # Determine cluster labels
        dipl <- dd$cluster[dd$nucl_sum_int == max(dd$nucl_sum_int)][1]
        dd$ploidy_clust <- ifelse(dd$cluster == dipl, "2n", "1n")
        
        
        dd$suns_for_clust <- paste(column_combinations[1, i], column_combinations[2, i])
        

        
        dd_comb <- rbind(dd_comb, dd)
      } else {
        next
      }
    } else {
      for (slide in unique(sp_by_sp_quart[, column_combinations[2, i]])) {
        local <- sp_by_sp_quart[sp_by_sp_quart[, column_combinations[2, i]] == slide,]
        
      
        if (nrow(local[local[, column_combinations[1, i]] == est_sp,]) > 2) {
          # Clustering
          km <- kmeans(local$nucl_sum_int[local[, column_combinations[1, i]] == est_sp], 2)
          dd <- cbind(local[local[, column_combinations[1, i]] == est_sp, ], cluster = km$cluster)
          
          # Determine cluster labels
          dipl <- dd$cluster[dd$nucl_sum_int == max(dd$nucl_sum_int)][1]
          dd$ploidy_clust <- ifelse(dd$cluster == dipl, "2n", "1n")
          
          dd$suns_for_clust <- paste(column_combinations[1, i], column_combinations[2, i])
          
          
          dd_comb <- rbind(dd_comb, dd)
        } else {
          next
        }
      }
    }
    
    kmean_comb_out <- rbind(kmean_comb_out, dd_comb)
  }
}
## !!! SUBSETTING K-MEAN CLUSTERING OU ONLY TO THAT WITHIN SPECIES!!!!
kmean_comb_out <- kmean_comb_out[grep("est_sp", kmean_comb_out$suns_for_clust),]

kmean_comb_out$bp_exp_ploidy_corr <- NA

kmean_comb_out$bp_exp_ploidy_corr[kmean_comb_out$ploidy_clust =="1n"] <- kmean_comb_out$bp_exp[kmean_comb_out$ploidy_clust =="1n"] 
kmean_comb_out$bp_exp_ploidy_corr[kmean_comb_out$ploidy_clust =="2n"] <- kmean_comb_out$bp_exp[kmean_comb_out$ploidy_clust =="2n"]*2


return(kmean_comb_out)
}

```


```{r, include=FALSE}

fluo_clustering_combiner_and_filtrator <- function(sp_by_sp_quart, combination_factors = c("est_sp", "slidedate",  "exp_series_date", ""), outlier_method = "IQR" ) {
  
column_combinations <- combn(combination_factors, 2)
## !!! SUBSETTING K-MEAN CLUSTERING OU ONLY TO THAT WITHIN SPECIES (not permitting across-species clusterisation, as it makes no sense)!!!!
column_combinations <- column_combinations[, which(column_combinations[1,]=="est_sp")]
## !!! SUBSETTING K-MEAN CLUSTERING OU ONLY TO THAT WITHIN SPECIES!!!!

kmean_comb_out <- data.frame()


for (i in 1:ncol(column_combinations)) {
  #debug
 #print(paste(column_combinations[1, i], column_combinations[2, i]))
  
  for (est_sp in unique(sp_by_sp_quart[, column_combinations[1, i]])) {
    
    #if ("sp" %in% column_combinations[1, i]) { CLUSTER <- 2 }else{ CLUSTER <- 2 }
    
    dd_comb <- data.frame()
    
    if (column_combinations[2, i] %in% c("", NA)) {  # Handle missing column case properly
      local <- sp_by_sp_quart[sp_by_sp_quart[, column_combinations[1, i]] == est_sp,]
      
      
############## 
#  print(column_combinations[, i])
#  print(unique(local$est_sp))
#  print(nrow( local[local[, column_combinations[1, i]] == est_sp, ] ) > 2 )      
############## Pre-filtering      
  quant_filtered_df <- local
  
  if (outlier_method == "QUART") {
    ##### filtering quartiles
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int >= summary(local$nucl_sum_int)[2]), ]
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int <= quantile(local$nucl_sum_int)[4]), ]
  }
  else if(outlier_method == "IQR") {
      ##### filtering outliers
       qnt <- quantile(quant_filtered_df$nucl_sum_int, probs=c(.25, .75))
   H <- 1.5*IQR(quant_filtered_df$nucl_sum_int)
   quant_filtered_df <- quant_filtered_df[intersect(which(quant_filtered_df$nucl_sum_int <= (qnt[2] + H) ) , which(quant_filtered_df$nucl_sum_int >= (qnt[1] - H) ) ), ]
 
  }
  else(print("outlier_method should be either IQR or QUART(default), no other values currently allowed!"))
     
  local <- quant_filtered_df
   
############## 
############## Clustering      
      if (nrow(local[local[, column_combinations[1, i]] == est_sp, ]) > 2) {
        # Clustering
        km <- kmeans(local$nucl_sum_int[local[, column_combinations[1, i]] == est_sp], 2)
        dd <- cbind(local[local[, column_combinations[1, i]] == est_sp, ], cluster = km$cluster)
        
        # Determine cluster labels
        dipl <- dd$cluster[dd$nucl_sum_int == max(dd$nucl_sum_int)][1]
        dd$ploidy_clust <- ifelse(dd$cluster == dipl, "2n", "1n")
        
        
        dd$suns_for_clust <- paste(column_combinations[1, i], column_combinations[2, i])
############## Clustering 

############## Post-filtering      
  quant_filtered_df <- dd
  
  if (outlier_method == "QUART") {
    ##### filtering quartiles
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int >= summary(local$nucl_sum_int)[2]), ]
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int <= quantile(local$nucl_sum_int)[4]), ]
  }
  else if(outlier_method == "IQR") {
      ##### filtering outliers
       qnt <- quantile(quant_filtered_df$nucl_sum_int, probs=c(.25, .75))
   H <- 1.5*IQR(quant_filtered_df$nucl_sum_int)
   quant_filtered_df <- quant_filtered_df[intersect(which(quant_filtered_df$nucl_sum_int <= (qnt[2] + H) ) , which(quant_filtered_df$nucl_sum_int >= (qnt[1] - H) ) ), ]
 
  }
  else(print("outlier_method should be either IQR or QUART(default), no other values currently allowed!"))
     
  dd <- quant_filtered_df
   
############## 
        
        dd_comb <- rbind(dd_comb, dd)
    
      } else {
        next
      }
    } else {
      for (slide in unique(sp_by_sp_quart[, column_combinations[2, i]])) {
        local <- sp_by_sp_quart[sp_by_sp_quart[, column_combinations[1, i]] == est_sp,]
        local <- local[local[, column_combinations[2, i]] == slide,]
##############
# print(column_combinations[, i])
# print(unique(local$est_sp))
# print(nrow( local[local[, column_combinations[1, i]] == est_sp, ] ) > 2 )
############## Pre-filtering      
  quant_filtered_df <- local

  if (outlier_method == "QUART") {
    ##### filtering quartiles
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int >= summary(local$nucl_sum_int)[2]), ]
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int <= quantile(local$nucl_sum_int)[4]), ]
  }
  else if(outlier_method == "IQR") {
      ##### filtering outliers
       qnt <- quantile(quant_filtered_df$nucl_sum_int, probs=c(.25, .75))
   H <- 1.5*IQR(quant_filtered_df$nucl_sum_int)
   quant_filtered_df <- quant_filtered_df[intersect(which(quant_filtered_df$nucl_sum_int <= (qnt[2] + H) ) , which(quant_filtered_df$nucl_sum_int >= (qnt[1] - H) ) ), ]
 
  }
  else(print("outlier_method should be either IQR or QUART(default), no other values currently allowed!"))
     
  local <- quant_filtered_df
   
############## 
############## Clustering      
        if (nrow(local[local[, column_combinations[1, i]] == est_sp,]) > 2) {
          # Clustering
          km <- kmeans(local$nucl_sum_int[local[, column_combinations[1, i]] == est_sp], 2)
          dd <- cbind(local[local[, column_combinations[1, i]] == est_sp, ], cluster = km$cluster)
          
          # Determine cluster labels
          dipl <- dd$cluster[dd$nucl_sum_int == max(dd$nucl_sum_int)][1]
          dd$ploidy_clust <- ifelse(dd$cluster == dipl, "2n", "1n")
          
          dd$suns_for_clust <- paste(column_combinations[1, i], column_combinations[2, i])
############## Clustering            
############## Post-filtering      
  quant_filtered_df <- dd
#  print(unique(quant_filtered_df$est_sp))

  if (outlier_method == "QUART") {
    ##### filtering quartiles
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int >= summary(local$nucl_sum_int)[2]), ]
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int <= quantile(local$nucl_sum_int)[4]), ]
  }
  else if(outlier_method == "IQR") {
      ##### filtering outliers
       qnt <- quantile(quant_filtered_df$nucl_sum_int, probs=c(.25, .75))
   H <- 1.5*IQR(quant_filtered_df$nucl_sum_int)
   quant_filtered_df <- quant_filtered_df[intersect(which(quant_filtered_df$nucl_sum_int <= (qnt[2] + H) ) , which(quant_filtered_df$nucl_sum_int >= (qnt[1] - H) ) ), ]
 
  }
  else(print("outlier_method should be either IQR or QUART(default), no other values currently allowed!"))
     
  dd <- quant_filtered_df
   
############## 
                  
          dd_comb <- rbind(dd_comb, dd)
        
          
        } else {
          next
        }
      }
    }
    
    kmean_comb_out <- rbind(kmean_comb_out, dd_comb)
  }
}


kmean_comb_out$bp_exp_ploidy_corr <- NA

kmean_comb_out$bp_exp_ploidy_corr[kmean_comb_out$ploidy_clust =="1n"] <- kmean_comb_out$bp_exp[kmean_comb_out$ploidy_clust =="1n"] 
kmean_comb_out$bp_exp_ploidy_corr[kmean_comb_out$ploidy_clust =="2n"] <- kmean_comb_out$bp_exp[kmean_comb_out$ploidy_clust =="2n"]*2


return(kmean_comb_out)
}


```


### Quantile regression-based functions

```{r, include=FALSE}

quantile_for_clustered_KINETOS <- function(SUBS_sp, slideORseries = "slide", dataset, debugplot = FALSE ){

to_return <- data.frame()
plotlist <- list()

    if(slideORseries == "slide")  { LIST <- unique(SUBS_sp$file_prefix) } 
    else if(slideORseries == "series") { LIST <- unique(SUBS_sp$exp_series) }
else if(slideORseries == "ALL") { LIST <- c("ALL") }
  else{stop("slideORseries must be either slide, series or ALL - no other options are supported so far.")}
    
  for (UNIT in LIST) {
  
  if(slideORseries == "slide")  {
    local1 <- SUBS_sp[SUBS_sp$file_prefix == UNIT,] } 
    else if(slideORseries == "series") {
    local1 <- SUBS_sp[SUBS_sp$exp_series == UNIT,]}
    else if(slideORseries == "ALL") {
    local1 <- SUBS_sp
  }
  
 testkin_sp <- unique(local1$sp[grep( paste( unique(SUBS_sp$test_species), collapse = "|") , local1$sp)])  
  for (sp in testkin_sp) {
    #taking out one of test species - !!! THE ORDER OF NEXT TWO LINES MATTERS !!!
    
    TEST <- local1[local1$sp == sp,]
    local <- local1[!(local1$sp %in% testkin_sp),]
    #debug below
    if (nrow(local) ==0) next
      
   # print(unique(local$file_prefix))#uncomment for debug 
  #  print(unique(local$suns_for_clust))
    
    
  local$bp_exp_ploidy_corr <- as.numeric(local$bp_exp_ploidy_corr)
  local$nucl_sum_int <- as.numeric(local$nucl_sum_int)
#building regression  
  qreg <- rq(bp_exp_ploidy_corr~nucl_sum_int, data=local)

 #p-value
  summary <- summary.rq(qreg, se='boot')
  pvalue <- summary$coefficients[8]
  SE <- summary$coefficients[4]
  
  #calculating Efron's pseudo r-squared
  R2 <- as.numeric(rcompanion::accuracy(qreg)[[2]][9])
  #preparing dataframes

 ## tmp <- tmp[tmp$estimate != "nucl_df", ] - ONLY RELEVANT WHEN COUNTING KINETOS
 ## TEST <- tmp[grep(unique(tmp$test_species),tmp$sp), ] - ONLY RELEVANT WHEN COUNTING KINETOS
  
  TEST$if_test_sp <- "TEST_kineto"
  local$if_test_sp <- "CONTROL_species"
  local$lower <- "NA"
  local$higher <- "NA"
  
  #prediction step
  new <- data.frame(nucl_sum_int = as.numeric(TEST$nucl_sum_int))
  prediction <- data.frame(predict.rq(qreg, new,  interval = "confidence"))
  TEST$bp_exp <- prediction$fit
  TEST$lower <- prediction$lower
  TEST$higher <- prediction$higher
  
  TEST$bp_exp_ploidy_corr <- prediction$fit
  TEST$suns_for_clust <- unique(local$suns_for_clust)
  #TEST$cluster <- "Undefined"
  #TEST$ploidy_clust <- "Undefined"
  
  
  
  #combining dataframes
   local_combined <- rbind(local, TEST)
     local_combined$pvalue <- pvalue
     local_combined$SE <- SE
     local_combined$coeficcient <- coef(qreg)[["nucl_sum_int"]]
     local_combined$intercept <- coef(qreg)[["(Intercept)"]]
      local_combined$R2 <- R2
     local_combined$regression_set <- paste(sp,  unique(local$suns_for_clust) , UNIT)
     
if(debugplot == TRUE) {
       gg <-  ggplot(local_combined[-grep("_KINnucl_df", local_combined$est_sp), ], aes(nucl_sum_int, bp_exp_ploidy_corr) )+
          geom_ribbon(aes(ymin=as.numeric(lower), ymax=as.numeric(higher)  ), fill="cyan", alpha=0.2)+
          geom_point(aes(shape = est_sp, colour = est_sp),size =3, alpha = 0.5)+
          geom_abline(slope = coef(qreg)[["nucl_sum_int"]], intercept = coef(qreg)[["(Intercept)"]] )+
          scale_x_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
          scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
    theme_linedraw()+
    ggtitle(paste(sp,  unique(local$suns_for_clust) , UNIT) )+
    scale_shape_manual(values=1:20) +
    geom_text(aes(label = paste0("p-value = ", round(pvalue, digits=5) ) ), x = mean(local$nucl_sum_int), y = max(local$bp_exp_ploidy_corr), check_overlap = TRUE, size=5)

print(gg)
plotlist[[length(plotlist) + 1]] <- (gg) }  
     
     #giant_fitted <- rbind(giant_fitted, local_combined) 
   to_return <- rbind(to_return, local_combined)
  }
}
#print(paste(sp,  unique(local$suns_for_clust) , UNIT) )
     
if(debugplot == TRUE) {
pdf(paste0("KINETO_reg_debug_", slideORseries, dataset, ".pdf") )
for (p in plotlist) {
    print(p) }
dev.off()
}

return(to_return)


}

```

```{r, include=FALSE}

quantile_for_clustered_BOOTSTRAPS <- function(SUBS_sp, slideORseries = "slide", bad_slide_list=c(), bad_species_list=c() , dataset, debugplot = FALSE ) { 
  
  giant_fitted <- data.frame()
  plotlist <- list()
  
    if(slideORseries == "slide")  { LIST <- unique(SUBS_sp$file_prefix) } 
    else if(slideORseries == "series") { LIST <- unique(SUBS_sp$exp_series) }
    else if(slideORseries == "ALL") { LIST <- c("ALL") }
  else{stop("slideORseries must be either slide, series or ALL - no other options are supported so far.")}
    
  for (UNIT in LIST) {
  
  if(slideORseries == "slide")  {
    local1 <- SUBS_sp[SUBS_sp$file_prefix == UNIT,] } 
    else if(slideORseries == "series") {
    local1 <- SUBS_sp[SUBS_sp$exp_series == UNIT,]}
    else if(slideORseries == "ALL") {
    local1 <- SUBS_sp
  }
  
  
  for (sp in unique(local1$sp)) { 
     print(paste(UNIT,  sp))
    
    #taking out one of test species - !!! THE ORDER OF NEXT TWO LINES MATTERS !!!
    
    TEST <- local1[local1$sp == sp,]
    local <- local1[local1$sp != sp,]
        if (nrow(local) <=1) next
    # print(nrow(local))
    
  local$bp_exp_ploidy_corr <- as.numeric(local$bp_exp_ploidy_corr)
  local$nucl_sum_int <- as.numeric(local$nucl_sum_int)
#building regression  
  qreg <- rq(bp_exp_ploidy_corr~nucl_sum_int, data=local)

 #p-value
  summary <- summary.rq(qreg, se='boot')
  pvalue <- summary$coefficients[8]
  SE <- summary$coefficients[4]
  #preparing dataframes

  #calculating Efron's pseudo r-squared
  R2 <- as.numeric(rcompanion::accuracy(qreg)[[2]][9])
 ## tmp <- tmp[tmp$estimate != "nucl_df", ] - ONLY RELEVANT WHEN COUNTING KINETOS
 ## TEST <- tmp[grep(unique(tmp$test_species),tmp$sp), ] - ONLY RELEVANT WHEN COUNTING KINETOS
  
  TEST$if_test_sp <- "TEST_bootstrap"
  local$if_test_sp <- "CONTROL_species"
  local$lower <- "NA"
  local$higher <- "NA"
  
  #prediction step
   #exception for one slide, otherwise throws the error
  new <- data.frame(nucl_sum_int = as.numeric(TEST$nucl_sum_int))
 
    if (UNIT %in% bad_slide_list  & sp %in% bad_species_list ) { prediction <- data.frame(predict.rq(qreg, new,  interval = "none"))
    TEST$bp_exp <- prediction$predict.rq.qreg..new..interval....none..
  TEST$lower <- NA
  TEST$higher <- NA 
  TEST$bp_exp_ploidy_corr <- prediction$predict.rq.qreg..new..interval....none..
  } else
    { prediction <- data.frame(predict.rq(qreg, new,  interval = "confidence"))
  TEST$bp_exp <- prediction$fit
  TEST$lower <- prediction$lower
  TEST$higher <- prediction$higher 
  TEST$bp_exp_ploidy_corr <- prediction$fit }
  
  TEST$suns_for_clust <- dataset
#  TEST$cluster <- "Undefined"
#  TEST$ploidy_clust <- "Undefined"
  


  
  #combining dataframes
   local_combined <- rbind(local, TEST)
     local_combined$pvalue <- pvalue
     local_combined$SE <- SE
     local_combined$coeficcient <- coef(qreg)[["nucl_sum_int"]]
     local_combined$intercept <- coef(qreg)[["(Intercept)"]]
     local_combined$R2 <- R2
     local_combined$bootstrap_setting <- paste(sp, dataset, UNIT)
     
if(debugplot == TRUE) {
       gg <-  ggplot(local_combined, aes(nucl_sum_int, bp_exp_ploidy_corr) )+
    geom_point(aes(shape = est_sp, colour = est_sp),size =3, alpha = 0.5)+geom_abline(slope = coef(qreg)[["nucl_sum_int"]] , intercept =  coef(qreg)[["(Intercept)"]])+
    theme_linedraw()+
    ggtitle(paste(sp,  unique(local$suns_for_clust) , UNIT) )+
  geom_ribbon(aes(ymin=as.numeric(lower), ymax=as.numeric(higher)  ), fill="cyan", alpha=0.2)+
    scale_shape_manual(values=1:20) +
    geom_text(aes(label = paste0("p-value = ", round(pvalue, digits=5) ) ), x = mean(local$nucl_sum_int), y = max(local$bp_exp_ploidy_corr), check_overlap = TRUE, size=5)

print(gg)
plotlist[[length(plotlist) + 1]] <- (gg) }  
     

giant_fitted <- rbind(giant_fitted, local_combined)

  }  
  }
  
if(debugplot == TRUE) {
pdf(paste0("Bootstrap_reg_debug_", slideORseries, dataset, ".pdf") )
for (p in plotlist) {
    print(p) }
dev.off()
}

  return(giant_fitted)
}

```



### Downstream processing

```{r, include=FALSE}

bind_pred_vs_actual_BOOTSTRAPS <- function(pred, actual, slideORseries = "slide") {

pred$ALL_PARAMS <- paste(pred$sp, pred$cell_coord, pred$suns_for_clust, pred$file_prefix, pred$ploidy_cluster)
actual$ALL_PARAMS <- paste(actual$sp, actual$cell_coord,actual$suns_for_clust, actual$file_prefix, pred$ploidy_cluster)


#debug
# pred$if_test_sp[which((pred$ALL_PARAMS %in% actual$ALL_PARAMS) == FALSE)]
# actual$ALL_PARAMS[actual$cell_coord == "[1149, 1736]"] 
# unique(pred$ALL_PARAMS[pred$cell_coord == "[1149, 1736]"])
 
 
actual$validation <- "ACTUAL"
pred$validation <- "BOOTSTRAP-LIKE_PREDICTED"

for (i in unique(pred$ALL_PARAMS)) {
  pred$diff_from_actual[pred$ALL_PARAMS == i] <- pred$bp_exp[pred$ALL_PARAMS == i] - actual$bp_exp_ploidy_corr[actual$ALL_PARAMS == i]
  pred$PERCENTdiff_from_actual[pred$ALL_PARAMS == i] <- (pred$bp_exp[pred$ALL_PARAMS == i] - actual$bp_exp_ploidy_corr[actual$ALL_PARAMS == i])/actual$bp_exp_ploidy_corr[actual$ALL_PARAMS == i]
}

actual[ , names(pred)[(names(pred) %in% names(actual) )==FALSE]] <- NA

pred_actual_toplot <- rbind(pred, actual)

pred_actual_toplot$suns_for_clust <- paste(pred_actual_toplot$suns_for_clust, slideORseries)

return(pred_actual_toplot)
}

```

```{r}

predictedbp_median_estimator_OLD <- function(test_kinetos_all_clust, subset_for_median_estimation = c("ploidy__exp_series_date", "ploidy__slidedate", "ploidy", "slidedate", "exp_series_date")  ) {

median_summary_regressed <- data.frame()

if (subset_for_median_estimation %in%  c("ploidy__exp_series_date", "ploidy__slidedate", "ploidy", "slidedate", "exp_series_date") == FALSE) {
  stop("subset_for_median_estimation must be one of: 'ploidy__exp_series_date', 'ploidy__slidedate', 'ploidy', 'slidedate', 'exp_series_date'")
  
}
group_vars <- switch(
  subset_for_median_estimation,
  "ploidy" = c("ploidy_clust"),
  "slidedate"  = c("exp_series_date", "file_prefix"),
  "exp_series_date"   = c("exp_series_date"),
  "ploidy__exp_series_date" = c("ploidy_clust", "exp_series_date"),
  "ploidy__slidedate" = c("ploidy_clust", "exp_series_date", "file_prefix")
)



test_kinetos_all_clust <- data.frame(test_kinetos_all_clust)


grouped_data <- split(test_kinetos_all_clust, test_kinetos_all_clust[group_vars])

# Loop through each group
for (group_name in names(grouped_data)) {
  
  group_df <- grouped_data[[group_name]]
  


  # debug
  #print(paste("Processing:", paste(keys, collapse = ", ")))

  for (reger_type in unique(group_df$suns_for_clust)) {
  for (species in unique(group_df$est_sp)) {
    
   tmp <- group_df[group_df$suns_for_clust == reger_type, ]
   tmp <- tmp[tmp$est_sp == species,]
      #debug
    # print(unique(tmp$sp))
   #  print( length(unique(tmp$cell_coord))==nrow(tmp) )
  if (nrow(tmp)==0) next
   
   MS <- data.frame("sp" = species, 
                    #"file_prefix" = paste( unique(tmp$file_prefix) ), 
                    #"exp_series_date" = unique(tmp$exp_series_date), 
                    "MEDIAN_int" = median(tmp$nucl_sum_int),
                    
                    "MEDIAN_bp_exp" = median(tmp$bp_exp_ploidy_corr), 
                    "quant_lower" = quantile((tmp$bp_exp_ploidy_corr))[2],
                    "quant_higher" = quantile((tmp$bp_exp_ploidy_corr))[4],
                    
                    "mean_lower" = mean(as.numeric(tmp$lower)), 
                    "mean_higher" = mean(as.numeric(tmp$higher)),  
                    "method"=reger_type, "if_test_sp" = unique(tmp$if_test_sp), 
                    #"ploidy" = unique(tmp$ploidy_clust), 
                    "cell_count"=length(unique(tmp$cell_coord)), 
                    "clustered_BY" = subset_for_median_estimation, 
                    "ploidy_clust" = paste( unique(tmp$ploidy_clust) ) )
   
   
   MS$set_for_median <- paste( unique( tmp[, group_vars, drop=TRUE] ), collapse = "__")
   
   for ( i in c( 1:length(str_split(unique(MS$clustered_BY), "__", simplify = T)) ) ){
        MS[, str_split(unique(MS$clustered_BY), "__", simplify = T)[1,i] ] <- str_split(unique(MS$set_for_median), "__", simplify = T)[i]
   }

   
   
   median_summary_regressed <- rbind(median_summary_regressed, MS) 

  
  }}  
  
 
  }

return(median_summary_regressed)
}


```


```{r}

predictedbp_median_estimator <- function(test_kinetos_all_clust, subset_for_median_estimation = c("ploidy__exp_series_date", "ploidy__slidedate", "ploidy", "slidedate", "exp_series_date"), BS_replicates = 5000 ) {

median_summary_regressed <- data.frame()

if (subset_for_median_estimation %in%  c("ploidy__exp_series_date", "ploidy__slidedate", "ploidy", "slidedate", "exp_series_date") == FALSE) {
  stop("subset_for_median_estimation must be one of: 'ploidy__exp_series_date', 'ploidy__slidedate', 'ploidy', 'slidedate', 'exp_series_date'")
  
}
group_vars <- switch(
  subset_for_median_estimation,
  "ploidy" = c("ploidy_clust"),
  "slidedate"  = c("exp_series_date", "file_prefix"),
  "exp_series_date"   = c("exp_series_date"),
  "ploidy__exp_series_date" = c("ploidy_clust", "exp_series_date"),
  "ploidy__slidedate" = c("ploidy_clust", "exp_series_date", "file_prefix")
)



test_kinetos_all_clust <- data.frame(test_kinetos_all_clust)


grouped_data <- split(test_kinetos_all_clust, test_kinetos_all_clust[group_vars])

# Loop through each group
for (group_name in names(grouped_data)) {
  
  group_df <- grouped_data[[group_name]]
  


  # debug
  #print(paste("Processing:", paste(keys, collapse = ", ")))

  for (reger_type in unique(group_df$suns_for_clust)) {
  for (species in unique(group_df$est_sp)) {
    
   tmp <- group_df[group_df$suns_for_clust == reger_type, ]
   tmp <- tmp[tmp$est_sp == species,]
      #debug
    # print(unique(tmp$sp))
   #  print( length(unique(tmp$cell_coord))==nrow(tmp) )
  if (nrow(tmp)<=1) next
    if ("file_prefix" %in% group_vars & nrow(tmp)<=2) next

   
   #print(c(species, reger_type, nrow(tmp)))
   
   BS <- groupwiseMedian(bp_exp_ploidy_corr~1, tmp, R = BS_replicates)
   
   MS <- data.frame("sp" = species, 
                    #"file_prefix" = paste( unique(tmp$file_prefix) ), 
                    #"exp_series_date" = unique(tmp$exp_series_date), 
                    "MEDIAN_int" = median(tmp$nucl_sum_int),
                    
                    "MEDIAN_bp_exp" = median(tmp$bp_exp_ploidy_corr), 
                    "quant_lower" = quantile((tmp$bp_exp_ploidy_corr))[2],
                    "quant_higher" = quantile((tmp$bp_exp_ploidy_corr))[4],
                    
                    #older version below
                   # "mean_lower" = mean(as.numeric(tmp$lower)), 
                   # "mean_higher" = mean(as.numeric(tmp$higher)),  
                   
                  "mean_lower" =  BS$Bca.lower ,
                   "mean_higher" = BS$Bca.upper ,
                    "method"=reger_type, "if_test_sp" = unique(tmp$if_test_sp), 
                    #"ploidy" = unique(tmp$ploidy_clust), 
                    "cell_count"=length(unique(tmp$cell_coord)), 
                    "clustered_BY" = subset_for_median_estimation, 
                    "ploidy_clust" = paste( unique(tmp$ploidy_clust) ) )
   
   
   MS$set_for_median <- paste( unique( tmp[, group_vars, drop=TRUE] ), collapse = "__")
   
   for ( i in c( 1:length(str_split(unique(MS$clustered_BY), "__", simplify = T)) ) ){
        MS[, str_split(unique(MS$clustered_BY), "__", simplify = T)[1,i] ] <- str_split(unique(MS$set_for_median), "__", simplify = T)[i]
   }

   
   
   median_summary_regressed <- rbind(median_summary_regressed, MS) 

  
  }}  
  
 
  }

return(median_summary_regressed)
}

```



### Plots

```{r, include=FALSE}

ks_medians_estimator_within <- function(toplot, between_reg = "within"){
 
pval_summary <- data.frame()

# if(between_reg == "within")  { LIST1 <- toplot$suns_for_clust  } else 
#   if(between_reg == "between") { LIST1 <- toplot$slidedate  } else
#     {print("between_reg must be either within or between - no other options are supported so far.")}

LIST1 <- toplot$suns_for_clust 
for(UNIT1 in unique(LIST1)){

  df1 <- toplot[LIST1 == UNIT1,]
  df1 <-df1[df1$if_test_sp=="TEST_bootstrap",]
  df1 <- df1[ which(is.na(df1$bp_exp_ploidy_corr) == FALSE), ]
  
#  gg <- ggplot(df1, aes(sp, bp_exp_ploidy_corr) )+
#  geom_pointrange(aes(ymin=as.numeric(lower),ymax=as.numeric(higher)),  linewidth=0.5, alpha=0.25, size=0.5, position=position_jitter(height=0, width=0.2))+
#    geom_violin( alpha=0)+geom_boxplot( alpha=0)+ scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))+theme_bw() + ggtitle(paste(UNIT1))
#print(gg)
#plotlist <- list(plotlist, gg)

  
  for(sp in unique(df1$sp) ){
      for(ploidy in unique(df1$ploidy_clust) ){
        df <-df1[df1$sp==sp,]
        df <-df[df$ploidy_clust==ploidy,]
        
#     if(between_reg == "within")  { LIST2 <- df$slidedate  } 
#    else if(between_reg == "between") { LIST2 <- df$suns_for_clust  }
#    else{print("between_reg must be either within or between - no other options are supported so far.")}
    
LIST2 <- df$slidedate
  if(length(unique(LIST2)) <= 1) next
      
  comb <- combn(unique(LIST2), 2)
  ks_test_df <- data.frame()
  median_test_df <- data.frame()
  
 for (i in seq(1, length(comb), by=2) ) {
   X1 <- comb[i]
   X2 <- comb[i+1]
   
   #### debug
   # print(df$bp_exp_ploidy_corr[LIST2 == comb[i] ] )
   # print(df$bp_exp_ploidy_corr[LIST2 == comb[i+1] ] )
   ####
   
   kstest<- ks.test( df$bp_exp_ploidy_corr[LIST2 == comb[i] ], df$bp_exp_ploidy_corr[LIST2 == comb[i+1]] )
   local_KS <- data.frame("X1" = paste(X1, ploidy), "X2" =paste(X2, ploidy), "KS_D" = kstest[1], "KS_p_value" = kstest[2] )
    ks_test_df <- rbind(ks_test_df, local_KS)
    
  mmm <- data.frame( "counts"= c( df$bp_exp_ploidy_corr[LIST2 == comb[i] ] , df$bp_exp_ploidy_corr[LIST2 == comb[i+1]] ),
                    "group" = c( rep(comb[i], length(df$bp_exp_ploidy_corr[LIST2 == comb[i] ]) ), rep(comb[i+1], length(df$bp_exp_ploidy_corr[LIST2 == comb[i+1] ]) )  ) )
  
  MS <- smmr::median_test(mmm, counts, group)
  
  local_MS <- data.frame("X1" = paste(X1, ploidy), "X2" =paste(X2, ploidy), "Median_D" = c(MS[[3]][[2]][[1]]), "Median_test_p_value" = c(MS[[3]][[2]][[3]]))
    median_test_df <- rbind(median_test_df, local_MS)
 } 
 stat_df <- merge(median_test_df, ks_test_df)
 stat_df$subset <- paste(ploidy, sp, UNIT1)
 
position <- seq(max(df$bp_exp_ploidy_corr) , max(df$bp_exp_ploidy_corr)+(nrow(stat_df)-1)*10000000, by=10000000)

pval_summary <- rbind(pval_summary, stat_df)

  }
 
}}


pval_summary$sp_out <- str_replace(pval_summary$subset, " .*.", "")
return(pval_summary)

}


ks_medians_estimator_between <- function(toplot){
 
pval_summary <- data.frame()

# if(between_reg == "within")  { LIST1 <- toplot$suns_for_clust  } else 
#   if(between_reg == "between") { LIST1 <- toplot$slidedate  } else
#     {print("between_reg must be either within or between - no other options are supported so far.")}

LIST1 <- toplot$slidedate

for(UNIT1 in unique(LIST1)){

  df1 <- toplot[LIST1 == UNIT1,]
  df1 <-df1[df1$if_test_sp=="TEST_bootstrap",]
  df1 <- df1[ which(is.na(df1$bp_exp_ploidy_corr) == FALSE), ]
  
#  gg <- ggplot(df1, aes(sp, bp_exp_ploidy_corr) )+
#  geom_pointrange(aes(ymin=as.numeric(lower),ymax=as.numeric(higher)),  linewidth=0.5, alpha=0.25, size=0.5, position=position_jitter(height=0, width=0.2))+
#    geom_violin( alpha=0)+geom_boxplot( alpha=0)+ scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))+theme_bw() + ggtitle(paste(UNIT1))
#print(gg)
#plotlist <- list(plotlist, gg)

  
  for(sp in unique(df1$sp) ){
        df <-df1[df1$sp==sp,]

#     if(between_reg == "within")  { LIST2 <- df$slidedate  } 
#    else if(between_reg == "between") { LIST2 <- df$suns_for_clust  }
#     else{print("between_reg must be either within or between - no other options are supported so far.")}
    
LIST2 <- df$suns_for_clust 
  if(length(unique(LIST2)) <= 1) next
      
  comb <- combn(unique(LIST2), 2)
  ks_test_df <- data.frame()
  median_test_df <- data.frame()
  
 for (i in seq(1, length(comb), by=2) ) {
   X1 <- comb[i]
   X2 <- comb[i+1]
   
   #### debug
   # print(df$bp_exp_ploidy_corr[LIST2 == comb[i] ] )
   # print(df$bp_exp_ploidy_corr[LIST2 == comb[i+1] ] )
   ####
   
   kstest<- ks.test( df$bp_exp_ploidy_corr[LIST2 == comb[i] ], df$bp_exp_ploidy_corr[LIST2 == comb[i+1]] )
   local_KS <- data.frame("X1" = X1, "X2" =X2, "KS_D" = kstest[1], "KS_p_value" = kstest[2] )
    ks_test_df <- rbind(ks_test_df, local_KS)
    
  mmm <- data.frame( "counts"= c( df$bp_exp_ploidy_corr[LIST2 == comb[i] ] , df$bp_exp_ploidy_corr[LIST2 == comb[i+1]] ),
                    "group" = c( rep(comb[i], length(df$bp_exp_ploidy_corr[LIST2 == comb[i] ]) ), rep(comb[i+1], length(df$bp_exp_ploidy_corr[LIST2 == comb[i+1] ]) )  ) )
  
  MS <- smmr::median_test(mmm, counts, group)
  
  local_MS <- data.frame("X1" = X1, "X2" =X2, "Median_D" = c(MS[[3]][[2]][[1]]), "Median_test_p_value" = c(MS[[3]][[2]][[3]]))
    median_test_df <- rbind(median_test_df, local_MS)
 } 
 stat_df <- merge(median_test_df, ks_test_df)
 stat_df$subset <- paste(sp, UNIT1)
 
position <- seq(max(df$bp_exp_ploidy_corr) , max(df$bp_exp_ploidy_corr)+(nrow(stat_df)-1)*10000000, by=10000000)

pval_summary <- rbind(pval_summary, stat_df)

  }
 
}


pval_summary$sp_out <- str_replace(pval_summary$subset, " .*.", "")
return(pval_summary)

}

```

```{r, include=FALSE}

ks_medians_intermediate_plots_between <- function(pval_summary){
  
gg <- ggplot(pval_summary, aes(p.value, subset, colour = sp_out))+geom_point()+facet_wrap(~X1+X2)+theme_bw()+ggtitle("KS-test p-values")+
  theme(axis.text.y = element_blank())+geom_vline(xintercept = 0.05, colour = "red")
  ggsave("between_LOG_KS-test_p-values.pdf", gg, width = 15, height =15)

gg <- ggplot(pval_summary, aes(Median_test_p_value, subset, colour = sp_out))+geom_point()+facet_wrap(~X1+X2)+theme_bw()+ggtitle("Median test p-values")+
  theme(axis.text.y = element_blank())+geom_vline(xintercept = 0.05, colour = "red")
  ggsave("between_LOG_Median_test_p-values.pdf", gg, width = 15, height =15)
  
  
gg <- ggplot(pval_summary, aes(p.value))+geom_histogram()+facet_wrap(~X1+X2)+theme_bw()+ggtitle("KS-test p-values")
ggsave("between_LOG_KS-test_p-values_hist.pdf", gg, width = 15, height =15)

gg <- ggplot(pval_summary, aes(Median_test_p_value))+geom_histogram()+facet_wrap(~X1+X2)+theme_bw()+ggtitle("Median test p-values")
ggsave("between_LOG_Median_test_p-values_hist.pdf", gg, width = 15, height =15)
  
}
  
ks_medians_intermediate_plots_within <- function(pval_summary){

   pval_summary$comp <- paste(pval_summary$X1, "//VS//", pval_summary$X2)
gg <- ggplot(pval_summary_within, aes(p.value, comp, colour = sp_out))+geom_point()+facet_wrap(~subset)+theme_bw()+ggtitle("KS-test p-values")+
  theme(axis.text.y = element_blank())+geom_vline(xintercept = 0.05, colour = "red")
  ggsave("within_LOG_KS-test_p-values.pdf", gg, width = 15, height =15)

gg <- ggplot(pval_summary_within, aes(Median_test_p_value, comp, colour = sp_out))+geom_point()+facet_wrap(~subset)+theme_bw()+ggtitle("KS-test p-values")+
  theme(axis.text.y = element_blank())+geom_vline(xintercept = 0.05, colour = "red")
  ggsave("within_LOG_Median_test_p-values.pdf", gg, width = 15, height =15)
  
  
gg <- ggplot(pval_summary, aes(p.value))+geom_histogram()+facet_wrap(~subset)+theme_bw()+ggtitle("KS-test p-values")
ggsave("within_LOG_KS-test_p-values_hist.pdf", gg, width = 15, height =15)

gg <- ggplot(pval_summary, aes(Median_test_p_value))+geom_histogram()+facet_wrap(~subset)+theme_bw()+ggtitle("Median test p-values")
ggsave("within_LOG_Median_test_p-values_hist.pdf", gg, width = 15, height =15)
  
}


```


```{r, include=FALSE}

anova_coef_within <- function(subs){

#plotlist <- list()
coeff_within <- data.frame()

for (set in unique(subs$suns_for_clust)) {
  data <- subs[subs$suns_for_clust == set, ] 
  res.aov <- aov(coeficcient~species_out, data)

  tuk <- TukeyHSD(res.aov)
 
  tuk <- data.frame(tuk$species_out)
  tuk$comb <- rownames(tuk)#str_replace(rownames(tuk), "sp ", "" )
  
 tuk <- cbind(tuk , data.frame(str_split(tuk$comb, "-" , simplify = TRUE)) )

  
#  gg1 <- ggplot(data, aes( species_out , coeficcient) )+
#  geom_violin()+geom_boxplot(alpha = 0.5)+
#   geom_pointrange(aes(ymin=as.numeric(coeficcient-SE), ymax=as.numeric(coeficcient+SE),colour = slidedate ),  linewidth=1, alpha=0.5, position=position_jitter(height=0, width=0.2, seed = 123))+
#    geom_bracket(xmin=tuk$X1, xmax=tuk$X2, y.position = c(15:24), label = round(tuk$p.adj, digits =3) )+ 
#    ggtitle(set)+
#  theme_classic() + theme(axis.text.x = element_text(angle =90))+
#    theme(legend.position = "none")
  
#print(gg1)
#   plotlist <- list(plotlist, gg1)
   
   tuk$suns_for_clust <- set
   coeff_within <- rbind(coeff_within, tuk)
}

 
#pdf("cefficients_within_Quantile_bootstr.pdf")
#for (p in plotlist) { print(p) }
#dev.off()

return(coeff_within)

}

anova_coef_between <- function(subs){
  
#plotlist <- list()
coeff_within <- data.frame()

for (species_out in na.omit(unique(subs$species_out)) ){
  data <- subs[subs$species_out == species_out, ] 
  res.aov <- aov(coeficcient~suns_for_clust, data)

  tuk <- TukeyHSD(res.aov)
 
  tuk <- data.frame(tuk$suns_for_clust)
  tuk$comb <- rownames(tuk)#str_replace(rownames(tuk), "sp ", "" )
  
 tuk <- cbind(tuk , data.frame(str_split(tuk$comb, "-" , simplify = TRUE)) )

  
#  gg1 <- ggplot(data, aes( suns_for_clust , coeficcient) )+
#  geom_violin()+geom_boxplot(alpha = 0.5)+
#   geom_pointrange(aes(ymin=as.numeric(coeficcient-SE), ymax=as.numeric(coeficcient+SE),colour = slidedate ),  linewidth=1, alpha=0.5, position=position_jitter(height=0, width=0.2, seed = 123))+
#    geom_bracket(xmin=tuk$X1, xmax=tuk$X2, y.position = c(15:length(unique(data$suns_for_clust))), label = round(tuk$p.adj, digits =3) )+ 
#    ggtitle(species_out)+
#  theme_classic() + theme(axis.text.x = element_text(angle =90))+
#    theme(legend.position = "none")
  
#print(gg1)

#plotlist <- list(plotlist, gg1)
   
   tuk$suns_for_clust <- species_out
   coeff_within <- rbind(coeff_within, tuk)
}

 
#pdf("cefficients_between_Quantile_bootstr.pdf")
#for (p in plotlist) { print(p)  }
#dev.off()

return(coeff_within)

}

```






# PRE-PROCESSING (GENERATING ALL_MEAS AND SUBSET FROM IT)

### Expected ratios dataframe

```{r, include=FALSE}
expected_ratios <- data.frame("sp"= c("Diplo_1", "Diplo_2", "Capsa_1", "Capsa_2", "Cflexa", "Sros_1", "Sros_2", "Theca", "Yeast"),
"bp" =  c(280283133, 107915492, 27967784, 27398043, 56404751, 55440309, 55448948, 28680627, 11529746) )


species_1 <- c()
species_2 <- c()
median_ratio <- c()

for ( i in unique(expected_ratios$sp)) {
  for ( e in unique(expected_ratios$sp)) {
    species_1 <- c(species_1, i)
      species_2 <- c(species_2, e)
    
  median_1 <- median(as.numeric(expected_ratios$bp[which(expected_ratios$sp == i)]))
    median_2 <- median(as.numeric(expected_ratios$bp[which(expected_ratios$sp == e)]))
  median_ratio <- c(median_ratio, median_1/median_2)
}
}

exp_rat <- data.frame("sp_1" = species_1, "sp_2" = species_2, "median_ratio_observed" = median_ratio)
exp_rat_backup <- exp_rat

```

### Expected genome sizes dataframe

```{r}

expected_sizes_full <- data.frame("sp"= c("Diplo_1", "Diplo_2", "Capsa_1", "Capsa_2", "Cflexa", "Sros_1", "Sros_2", "Theca", "Yeast"),
"bp" =  c(280283133, 107915492, 27967784, 27398043, 56404751, 55440309, 55448948, 28680627, 11529746) )

expected_sizes_trunc <- data.frame("sp"= c("Diplo", "Capsa", "Cflexa", "Sros", "Theca", "Yeast"),
"bp" =  c(280283133, 27398043, 56404751, 55448948, 28680627, 11529746) )

```

### Overview of available measurements

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
list <- c("20032025_thickstacks_noHU","25022025","1022025", "14022025", "21122024", "12042025", "27052025", "7052025", "25042025", "12072025", "5072025", "12092025", "6092025", "9102025", "26092025","22112025")
# "20032025_thickstacks_HU", 

# , "26092025"
#, "7052025", "31032025" 
#list <- c("20032025")
meas_files <- data.frame("exp_series_date"=c(), "file"=c())

for (suffix in list) {
files <- list.files(paste0("/home/dzavadska/Data/kDNA/born_again_meas/MEAS_ANALYSIS/", suffix), pattern = "nucl_df")
tmp <- data.frame("exp_series_date"=rep(suffix, length(files)), "file"=c(files)) 
meas_files <- rbind(meas_files, tmp)
}

meas_files$file_prefix <- str_replace(meas_files$file, "-.*", "")
meas_files$date_prefix <- paste0(meas_files$exp_series_date, "_", meas_files$file_prefix )

parameters <- fread("UPDATED_images_taken_data.csv")
parameters$date_prefix <- paste0(parameters$exp_series_date, "_", parameters$file_prefix )

to_see <- data.frame()

for (i in unique(meas_files$date_prefix)) {
 print(i)
  tmp <- data.frame(parameters[grep(i,parameters$date_prefix)])
  tmp$images_processed <- nrow(meas_files[meas_files$date_prefix==i,])
  to_see <- rbind(to_see, tmp)
}

#to_see 

gg1 <- ggplot(to_see, aes(images_processed, date_prefix))+geom_point(aes(colour = test_species, shape = as.factor(quant_stain_exp)), size =3)+theme_linedraw()+theme(axis.text.x = element_text(angle =90))

ggsave("slides_measured_counts.png", gg1, height = 10, width = 10)

```




```{r, results='hide', message=FALSE, warning=FALSE}
list <- c("20032025_thickstacks_noHU", "25022025","1022025", "14022025", "21122024", "12042025", "27052025", "7052025", "25042025", "12072025", "5072025", "12092025", "6092025", "9102025", "26092025","22112025")
#"20032025_thickstacks_HU",

# , "26092025"
#, "7052025", "31032025" 
#list <- c("20032025")

all_meas <- data.frame()

for (suffix in list) {
  

KIN_nuclDNA_raw_meas <- list.files(paste0("/home/dzavadska/Data/kDNA/born_again_meas/MEAS_ANALYSIS/", suffix), pattern = "KIN_nuclDNA")

KIN_nuclDNA_all_meas <- data.frame()

for (i in KIN_nuclDNA_raw_meas) {
  df <- fread(paste0("/home/dzavadska/Data/kDNA/born_again_meas/MEAS_ANALYSIS/", suffix, "/", i))
  
  if (length(KIN_nuclDNA_raw_meas)==0) {print( paste("Warning! No Kinetoplastid nuclear DNA measurements found for slide", suffix));
    next}
  
  KIN_nuclDNA_all_meas <- rbind(KIN_nuclDNA_all_meas, df)
}

if (length(KIN_nuclDNA_raw_meas)==0) {}else{
KIN_nuclDNA_all_meas$estimate <- "KIN_nuclDNA"
KIN_nuclDNA_all_meas <- KIN_nuclDNA_all_meas[!is.na(KIN_nuclDNA_all_meas$sp),]
KIN_nuclDNA_all_meas$exp_series <- suffix
}

############
kDNA_raw_meas <- list.files(paste0("/home/dzavadska/Data/kDNA/born_again_meas/MEAS_ANALYSIS/", suffix), pattern = "kDNA_df")
kDNA_all_meas <- data.frame()

for (i in kDNA_raw_meas) {
  df <- fread(paste0("/home/dzavadska/Data/kDNA/born_again_meas/MEAS_ANALYSIS/", suffix, "/", i))
  
  if (length(kDNA_raw_meas)==0) {print( paste("Warning! No Kinetoplastid kDNA measurements found for slide", suffix));
    next}
  
  kDNA_all_meas <- rbind(kDNA_all_meas, df)
}

if (length(KIN_nuclDNA_raw_meas)==0) {}else{
kDNA_all_meas$estimate <- "kDNA_df"
kDNA_all_meas <- kDNA_all_meas[!is.na(kDNA_all_meas$sp),]
kDNA_all_meas$exp_series <- suffix
}
##############
nucl_df_raw_meas <- list.files(paste0("/home/dzavadska/Data/kDNA/born_again_meas/MEAS_ANALYSIS/", suffix), pattern = "nucl_df")
nucl_df_all_meas <- data.frame()

for (i in nucl_df_raw_meas) {
  df <- fread(paste0("/home/dzavadska/Data/kDNA/born_again_meas/MEAS_ANALYSIS/", suffix, "/", i))
  nucl_df_all_meas <- rbind(nucl_df_all_meas, df)
}
nucl_df_all_meas$estimate <- "nucl_df"
nucl_df_all_meas <- nucl_df_all_meas[!is.na(nucl_df_all_meas$sp),]
nucl_df_all_meas$exp_series <- suffix
###

nucl_df_all_meas_to_bind <- nucl_df_all_meas[,c( "V1", "id", "cell", "sp", "cell_coord", "cyto_mean_int", "nucl_sum_int", "nucl_sum_int_corr", "estimate", "exp_series" )]

  if (length(kDNA_raw_meas)==0 & length(KIN_nuclDNA_raw_meas)==0 ) {all_meas <- rbind(data.table(all_meas), nucl_df_all_meas_to_bind, use.names=FALSE)}
  else{
    all_meas <- rbind(data.table(all_meas), nucl_df_all_meas_to_bind, KIN_nuclDNA_all_meas, kDNA_all_meas,  use.names=FALSE)
    }



}
###

if (length(grep("ch1", all_meas$id))<nrow(all_meas)) {
  print("Warning! Some channels might have been incorrectly assigned!")
  print("check following images:")
  print(unique(all_meas$id[!str_detect(all_meas$id, "ch1")]))
}

all_meas$est_sp <- paste0(all_meas$sp, all_meas$estimate)

all_meas$expected_sizes <- NA 

for (sp in expected_sizes_trunc$sp) {
  all_meas$expected_sizes[all_meas$sp ==sp] <- expected_sizes_trunc$bp[expected_sizes_trunc$sp==sp]
}

all_meas_backup <- all_meas

# unique(all_meas_backup$exp_series[grep("Gemkin",all_meas_backup$sp)] )


```

### To rename shit filenames which cause tremendous mental damage and time loss, run text output of something like this in commandline, in corresponding folder.
Do not forget to update UPDATED_images_taken_data.csv with corresponding names.
```{bash, eval = FALSE , echo=FALSE, results='hide', message=FALSE, warning=FALSE, include=FALSE}

for f in exp_12_5_Gemkin_Cflexa_DAPI600ngml_thickstack-*_ch1_*_20032025.csv; do
  base="exp_12_5_Gemkin_Cflexa"
  date="20032025"
  # Extract stack part
  stack=$(echo "$f" | grep -o "thickstack-[^_]*")
  # Extract ch1 tag
  ch1=$(echo "$f" | grep -o "ch1_[^_]*")
  # Construct new filename
  new="${base}_${date}_DAPI600ngml_${stack}_${ch1}_nuclDNA_df.csv"
  echo mv "$f" "$new"
done

#Also, to substitute ids inside every file
for f in 25022025_exp_12_5_Gemkin_Cflexa*.csv; do sed -i 's/exp/25022025_exp/g' $f ; done

for f in exp_12_5_Gemkin_Cflexa_20032025_DAPI600ngml_thickstack*.csv; do sed -i 's/Cflexa_/Cflexa_20032025_/g' $f ; done

for f in exp_12_5_Gre65_Cflexa_20032025_DAPI600ngml_thickstack*.csv; do sed -i 's/DAPI/20032025_DAPI/g' $f ; done
for f in exp_12_5_Gre65_Sros_20032025_DAPI600ngml_thickstack*.csv; do sed -i 's/DAPI/20032025_DAPI/g' $f ; done


```


### Adding staining parameters

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
all_meas <- data.frame(all_meas)

all_meas$exp_series_date <- NA
all_meas$file_prefix <- NA
all_meas$AT_stain <- NA
all_meas$AT_stain_conc <- NA
all_meas$quant_stain <- NA
all_meas$quant_stain_conc <- NA
all_meas$AT_stain_laser_int <- NA
all_meas$AT_stain_exp <- NA
all_meas$quant_stain_laser_int <- NA
all_meas$quant_stain_exp <- NA
all_meas$replicate <- NA
all_meas$other <- NA
all_meas$test_species <- NA
all_meas$images_taken <- NA
all_meas$images_processed <- NA

parameters <- fread("UPDATED_images_taken_data.csv")
# !!!!!!!!!!! SUBSET PARAMETERS IF NEEDED (IN CASE OF IDENTICAL NAMES IN DIFFERENT SERIES TROWS ERRORS - BECAUSE YOU SHOULDNT NAME THEM THAT WAY)



for(pref in unique(parameters$file_prefix)){
  
  all_meas[grep(paste0(pref), all_meas$id), 13:ncol(all_meas)] <- parameters[grep(paste0("\\b", pref, "\\b"), parameters$file_prefix)]
}

#all_meas[grep("Gemkin", all_meas$sp), ]
#all_meas[grep(paste0(pref), all_meas$id),]

```




### Subsetting dataset

```{r, results='hide', message=FALSE, warning=FALSE}
sp_by_sp <- all_meas
#sp_by_sp <- sp_by_sp[-grep("_B" , sp_by_sp$file_prefix ),]

sp_by_sp <- sp_by_sp[sp_by_sp$quant_stain_exp == 12.5, ]
sp_by_sp <- sp_by_sp[sp_by_sp$quant_stain == "PI",]

#by amount of slides processed
tmp <- to_see[to_see$file_prefix %in% unique(sp_by_sp$file_prefix),]
sp_by_sp$images_processed <- tmp$images_processed[match(sp_by_sp$file_prefix, tmp$file_prefix)]


sp_by_sp <- sp_by_sp[which(sp_by_sp$sp != "Theca"),]

sp_by_sp <- sp_by_sp[which(sp_by_sp$sp != "Cflexa"),]

#sp_by_sp <- sp_by_sp[which(sp_by_sp$sp != "Yeast"),]

#sp_by_sp <- sp_by_sp[sp_by_sp$images_processed >= 5,]

```


### Matching with bp_exp and backing up

```{r, results='hide', message=FALSE, warning=FALSE}

sp_by_sp$bp_exp <- expected_sizes_trunc$bp[match(sp_by_sp$sp, expected_sizes_trunc$sp)]
sp_by_sp$slidedate <- paste(sp_by_sp$file_prefix, sp_by_sp$exp_series_date)

sp_by_sp_backup <- sp_by_sp

# unique(sp_by_sp_backup$exp_series[grep("Gemkin",sp_by_sp_backup$sp)] )
# sp_by_sp_backup[grep("Gemkin",sp_by_sp_backup$sp) ,] 
```


### here, there was subsetting to only control species, but removed in this version

at this point, and up till Bootstrap control start, sp_by_sp_backup == sp_by_sp. Both kinetos (actual test) and controls are IQR-filtered and clustered.

```{r}
#sp_by_sp <- sp_by_sp #[sp_by_sp$sp %in% expected_sizes_trunc$sp,]
```


# END OF PREPROCESSING, REGRESSIONS BELOW

### preprocessing
In this version, we apply (pre-prediction) clusterisation to outlier-filtered controls, as well as to kinetos - therefore, subsetting to only control species for clusterisation and assigning "Undefined" ploidy in this case makes few sense

```{r}

sp_by_sp_by_kmeans <- fluo_clustering_combiner_and_filtrator(sp_by_sp, combination_factors = c("est_sp", "slidedate",  "exp_series_date", ""), outlier_method = "IQR" ) 


# older version below, sp_by_sp_quart still used for pruducing unclustered dataset
 sp_by_sp_quart <- fluo_preprocess_adjustable(sp_by_sp, debugplot=FALSE, slideORseries = "series", outlier_method = "IQR")
# sp_by_sp_by_kmeans_unfilt <- fluo_clustering_combiner(sp_by_sp_quart)
# sp_by_sp_by_kmeans <- data.frame()
# for (ploidy in unique(sp_by_sp_by_kmeans_unfilt$ploidy_clust)) {
#  tmp <- sp_by_sp_by_kmeans_unfilt[sp_by_sp_by_kmeans_unfilt$ploidy_clust == ploidy, ]
#  tobind <- fluo_preprocess_adjustable(tmp, debugplot=FALSE, slideORseries = "series", outlier_method = "IQR")
#  sp_by_sp_by_kmeans <- rbind(sp_by_sp_by_kmeans, tobind) }



tmp <- sp_by_sp_quart
tmp$cluster <-  "Undefined" 
tmp$ploidy_clust <-  "Undefined" 
tmp$suns_for_clust <-  "NOT CLUSTERED" 
tmp$bp_exp_ploidy_corr <-  tmp$expected_sizes

sp_by_sp_by_kmeans_plus_uncl <- rbind(tmp, sp_by_sp_by_kmeans)

#in this version - must print both kinetos and control species
unique(sp_by_sp_by_kmeans$est_sp)

#debug for ploidy correction
#sp_by_sp_by_kmeans_plus_uncl[sp_by_sp_by_kmeans_plus_uncl$est_sp=="Gemkin_KINnucl_df",c("ploidy_clust", "suns_for_clust")]
```

### check how the outlier filtering and clusterisation went

```{r}

ggplot(sp_by_sp_by_kmeans[sp_by_sp_by_kmeans$suns_for_clust == "est_sp slidedate",], aes(interaction(ploidy_clust, est_sp), nucl_sum_int))+geom_boxplot()+
  geom_point(aes( colour=sp ), size =1, alpha = 0.15, position=position_jitter(height=0, width=0.2))+
  theme_bw()+ ggtitle("est_sp slidedate")+theme(axis.text.x = element_blank())+
  facet_wrap(~suns_for_clust)+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))


ggplot(sp_by_sp_by_kmeans[sp_by_sp_by_kmeans$suns_for_clust == "est_sp exp_series_date",], aes(interaction(ploidy_clust, est_sp), nucl_sum_int))+geom_boxplot()+
  geom_point(aes( colour=sp ), size =1, alpha = 0.15, position=position_jitter(height=0, width=0.2))+
  theme_bw()+ ggtitle("est_sp slidedate")+theme(axis.text.x = element_blank())+
  facet_wrap(~suns_for_clust)+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))

ggplot(sp_by_sp_by_kmeans[sp_by_sp_by_kmeans$suns_for_clust == "est_sp ",], aes(interaction(ploidy_clust, est_sp), nucl_sum_int))+geom_boxplot()+
  geom_point(aes( colour=sp ), size =1, alpha = 0.15, position=position_jitter(height=0, width=0.2))+
  theme_bw()+ ggtitle("est_sp slidedate")+theme(axis.text.x = element_blank())+
  facet_wrap(~suns_for_clust)+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))


```

prints the list of slides with 1 or less points of a given species per slide; 
such points might cause errors when attempting to run the regression across slides, so, for the regression across slides, they should be removed.


```{r}

sp_by_sp_by_kmeans_slidedate <- sp_by_sp_by_kmeans[grep("est_sp exp_series_date", sp_by_sp_by_kmeans$suns_for_clust), ]

for (sp in unique(sp_by_sp_by_kmeans_slidedate$sp)) {
  tmp1 <- sp_by_sp_by_kmeans_slidedate[sp_by_sp_by_kmeans_slidedate$sp == sp, ]
  for (slide in unique(tmp1$file_prefix)) {
         if ( nrow( tmp1[tmp1$file_prefix == slide, ] ) <= 1) {
         
         print(paste(sp, slide))  
         #print ( ggplot(tmp1[tmp1$file_prefix == slide, ], aes(file_prefix, nucl_sum_int))+
         # geom_point(aes(colour = ploidy_clust), alpha=0.15)+
         # facet_wrap(~ploidy_clust)+         # ggtitle(paste(sp, slide))+         # theme_bw()+theme(axis.text.x = element_text(angle =90, size =5)))
         
       }}}
```





### Quantile regression-based prediction of test KINETOS, cluster-based

```{r}

  test_kinetos_all <- data.frame()
 test_kinetos_slidebyslide <- data.frame()
test_kinetos_series <- data.frame()

for (dataset in unique(sp_by_sp_by_kmeans$suns_for_clust) ) {
  
  SUBS_sp <- sp_by_sp_by_kmeans[sp_by_sp_by_kmeans$suns_for_clust == dataset,]

  #used earlier for unclustered version, not required anymore
  # to_bind_kin <- sp_by_sp[grep(paste(unique(sp_by_sp$test_species), collapse = "|"), sp_by_sp$est_sp),]
  # to_bind_kin[ , names(SUBS_sp)[ !( names(SUBS_sp) %in% names(to_bind_kin) ) ] ] <- "Undefined"
  # SUBS_sp <- rbind(to_bind_kin, SUBS_sp)
  
  allreg <- quantile_for_clustered_KINETOS(SUBS_sp, slideORseries = "ALL", dataset, debugplot = F)
 slidereg <- quantile_for_clustered_KINETOS(SUBS_sp, slideORseries = "slide", dataset, debugplot = F)
seriesreg <- quantile_for_clustered_KINETOS(SUBS_sp, slideORseries = "series", dataset, debugplot = F)

test_kinetos_all <- rbind( test_kinetos_all, allreg)
test_kinetos_slidebyslide <- rbind(test_kinetos_slidebyslide,  slidereg)
test_kinetos_series <- rbind(test_kinetos_series, seriesreg )

}

#debug for ploidy correction
#test_kinetos_series[test_kinetos_series$est_sp=="Gemkin_KINnucl_df",c("ploidy_clust", "suns_for_clust")]

#ggplot(SUBS_sp[SUBS_sp$file_prefix == "exp_12_5_Gre65_Cflexa_20032025_DAPI600ngml_thickstack",], aes(est_sp, nucl_sum_int))+
#  geom_point(aes(colour = ploidy_clust))+
#  facet_wrap(~suns_for_clust)


```


### Summarising kineto predictions, creating output for INFERRED_MEDIANS_downstream.Rmd

```{r}

test_kinetos_all$suns_for_clust <- paste(test_kinetos_all$suns_for_clust, "ALL")
test_kinetos_slidebyslide$suns_for_clust <- paste(test_kinetos_slidebyslide$suns_for_clust, "slide")
test_kinetos_series$suns_for_clust <- paste(test_kinetos_series$suns_for_clust, "series")


#unique(test_kinetos_slidebyslide$exp_series_date[grep("Gemkin", test_kinetos_slidebyslide$sp)] )
```


### saving cell-by-cell dataframe (median summary function was here, now moved to another parser)

```{r}

 test_kinetos_all_clust <- rbind(test_kinetos_slidebyslide, test_kinetos_series, test_kinetos_all)
#test_kinetos_all_clust <- rbind(test_kinetos_series)

fwrite(test_kinetos_all_clust, "separate_cells_test_kinetos_all_clust.csv")

#median_summary_regressed[grep("Gemkin", median_summary_regressed$sp), ]

```


### dataframe for regression SE, p-value, R^2 etc.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#prepare corresponding kineto dataframe
test_kinetos_all_clust$merger1 <- paste(test_kinetos_all_clust$suns_for_clust, test_kinetos_all_clust$slidedate)
test_kinetos_all_clust1 <- test_kinetos_all_clust[!duplicated(test_kinetos_all_clust$merger1), ]

fwrite(test_kinetos_all_clust1, "reg_metrics_test_kinetos_all_clust.csv")

```
















# Quantile regression-based prediction - bootstrap-like

## Now,we actually subset to control species and keep operation on that dataset 
### !!! Subsetting to only control species !!! Required for clusterisation, BUT dont forget to merge back when predicting kinetos

```{r}
sp_by_sp_by_kmeans_plus_uncl <- sp_by_sp_by_kmeans_plus_uncl[sp_by_sp_by_kmeans_plus_uncl$sp %in% expected_sizes_trunc$sp,]

#now it should return only the control species
unique(sp_by_sp_by_kmeans_plus_uncl$est_sp)

```


```{r}

bad_slide_list <- c("D44_Sros_thickstack_A", "D44_Sros_thickstack_B", "20032025_thickstacks_HU", "13G_Sros_06092025_DAPI600ngml_thickstack","exp_12_5_Rhyncho_noHU_Capsa_noHU_Cflexa_DAPI600ngml_thickstack", "20032025_thickstacks_noHU", "26092025", "exp_12_5_Gre65_Cflexa_DAPI600ngml", "10D_Sros_06092025_DAPI600ngml_thickstack", "25022025_exp_12_5_Gemkin_Cflexa_DAPI600ngml", "Rhyncho_Cflexa_12092025_DAPI600ngml_thickstack", "OB23_Cflexa_09102025_DAPI600ngml_thickstack", "exp_12_5_Rhyncho_40nmHU_Capsa_100mMHU_Sros_DAPI600ngml_thickstack", "exp_12_5_INUKIN1_Cflexa_thickstack", "INUKIN_Cflexa_12092025_DAPI600ngml_thickstack", "INUKIN_Sros_22112025_thickstack", "INUKIN_Cflexa_09102025_DAPI600ngml_thickstack")
bad_species_list <-c("Capsa", "Sros", "Diplo", "Yeast")


#bad_slide_list_few_sp <- c("D44_Sros_thickstack_A", "D44_Sros_thickstack_B", "13G_Sros_06092025_DAPI600ngml_thickstack","exp_12_5_Rhyncho_noHU_Capsa_noHU_Cflexa_DAPI600ngml_thickstack", "20032025", "exp_12_5_Gre65_Cflexa_DAPI600ngml","10D_Sros_06092025_DAPI600ngml_thickstack", "exp_12_5_Gemkin_Cflexa_20032025_DAPI600ngml_thickstack", "exp_12_5_Rhyncho_40nmHU_Capsa_100mMHU_Sros_DAPI600ngml_thickstack","25022025_exp_12_5_Gemkin_Cflexa_Cruz_DAPI600ngml","25022025_exp_12_5_Gemkin_Cflexa_DAPI600ngml","exp_12_5_Rhyncho_Cflexa_DAPI600ngml","Rhyncho_Cflexa_12092025_DAPI600ngml_thickstack","10D_Cflexa_06092025_DAPI600ngml_thickstack","13G_Sros_12072025_DAPI600ngml_thickstack", "13G_Sros_26092025_thickstack", "exp_12_5_Rhyncho_40nmHU_Capsa_100mMHU_Sros_DAPI600ngml_thickstack", "exp_12_5_Rhyncho_40nmHU_Capsa_100mMHU_Sros_DAPI600ngml_thickstack", "exp_12_5_Rhyncho_40nmHU_Capsa_100mMHU_Sros_DAPI600ngml_thickstack", "Rhyncho_Cflexa_12092025_DAPI600ngml_thickstack")
# bad_species_list <-c("Capsa", "Sros", "Diplo", "Yeast")

bootstr_all <- data.frame()
bootstr_slide <- data.frame()
bootstr_series <- data.frame()

for (dataset in unique(sp_by_sp_by_kmeans_plus_uncl$suns_for_clust) ) {
  
SUBS_sp <- sp_by_sp_by_kmeans_plus_uncl[sp_by_sp_by_kmeans_plus_uncl$suns_for_clust == dataset,]

#SUBS_sp[grep("exp_12_5_Gemkin_Cflexa_20032025_DAPI600ngml_thickstack", SUBS_sp$id),]

reg_all <- quantile_for_clustered_BOOTSTRAPS(SUBS_sp, slideORseries = "ALL", dataset = dataset, debugplot = F)
reg_slide <- quantile_for_clustered_BOOTSTRAPS(SUBS_sp, slideORseries = "slide", bad_slide_list = bad_slide_list, bad_species_list = bad_species_list, dataset, debugplot = F)
reg_series <- quantile_for_clustered_BOOTSTRAPS(SUBS_sp, slideORseries = "series", bad_slide_list = bad_slide_list, bad_species_list = bad_species_list, dataset = dataset, debugplot = F)

bootstr_all <- rbind(bootstr_all, reg_all)
bootstr_slide <- rbind(bootstr_slide, reg_slide)
bootstr_series <- rbind(bootstr_series, reg_series)

}

```


## In the old version, regressions were filtered by p-value at this step. Now they are filtered later, so this chunk is kept as renment - and later re-ran, with manual specification of ifts that should be deleted
Manually inspect graphs generated in pdfs for the ones with flat slopes, shit p-values etc.
 Alternatively, filter by p-value as is done automatically below. regressions with p-values \> 0.05 as a rule have near-zero coefficient

```{r}

bootstr_all_pfilt <- bootstr_all#[bootstr_slide$pvalue <= 0.05, ]
bootstr_slide_pfilt <- bootstr_slide#[bootstr_slide$pvalue <= 0.05, ]
bootstr_series_pfilt <- bootstr_series#[bootstr_series$pvalue <= 0.05, ]

```

## Postprocess predictions so that they are put in a same dataframe as actual values, and bind regressions over different subsets (slide or series) in the same dataframe

```{r}
to_bind <- sp_by_sp_by_kmeans_plus_uncl[sp_by_sp_by_kmeans_plus_uncl$sp %in% unique(expected_sizes_trunc$sp), ]
###
#put in a same dataframe as actual values
toplot_all <- bind_pred_vs_actual_BOOTSTRAPS(bootstr_all_pfilt, to_bind, slideORseries = "ALL")
toplot_slide <- bind_pred_vs_actual_BOOTSTRAPS(bootstr_slide_pfilt, to_bind, slideORseries = "slide")
toplot_series <- bind_pred_vs_actual_BOOTSTRAPS(bootstr_series_pfilt, to_bind, slideORseries = "series")

#bind regressions over different subsets (slide or series) in the same dataframe

toplot <- rbind(toplot_slide, toplot_series, toplot_all)
# toplot <- rbind(toplot_series)

#Check if settings actually had different values
# length( which((toplot_slide$bp_exp_ploidy_corr == toplot_series$bp_exp_ploidy_corr) == FALSE) ) > 0

unique(toplot$ploidy_clust[toplot$if_test_sp == "TEST_bootstrap"])


```



# PLOTS and summary

### adding variables which will be used for "beatiful" plots

```{r}


###!!! Order of the next lines matters!                       
toplot$beatiful_suns <- toplot$suns_for_clust

toplot$beatiful_suns <- str_replace(toplot$suns_for_clust, "NOT CLUSTERED " , "NOT CLUSTERED; ")
toplot$beatiful_suns <- str_replace(toplot$beatiful_suns, "est_sp exp_series_date " , "II; ")
toplot$beatiful_suns <- str_replace(toplot$beatiful_suns, "est_sp slidedate " , "III; ")
toplot$beatiful_suns <- str_replace(toplot$beatiful_suns, "est_sp " , "I; ")

toplot$beatiful_suns <- str_replace(toplot$beatiful_suns, "ALL" , "I")
toplot$beatiful_suns <- str_replace(toplot$beatiful_suns, "series" , "II")
toplot$beatiful_suns <- str_replace(toplot$beatiful_suns, "slide" , "III")

toplot$beatiful_sp <- toplot$sp

toplot$beatiful_sp <- str_replace(toplot$sp, "Sros" , "Salpingoeca rosetta")
toplot$beatiful_sp <- str_replace(toplot$beatiful_sp, "Capsa" , "Capsaspora owczarzaki")
toplot$beatiful_sp <- str_replace(toplot$beatiful_sp, "Diplo" , "Diplonema papillatum")
toplot$beatiful_sp <- str_replace(toplot$beatiful_sp, "Cflexa" , "Choanoeca flexa")
toplot$beatiful_sp <- str_replace(toplot$beatiful_sp, "Yeast" , "Saccharomyces cerevisiae")


```


### Sanity check - if the clusterisation was done correctly

```{r}

ggplot(toplot, aes(interaction(ploidy_clust, sp), bp_exp))+
  geom_point(aes( colour=sp ), size =1, alpha = 0.1, position=position_jitter(height=0, width=0.2))+
  geom_boxplot( alpha = 0)+
  theme_bw()+theme(axis.text.x = element_blank())+
  facet_wrap(~suns_for_clust)+
   scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))

```



## Coefficients of regressions +/- SE, bootstrap-like ALL control species - i.e. KINETO regressions





```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg <- ggplot(test_kinetos_all_clust1, aes( suns_for_clust , coeficcient) )+ 
  geom_violin()+ 
  geom_boxplot(alpha =0.2, width = 0.2)+
  #geom_point(aes(colour = species_out),size =4, alpha = 0.5,  position=position_jitter(height=0, width=0.2, seed = 123) )+
   geom_pointrange(aes(ymin=as.numeric(coeficcient-SE), ymax=as.numeric(coeficcient+SE), shape = test_species , colour = slidedate),  linewidth=1, alpha=0.5, position=position_jitter(height=0, width=0.2, seed = 123))+
  theme_linedraw()+theme(axis.text.x = element_text(angle =90))+
   theme(legend.position = "none")+
   ggtitle("Coefficients of regressions +/- SE (colour = slidedate)")
 
 print(gg)
 
##save the same plot, but with the legend - to spot suspicious slides/species 
 gg <- ggplot(test_kinetos_all_clust1, aes( suns_for_clust , coeficcient) )+ geom_violin()+ geom_boxplot(alpha =0.2, width = 0.2)+
  #geom_point(aes(colour = species_out),size =4, alpha = 0.5,  position=position_jitter(height=0, width=0.2, seed = 123) )+
   geom_pointrange(aes(ymin=as.numeric(coeficcient-SE), ymax=as.numeric(coeficcient+SE), shape = test_species , colour = slidedate),  linewidth=1, alpha=0.5, position=position_jitter(height=0, width=0.2, seed = 123))+
  theme_linedraw()+theme(axis.text.x = element_text(angle =90))+
   #theme(legend.position = "bottom")+
   ggtitle("Coefficients of regressions +/- SE (colour = slidedate)")

# ggsave("coeff_single_plot_comp_KINETO.png", height = 20, width = 50, unit ="cm", gg)

```


## Coefficients of regressions +/- SE, bootstrap-like

```{r, echo=FALSE, message=FALSE, warning=FALSE}

subs <- toplot[!duplicated(toplot$bootstrap_setting), ]

subs$species_out <- str_replace(subs$bootstrap_setting, " .*", "")

 gg <- ggplot(subs, aes( suns_for_clust , coeficcient) )+ geom_violin()+ geom_boxplot(alpha =0.2, width = 0.2)+
  #geom_point(aes(colour = species_out),size =4, alpha = 0.5,  position=position_jitter(height=0, width=0.2, seed = 123) )+
   geom_pointrange(aes(ymin=as.numeric(coeficcient-SE), ymax=as.numeric(coeficcient+SE), shape = species_out , colour = slidedate),  linewidth=1, alpha=0.5, position=position_jitter(height=0, width=0.2, seed = 123))+
  theme_linedraw()+theme(axis.text.x = element_text(angle =90))+
   theme(legend.position = "none")+
   ggtitle("Coefficients of regressions +/- SE (colour = slidedate)")
 
 print(gg)
 
 
##save the same plot, but with the legend - to spot suspicious slides/species 
 gg <- ggplot(subs, aes( suns_for_clust , coeficcient) )+ geom_violin()+ geom_boxplot(alpha =0.2, width = 0.2)+
  #geom_point(aes(colour = species_out),size =4, alpha = 0.5,  position=position_jitter(height=0, width=0.2, seed = 123) )+
   geom_pointrange(aes(ymin=as.numeric(coeficcient-SE), ymax=as.numeric(coeficcient+SE), shape = species_out , colour = slidedate),  linewidth=1, alpha=0.5, position=position_jitter(height=0, width=0.2, seed = 123))+
  theme_linedraw()+theme(axis.text.x = element_text(angle =90))+
  # theme(legend.position = "bottom")+
   ggtitle("Coefficients of regressions +/- SE (colour = slidedate)")

ggsave("coeff_single_plot_comp_Quantile_bootstr.png", height = 20, width = 50, unit ="cm", gg)

```





## Efron's pseudo R-squared (how good is the fit?), for regeressiong the account ALL control species - i.e. KINETO regressions


```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(test_kinetos_all_clust1, aes( suns_for_clust , as.numeric(R2)) )+
  geom_violin()+
   geom_point(aes(colour = test_species), alpha = 0.5, position=position_jitter(width=0.2, height = 0))+
  theme_linedraw()+theme(axis.text.x = element_text(angle =90))+
   #theme(legend.position = "bottom")+
   ggtitle("Efron's pseudo R-squared, test KINETOS")

```




pseudo-R-squred for Bootstrap-like validations. In makes few sense - it is sensitive to the absolute genome sizes,
because it equals 1-(RSS/TSS), where RSS is residual sum of squares, and TSS is total sum of squares. Since TSS is computed from MEAN, adding species with very large genome (like DIPLO) increases MEAN disproportionately. 

Note the log scale and the fact that Diplo seems critical for the prediction

Uncomment if really interested

```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg <- ggplot(subs, aes( suns_for_clust , as.numeric(R2)) )+
  geom_violin()+
   geom_point(aes(colour = species_out), alpha = 0.5, position=position_jitter(width=0.2, height = 0))+
  theme_linedraw()+theme(axis.text.x = element_text(angle =90))+
   #theme(legend.position = "bottom")+
   ggtitle("Efron's pseudo R-squared, bootstrap-like (NOT RELIABLE)")

print(gg)
subs1 <-  subs[which(subs$R2 >0.15),]
 
gg <- ggplot(subs1[which(subs1$species_out != "Diplo"),], aes( suns_for_clust , as.numeric(R2)) )+
  geom_violin()+
   geom_point(aes(colour = species_out), alpha = 0.5, position=position_jitter(width=0.2, height = 0))+
  theme_linedraw()+theme(axis.text.x = element_text(angle =90))+
   #theme(legend.position = "bottom")+
   ggtitle("Efron's pseudo R-squared, filtered>0.15, no Diplo")
#print(gg)

```




## SE and p-values of regression fits (how good is the fit?), for regeressiong tha account ALL control species - i.e. KINETO regressions


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#test_kinetos_all_clust$merger1 <- paste(test_kinetos_all_clust$suns_for_clust, test_kinetos_all_clust$slidedate)
#test_kinetos_all_clust1 <- test_kinetos_all_clust[!duplicated(test_kinetos_all_clust$merger1), ]


ggplot(test_kinetos_all_clust1, aes( suns_for_clust , as.numeric(SE)) )+
  geom_violin()+
   geom_point(aes(colour = test_species), alpha = 0.5, position=position_jitter(width=0.2, height = 0))+
  theme_linedraw()+theme(axis.text.x = element_text(angle =90))+
   #theme(legend.position = "bottom")+
   ggtitle("SE, test KINETOS")
 
which(as.numeric(test_kinetos_all_clust1$pvalue) > 0.05)

ggplot(test_kinetos_all_clust1, aes( suns_for_clust , as.numeric(pvalue)) )+
  geom_violin()+
   geom_point(aes(colour = test_species), alpha = 0.5, position=position_jitter(width=0.2, height = 0))+
  theme_linedraw()+theme(axis.text.x = element_text(angle =90))+
   #theme(legend.position = "bottom")+
  #scale_y_continuous(labels = abs, limits = c(0, 100))+
   ggtitle("p-value, test KINETOS")

```


## SE and p-values and CIs of BOOTSTRAP-like regressions


```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(subs, aes( suns_for_clust , as.numeric(SE)) )+
  geom_violin()+
   geom_point(aes(colour = species_out), alpha = 0.5, position=position_jitter(width=0.2, height = 0))+
  theme_linedraw()+theme(axis.text.x = element_text(angle =90))+
   #theme(legend.position = "bottom")+
   ggtitle("SE, bootstrap-like test")

ggplot(subs, aes( suns_for_clust , as.numeric(pvalue)) )+
  geom_violin()+
   geom_point(aes(colour = species_out), alpha = 0.5, position=position_jitter(width=0.2, height = 0))+
  theme_linedraw()+theme(axis.text.x = element_text(angle =90))+
   #theme(legend.position = "bottom")+
   ggtitle("p-value, bootstrap-like test")


```











## filtering regressions by p-value, R2, SE, other factors, for bootstrap-like regressions and saving filtered df

```{r}

toplot <- toplot[!is.na(toplot$bp_exp_ploidy_corr),]

toplot <- toplot[ -intersect( which(toplot$pvalue > 0.05), which(toplot$validation == "BOOTSTRAP-LIKE_PREDICTED") ),]
#toplot <- toplot[ -(which(toplot$validation == "BOOTSTRAP-LIKE_PREDICTED") ),]


#toplot <- toplot[which(toplot$R2 >= 0.15),]
#toplot <- toplot[which(toplot$SE <= 0.1),]
#toplot <- toplot[!grep("", toplot),]
#unique(toplot$validation)

fwrite(toplot, "TOPLOT_bootstrap_out_clustered.csv")

```






## How close are the inferred medians to the expected counts?
## Mean(Predicted medians - expected_bp)/expected_bp
Done accounting pre-prediction clustering data.
When suns_for_clust is clustered - the predicted value is substracted from the expected value, and module(absolute value) of difference is used.
### !!! Two values are produced - normalised by total bp_exp (so different species with equal genome sizes are comparable), and non-normalised. Normalised is taken for subsequent comparions.
Non-normalised was used in an older version. If you want to change back to it change "metrics_value" in tmp dataframe from diff_big_norm ==> diff_big


```{r, include=FALSE}
MEDIAN_summary <- data.frame()

for(sp in unique(toplot$sp)){
  
  df <- toplot[toplot$sp == sp,]
  actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  
  df1 <- df[df$if_test_sp=="TEST_bootstrap",]
  #df1$for_medians <- paste(df1$suns_for_clust, df1$slidedate)
  df1 <- df1[ which(is.na(df1$bp_exp_ploidy_corr) == FALSE), ]
  
  for (setting in unique(df1$suns_for_clust)) {
    df3 <- df1[df1$suns_for_clust == setting,]
  for (ploidy in unique(df3$ploidy_clust)) {
    df2 <- df3[df3$ploidy_clust == ploidy,]
    diff_big <- c()
    diff_big_norm <- c()
  for (i in unique(df2$slidedate)) {
    if (nrow(df2[df2$slidedate==i,]) == 0 ) next 
    
    med <- median(df2$bp_exp_ploidy_corr[df2$slidedate==i])
  
    if (ploidy == "1n") { 
      diff <- min(actual_pos) - med 
      diff_norm <-(min(actual_pos) - med)/min(actual_pos) } else 
    if (ploidy == "2n") { 
      diff <- med - max(actual_pos) 
      diff_norm <- (med - max(actual_pos))/max(actual_pos) } else 
    if (ploidy == "Undefined") {
    if (med < min(actual_pos)) {
      diff <-  min(actual_pos) - med 
      diff_norm <-  (min(actual_pos) - med)/min(actual_pos) } 
    else if(med > max(actual_pos)){
      diff <- med - max(actual_pos)  
      diff_norm <- (med - max(actual_pos))/max(actual_pos) }
    else if(med > min(actual_pos) & med < max(actual_pos)) {
      diff <- mean( abs(min(actual_pos) - med), abs(med - max(actual_pos)) )
      diff_norm <- mean( abs(min(actual_pos) - med), abs(med - max(actual_pos)) )/mean((min(actual_pos)),(max(actual_pos))) } 
    }
    # for debug purposes
    #print(ploidy)
    #print(diff_norm)
    diff_big <- c(diff_big, diff)
    diff_big_norm <- c(diff_big_norm, diff_norm)
  }
    #print(ploidy)
    #print(diff_big_norm)
    ################### NORMALISING BY THE NUMBER OF DATAPOINTS 
    
    
   tmp <-   data.frame("metrics" = "Mean(|Predicted medians - expected_bp|)/bp_exp",
                       "suns_for_clust" = setting,
                        "sp" = sp,
                         "metrics_value" = mean(abs(diff_big_norm)) )
   MEDIAN_summary <- rbind(MEDIAN_summary, tmp)
  } }
}


```

### corresponding plots

```{r, echo=FALSE, message=FALSE, warning=FALSE}
datatable(MEDIAN_summary)

for(sp in unique(toplot$sp)){
  
  df <- toplot[toplot$sp == sp,]
    actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  
  df1 <-df[df$if_test_sp=="TEST_bootstrap",]
  df1$for_medians <- paste(df1$suns_for_clust, df1$exp_series_date)
  
  # df1 <- df1[ df1$ploidy_clust == "1n", ]
  
  gg <- ggplot(df1[df1$sp==sp,], aes(for_medians, bp_exp_ploidy_corr ) )+
    #geom_boxplot(aes(fill = as.factor(exp_series_date)) )+#geom_point(alpha = 0.1)+
    stat_summary(fun = median, geom = "point", aes(colour = ploidy_clust)) +
     theme_bw()+facet_wrap(~suns_for_clust, scales = "free_x")+
    geom_hline( aes(yintercept = actual_pos[1]), colour ="red" )+ geom_hline( aes(yintercept = actual_pos[2]), colour ="red" )+
  theme(axis.text.x = element_blank())+
     ggtitle(paste(sp, "medians for one experimental series"))
  
  print(gg)

}




for(sp in unique(toplot$sp)){
  
  df <- toplot[toplot$sp == sp,]
    actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  
  df1 <-df[df$if_test_sp=="TEST_bootstrap",]
  df1$for_medians <- paste(df1$suns_for_clust, df1$slidedate)
  
  # df1 <- df1[ df1$ploidy_clust == "1n", ]
  
  gg <- ggplot(df1[df1$sp==sp,], aes(for_medians, bp_exp_ploidy_corr ) )+
    #geom_boxplot(aes(fill = as.factor(exp_series_date)) )+#geom_point(alpha = 0.1)+
    stat_summary(fun = median, geom = "point", aes(colour = ploidy_clust)) +
     theme_bw()+facet_wrap(~suns_for_clust, scales = "free_x")+
    geom_hline( aes(yintercept = actual_pos[1]), colour ="red" )+ geom_hline( aes(yintercept = actual_pos[2]), colour ="red" )+
  theme(axis.text.x = element_blank())+
     ggtitle(paste(sp, "medians for one slide"))
  
  print(gg)

}


which(is.na(toplot$suns_for_clust) ==TRUE)

```

## How close are the inferred bp counts to the expected counts?
## |(bp_expected - bp_predicted)|/bp_exp

```{r, include=FALSE}
Pred_bp_delta_summary <- data.frame("metrics" = NA,
                               "suns_for_clust" = NA,
                               "sp" = NA,
                               "metrics_value" = NA)

for(sp in unique(toplot$sp)){
  
  df <- toplot[toplot$sp == sp,]
  actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  
  df1 <- df[df$if_test_sp=="TEST_bootstrap",]
  #df1$for_medians <- paste(df1$suns_for_clust, df1$slidedate)
  df1 <- df1[ which(is.na(df1$bp_exp_ploidy_corr) == FALSE), ]
  
  for (setting in unique(df1$suns_for_clust)) {
  for (ploidy in unique(df1$ploidy_clust)) {
    df2 <- df1[df1$suns_for_clust == setting,]
    df2 <- df2[df2$ploidy_clust == ploidy,]
    
    diff_big <- mean(abs(df2$PERCENTdiff_from_actual))
    
   tmp <-   data.frame("metrics" = "|(bp_expected - bp_predicted)|/bp_exp",
                       "suns_for_clust" = setting,
                        "sp" = sp,
                         "metrics_value" = diff_big )
   Pred_bp_delta_summary <- rbind(Pred_bp_delta_summary, tmp)
  } }
}

```

### corresponding plots

```{r, echo=FALSE, message=FALSE, warning=FALSE}
datatable(Pred_bp_delta_summary)

df <- toplot[toplot$if_test_sp == "TEST_bootstrap",]

gg <- ggplot(df, aes( PERCENTdiff_from_actual*100 ) )+geom_histogram(aes(fill =sp ), alpha = 0.4)+
     theme_bw()+facet_wrap(~suns_for_clust)+
  #theme(legend.position = "none")+
     ggtitle("% difference of predicted bp counts from actual expected genome size")

print(gg)

gg <- ggplot(df, aes(sp, PERCENTdiff_from_actual*100 ) )+geom_violin()+geom_point(aes(colour =file_prefix ), alpha = 0.1)+
     theme_bw()+facet_wrap(~suns_for_clust)+
  theme(legend.position = "none")+
     ggtitle("% difference of predicted bp counts from actual expected genome size")

print(gg)

gg <- ggplot(df[df$sp != "Yeast",], aes( PERCENTdiff_from_actual*100 ) )+geom_histogram(aes(fill =sp ), alpha = 0.4)+
     theme_bw()+facet_wrap(~suns_for_clust)+
  #theme(legend.position = "none")+
     ggtitle("% difference of predicted bp counts from actual expected genome size, no Yeast")

print(gg)

gg <- ggplot(df[df$sp != "Yeast",], aes(sp, PERCENTdiff_from_actual*100 ) )+geom_violin()+geom_point(aes(colour =file_prefix ), alpha = 0.1)+
     theme_bw()+facet_wrap(~suns_for_clust)+
  theme(legend.position = "none")+
     ggtitle("% difference of predicted bp counts from actual expected genome size, no Yeast")

print(gg)



```

### corresponding plots 

```{r, echo=FALSE, message=FALSE, warning=FALSE}

plotlist <- list()
for(sp in unique(toplot$sp) ){

  df <- toplot[toplot$sp == sp,]
  #
    #df <- df[df$ploidy_clust == "1n",]
    actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  
    df <- df[df$if_test_sp == "TEST_bootstrap",]
df <- df[!is.na(df$sp) & !is.na(df$suns_for_clust),]
   gg <- ggplot(df, aes( bp_exp_ploidy_corr ) )+geom_histogram(aes(fill = ploidy_clust), bins = 30, alpha = 0.4, position = "identity")+
     theme_bw()+geom_vline( aes(xintercept = actual_pos[1]) )+ geom_vline( aes(xintercept = actual_pos[2]) )+
    scale_x_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))+facet_wrap(~suns_for_clust)+
  theme(legend.position = "none")+
     ggtitle(sp)
  
print(gg)
  plotlist[[length(plotlist) + 1]] <- gg
 
}

#type of visualisation that might be interesting but not reallly essential - uncomment if needed
#plotlist <- list()
#for(sp in unique(toplot$sp) ){
#  df <- toplot[toplot$sp == sp,]
#    actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
#    df <- df[df$if_test_sp == "TEST_bootstrap",]
#   gg <- ggplot(df, aes( diff_from_actual ) )+geom_histogram(aes(fill = validation), alpha = 0.4, position = "identity")+
#     theme_bw()+geom_vline( aes(xintercept = actual_pos[1]) )+ geom_vline( aes(xintercept = actual_pos[2]) )+
#    scale_x_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))+facet_wrap(~suns_for_clust)+
#  theme(legend.position = "none")+
#     ggtitle(sp)
#print(gg)
# plotlist <- list(plotlist, gg)
#}

#unique(toplot$ploidy_clust)

#

```


## PLOTS FOR MANUSCRIPT


```{r}

plotlist <- list()
for(sp in unique(toplot$beatiful_sp) ){

  df <- toplot[toplot$beatiful_sp == sp,]
  #
    #df <- df[df$ploidy_clust == "1n",]
    actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  
    df <- df[df$if_test_sp == "TEST_bootstrap",]
df <- df[!is.na(df$sp) & !is.na(df$beatiful_suns),]


   gg <- ggplot(df, aes( bp_exp_ploidy_corr ) )+geom_histogram(aes(fill = ploidy_clust), bins = 30, alpha = 0.4, position = "identity")+
     theme_bw()+    scale_x_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
     geom_vline(xintercept = actual_pos[1])+
     geom_vline(xintercept = actual_pos[2])+
     facet_wrap(~beatiful_suns)+
  theme(legend.position = "none")+
     xlab("predicted genome size, bp")+ylab("number of cells")+
     ggtitle(sp)
  
#print(gg)
  plotlist[[length(plotlist) + 1]] <- gg
 
}


ggexport(plotlist = plotlist, filename = "distribution_colntrols_suppl.pdf", ncol = 1, height = 20)

```



```{r}

plotlist <- list()
for(sp in unique(toplot$beatiful_sp) ){

  df <- toplot[toplot$beatiful_sp == sp,]

  df <- df[df$suns_for_clust == "est_sp  ALL",]
  #
    #df <- df[df$ploidy_clust == "1n",]
    actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  
    df <- df[df$if_test_sp == "TEST_bootstrap",]
    
  #  df <- df[!is.na(df$sp) & !is.na(df$suns_for_clust),]

   gg <- ggplot(df, aes( bp_exp_ploidy_corr ) )+geom_histogram(aes(fill = ploidy_clust), bins = 30, alpha = 0.4, position = "identity")+
     theme_bw()+ scale_x_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
     geom_vline(xintercept = actual_pos[1] )+ 
     geom_vline(xintercept = actual_pos[2] )+
  theme(legend.position = "none")+
     xlab("predicted genome size, bp")+ylab("number of cells")+
     ggtitle(sp)
  
#print(gg)

 plotlist[[length(plotlist) + 1]] <- gg

}


ggexport(plotlist = plotlist, filename = "distribution_colntrols_main.pdf", nrow = 2, ncol = 2)

```



```{r}


median_summary_control <- predictedbp_median_estimator(data.frame(toplot[toplot$suns_for_clust == "est_sp  ALL" & toplot$if_test_sp == "TEST_bootstrap",]), subset_for_median_estimation = "ploidy"  )
median_summary_control$sp <- str_replace(median_summary_control$sp, "nucl_df", "")

tmp <- toplot[toplot$validation == "ACTUAL", c("sp", "bp_exp_ploidy_corr", "ploidy_clust", "beatiful_sp")]
tmp <- tmp[!duplicated(tmp),]



median_summary_control_act <- median_summary_control %>% full_join(tmp) 

median_summary_control_act <- median_summary_control_act[median_summary_control_act$ploidy_clust != "Undefined",]
median_summary_control_act$ploidy_clust <- str_replace(median_summary_control_act$ploidy_clust, "1n", "G1")
median_summary_control_act$ploidy_clust <- str_replace(median_summary_control_act$ploidy_clust, "2n", "G2")

#adding means of medians for each species - so far, done because this is a way to obtain a single estimate of the genome size
median_summary_control_act <- median_summary_control_act %>% group_by(beatiful_sp, ploidy_clust) %>%  mutate(mean_MEDIAN_bp_exp = mean(MEDIAN_bp_exp, na.rm = TRUE))%>%ungroup()


gg_mainB <- ggplot(median_summary_control_act, aes(factor(beatiful_sp, 
                                              levels = c("Saccharomyces cerevisiae", "Capsaspora owczarzaki","Salpingoeca rosetta", "Diplonema papillatum" )), (MEDIAN_bp_exp-bp_exp_ploidy_corr)/bp_exp_ploidy_corr) )+
    #geom_boxplot(aes(fill = as.factor(exp_series_date)) )+#geom_point(alpha = 0.1)+
    geom_violin( ) +
   geom_point(alpha=0.25 ) +
     geom_point( aes( x=beatiful_sp, y = (mean_MEDIAN_bp_exp-bp_exp_ploidy_corr)/bp_exp_ploidy_corr ), colour = "coral2", shape =8, size=3 ) +

#   geom_pointrange( aes( 
#   ymin= (MEDIAN_bp_exp-as.numeric(mean_lower)-bp_exp_ploidy_corr)/bp_exp_ploidy_corr,
#  ymax= (MEDIAN_bp_exp+as.numeric(mean_higher)-bp_exp_ploidy_corr)/bp_exp_ploidy_corr ), 
#                 colour = "coral2", shape =8)+
 
  # geom_line( aes(group=exp_series_date), alpha=0.15)+
  facet_wrap(~ploidy_clust, scales = "free_y", ncol=1)+
     theme_bw()+theme(axis.text.x = element_text(angle =45, hjust = 1))+
     ylab("difference of predicted median from actual genome size")+
     xlab("Control species predicted")



ggsave(gg_mainB, filename = "median_distr_controls_main.pdf")


```


```{r}
library(patchwork)


figA <- (plotlist[[3]] | plotlist[[1]]) /
        (plotlist[[4]] | plotlist[[2]])
gg_mainB <- gg_mainB + theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.2), "cm"))

final_fig <- figA / gg_mainB +
  plot_layout(widths = c(1.3, 1))  
  #plot_annotation(tag_levels = c("A", "B"))

ggsave("controls_main.pdf", final_fig,  height = 10)

```



```{r}

plotlist <- list()

for(suns in unique(toplot$suns_for_clust)){
  
suns1 <- str_replace(suns, "est_sp .* ", "ploidy")
suns1 <- str_replace(suns1, "NOT CLUSTERED ", "ploidy")

suns1 <- str_replace(suns1, "slide", "__slidedate")
suns1 <- str_replace(suns1, "series", "__exp_series_date")
suns1 <- str_replace(suns1, "ALL", "")
  
print(c(suns, suns1))

median_summary_control <- predictedbp_median_estimator(data.frame(toplot[toplot$suns_for_clust == suns & toplot$if_test_sp == "TEST_bootstrap",]), subset_for_median_estimation = suns1 , BS_replicates = 100  )
median_summary_control$sp <- str_replace(median_summary_control$sp, "nucl_df", "")

tmp <- toplot[toplot$validation == "ACTUAL", c("sp", "bp_exp_ploidy_corr", "ploidy_clust", "beatiful_sp")]
tmp <- tmp[!duplicated(tmp),]



median_summary_control_act <- median_summary_control %>% full_join(tmp) 
if(length(grep("NOT CLUSTERED", suns)) ==0) {
  median_summary_control_act <- median_summary_control_act[median_summary_control_act$ploidy_clust != "Undefined",] }
else {median_summary_control_act <- median_summary_control_act[median_summary_control_act$ploidy_clust == "Undefined",] }
median_summary_control_act$ploidy_clust <- str_replace(median_summary_control_act$ploidy_clust, "1n", "G1")
median_summary_control_act$ploidy_clust <- str_replace(median_summary_control_act$ploidy_clust, "2n", "G2")

#adding means of medians for each species - so far, done because this is a way to obtain a single estimate of the genome size
median_summary_control_act <- median_summary_control_act %>% group_by(beatiful_sp, ploidy_clust) %>%  mutate(mean_MEDIAN_bp_exp = mean(MEDIAN_bp_exp, na.rm = TRUE))%>%ungroup()


gg <- ggplot(median_summary_control_act, aes(factor(beatiful_sp, 
                                              levels = c("Saccharomyces cerevisiae" , "Capsaspora owczarzaki","Salpingoeca rosetta", "Diplonema papillatum" )), (MEDIAN_bp_exp-bp_exp_ploidy_corr)/bp_exp_ploidy_corr) )+
    #geom_boxplot(aes(fill = as.factor(exp_series_date)) )+#geom_point(alpha = 0.1)+
    geom_violin( ) +
   geom_point(alpha=0.25 ) +
     geom_point( aes( x=beatiful_sp, y = (mean_MEDIAN_bp_exp-bp_exp_ploidy_corr)/bp_exp_ploidy_corr ), colour = "red", shape =18, size=1 ) +
 # geom_line( aes(group=exp_series_date), alpha=0.15)+
  facet_wrap(~ploidy_clust, scales = "free_x")+
     theme_bw()+
  theme(axis.text.x = element_text(angle =45, hjust = 1),
         axis.text = element_text(size = 7),
    axis.title = element_text(size = 7))+
     ylab("predicted median VS actual genome size")+
     xlab("Control species predicted")+
  ggtitle(toplot$beatiful_suns[toplot$suns_for_clust == suns])+scale_y_continuous(trans = "asinh")

gg
  plotlist[[length(plotlist) + 1]] <- gg
 
}


ggexport(plotlist = plotlist, filename = "median_distr_controls_suppl.pdf", ncol = 4, nrow =ceiling(length(plotlist) / 3), height = 15, width = 9)

#for (i in length(plotlist)) {  print(plotlist[i])}

```









## KS and median tests comparing predicted counts for each slide WITHIN different regressions - "How consistent are the results of each regression method when applied to different datasets(=SLIDES)?"

modified here - KS-test now compares distributions of "1n" and "2n" clusters separately.
```{r, include=FALSE}
pval_summary_within <- ks_medians_estimator_within(toplot)
```




GENERATING INTERMEDIATE PLOTS - uncomment content of chunk below upon need

```{r}
#ks_medians_intermediate_plots_within(pval_summary_within)
```


## Median test, how uniform are predicted count medians for each setting when comparing predictions for different slides 
## number of p-values <=0.05 ("significant") / total number of sets compared for each combination of settings + CORRESPONDING PLOTS

```{r echo=FALSE, message=FALSE, warning=FALSE}

#########Actual summary plot
pval_summary_within$comp <- paste(pval_summary_within$X1, "//VS//", pval_summary_within$X2)

pval_summary_within$median_signif <- "p-value > 0.05"
pval_summary_within$median_signif[pval_summary_within$Median_test_p_value <= 0.05] <- "p-value <= 0.05"

median_summaries <- data.frame("subset" = unique(pval_summary_within$subset), "p_val_signif" = NA, "p_val_Insignif" = NA, "significance_norm" = NA )
subset_split <- str_split(median_summaries$subset,  " ", simplify=TRUE)
median_summaries$ploidy <- subset_split[,1]
median_summaries$sp <- subset_split[,2]
median_summaries$reg_across <- subset_split[,5]
median_summaries$clustering_setting <- paste(subset_split[,1], subset_split[,3], subset_split[,4])

for (i in unique(pval_summary_within$subset)) {

  df <- pval_summary_within[pval_summary_within$subset == i, ]  

  median_summaries$p_val_Insignif[median_summaries$subset ==i] <- nrow(df[df$median_signif == "p-value > 0.05",])
  median_summaries$p_val_signif[median_summaries$subset ==i] <- nrow(df[df$median_signif == "p-value <= 0.05",])
  median_summaries$significance_norm[median_summaries$subset ==i] <- nrow(df[df$median_signif == "p-value <= 0.05",])/nrow(df)
}
 

symmetry <- median_summaries

ggplot(rbind(median_summaries, symmetry), mapping=aes(clustering_setting, sp, fill = as.numeric(significance_norm )) ) + 
  geom_tile( alpha = 0.8)+ 
   geom_text( aes(label=round(significance_norm, digits = 3)), size = 2 )+ 
  theme(axis.text.x = element_text(angle =90))+
    scale_fill_gradientn(colors = hcl.colors(20, "Cyan-Magenta"))+
  theme_classic()+ theme(axis.text.x = element_text(angle =90))+ggtitle(paste("Median test, p-value <=0.05/total number of slides+species compared"))+
  facet_grid(~reg_across)

testpval_MEDIAN_summary <- data.frame("metrics" = "Median test, p-value <=0.05/total number of slides+species compared",
                       "suns_for_clust" = paste(median_summaries$clustering_setting, median_summaries$reg_across),
                        "sp" = median_summaries$sp,
                         "metrics_value" = median_summaries$significance_norm )
 
datatable(testpval_MEDIAN_summary)
```

supporting slide-by-slide plots, uncomment if really interested
```{r, include=FALSE}

pval_summary_within$comp <- paste(pval_summary_within$X1, "//VS//", pval_summary_within$X2)
pval_summary_within$median_signif <- "p-value > 0.05"
pval_summary_within$median_signif[pval_summary_within$Median_test_p_value <= 0.05] <- "p-value <= 0.05"

for (i in unique(pval_summary_within$subset)) {
  
  df <- pval_summary_within[pval_summary_within$subset ==i, ]
  
  gg <-  ggplot(df, aes(X1, X2, fill = as.numeric(Median_D )) ) +  geom_tile( alpha = 0.8)+ 
   geom_text( aes(label=round(Median_test_p_value, digits = 3), colour = as.factor(median_signif)) )+ 
  theme(axis.text.x = element_text(angle =90))+
    scale_fill_gradientn(colors = hcl.colors(20, "Cyan-Magenta"))+
  theme_classic()+ theme(axis.text.x = element_blank(), axis.text.y = element_blank())+ggtitle(paste("Median test", i))
  print(gg)
}  
########

```


## KS test, how uniform are predicted count distributions for each setting when comparing predictions for different slides 
## number of p-values <=0.05 ("significant") / total number of sets compared for each combination of settings  + CORRESPONDING PLOTS

```{r, echo=FALSE, message=FALSE, warning=FALSE}


#########Actual summary plot
pval_summary_within$comp <- paste(pval_summary_within$X1, "//VS//", pval_summary_within$X2)

pval_summary_within$ks_signif <- "p-value > 0.05"
pval_summary_within$ks_signif[pval_summary_within$p.value <= 0.05] <- "p-value <= 0.05"

ks_summaries <- data.frame("subset" = unique(pval_summary_within$subset), "p_val_signif" = NA, "p_val_Insignif" = NA, "significance_norm" = NA )
subset_split <- str_split(ks_summaries$subset,  " ", simplify=TRUE)

ks_summaries$ploidy <- subset_split[,1]
ks_summaries$sp <- subset_split[,2]
ks_summaries$reg_across <- subset_split[,5]
ks_summaries$clustering_setting <- paste(subset_split[,1], subset_split[,3], subset_split[,4])

for (i in unique(pval_summary_within$subset)) {

  df <- pval_summary_within[pval_summary_within$subset == i, ]  

  ks_summaries$p_val_Insignif[ks_summaries$subset ==i] <- nrow(df[df$ks_signif == "p-value > 0.05",])
  ks_summaries$p_val_signif[ks_summaries$subset ==i] <- nrow(df[df$ks_signif == "p-value <= 0.05",])
  ks_summaries$significance_norm[ks_summaries$subset ==i] <- nrow(df[df$ks_signif == "p-value <= 0.05",])/nrow(df)
}
 

symmetry <- ks_summaries

ggplot(rbind(ks_summaries, symmetry), mapping=aes(clustering_setting, sp, fill = as.numeric(significance_norm )) ) + 
  geom_tile( alpha = 0.8)+ 
   geom_text( aes(label=round(significance_norm, digits = 3)), size =2 )+ 
  theme(axis.text.x = element_text(angle =90))+
    scale_fill_gradientn(colors = hcl.colors(20, "Cyan-Magenta"))+
  theme_classic()+ theme(axis.text.x = element_text(angle =90))+ggtitle(paste("KS test, p-value <=0.05/total number of slides+species compared"))+
  facet_grid(~reg_across)
 
testpval_KS_summary <- data.frame("metrics" = "KS test, p-value <=0.05/total number of slides+species compared",
                       "suns_for_clust" = paste(ks_summaries$clustering_setting, ks_summaries$reg_across),
                        "sp" = ks_summaries$sp,
                         "metrics_value" = ks_summaries$significance_norm )
 
datatable(testpval_KS_summary)

```


 supporting slide-by-slide plots, uncomment if really interested
```{r, include=FALSE}
pval_summary_within$comp <- paste(pval_summary_within$X1, "//VS//", pval_summary_within$X2)
pval_summary_within$ks_signif <- "p-value > 0.05"
pval_summary_within$ks_signif[pval_summary_within$p.value <= 0.05] <- "p-value <= 0.05"

for (i in unique(pval_summary_within$subset)) {
  
  df <- pval_summary_within[pval_summary_within$subset ==i, ]
  
  gg <-  ggplot(df, aes(X1, X2, fill = as.numeric(Median_D )) ) +  geom_tile( alpha = 0.8)+ 
   geom_text( aes(label=round(p.value, digits = 3), colour = as.factor(ks_signif)) )+ 
  theme(axis.text.x = element_text(angle =90))+
    scale_fill_gradientn(colors = hcl.colors(20, "Cyan-Magenta"))+
  theme_classic()+ theme(axis.text.x = element_blank(), axis.text.y = element_blank())+ggtitle(paste("Median test", i))
  print(gg)
}  
########
```



## Comparing 95% Confidence Intervals of predictions
## Mean(Highest possible predicted bp_exp - Lowest possible predicted bp_exp)

```{r, echo=FALSE, message=FALSE, warning=FALSE}

CI_summary <- data.frame()

for(sp in unique(toplot$sp)){

  df <- toplot[toplot$sp == sp,]
  
  actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  df1 <- df[df$if_test_sp=="TEST_bootstrap",]

  #df1$for_medians <- paste(df1$suns_for_clust, df1$slidedate)
  df1 <- df1[ which(is.na(df1$bp_exp_ploidy_corr) == FALSE), ]
  
  for (setting in unique(df1$suns_for_clust)) {
   df3 <- df1[df1$suns_for_clust == setting,]
  # diff_big <- c()
  for (ploidy in unique(df3$ploidy_clust)) {
    
    df2 <- df3[df3$ploidy_clust == ploidy,]
    
    mean <- mean(as.numeric(df2$higher)-as.numeric(df2$lower) , na.rm = TRUE)
    med <- median( c(as.numeric(df2$higher), as.numeric(df2$lower)) , na.rm = TRUE)
    
      if (ploidy == "1n") { 
      diff_norm <- mean/min(actual_pos) } else 
    if (ploidy == "2n") { 
      diff_norm <- mean/max(actual_pos) } else 
    if (ploidy == "Undefined") {
    if (med < min(actual_pos)) {
      diff_norm <-  mean/min(actual_pos) } 
    else if(med > max(actual_pos)){
      diff <- med - max(actual_pos)  
      diff_norm <- (mean - max(actual_pos))/max(actual_pos) }
    else if(med > min(actual_pos) & med < max(actual_pos)) {
      diff <- mean( abs(min(actual_pos) - mean), abs(mean - max(actual_pos)) )
      diff_norm <- mean( abs(min(actual_pos) - mean), abs(mean - max(actual_pos)) )/mean((min(actual_pos)),(max(actual_pos))) } 
    }
    

    
   #diff_big <- c(diff_big, diff_norm)
    
    #for debug
    # print(paste(setting, sp, ploidy))
    # print(as.numeric(df2$higher)-as.numeric(df2$lower))
    
   tmp <-   data.frame("metrics" = "CI = mean(upper-lower)/bp_exp",
                       "suns_for_clust" = paste(ploidy, setting),
                        "sp" = sp,
                         "metrics_value" = diff_norm )
   CI_summary <- rbind(CI_summary, tmp)
 } }
}



datatable(CI_summary)
```

## corresponding plots ---- Comparing 95% Confidence Intervals of predictions
## Mean(Highest possible predicted bp_exp - Lowest possible predicted bp_exp)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
gg <- ggplot(toplot, aes( bp_exp_ploidy_corr,nucl_sum_int, colour = sp,) )+geom_pointrange(aes(xmin = as.numeric(lower), xmax =  as.numeric(higher)),  alpha = 0.1, size=0.05, linewidth=2)+
     theme_bw()+facet_wrap(~suns_for_clust)+ 
     scale_x_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
  scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))+
   #scale_shape_manual(values=c())+
     ggtitle("CI full")
  
print(gg)


#gg <- ggplot(toplot, aes( bp_exp_ploidy_corr,nucl_sum_int) )+geom_pointrange(aes(xmin = as.numeric(lower), xmax =  as.numeric(higher)),  alpha = 0.2, size=0.05, linewidth=1)+
#     theme_bw()+facet_wrap(~suns_for_clust+sp, scales = "free_x", ncol= length(unique(toplot$sp)))+ 
#     scale_x_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
#  scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))+
   #scale_shape_manual(values=c())+
#     ggtitle("Ci full")
  
#print(gg)
#ggsave("CI_values_all_Quantile_bootstr.pdf", gg, width = 15, height =15)

toplot$CI_diff <- as.numeric(toplot$higher) - as.numeric(toplot$lower)
toplot$CI_diff <- log2( as.numeric(toplot$higher) ) - log2(  as.numeric(toplot$lower) )

gg <- ggplot(toplot, aes( suns_for_clust, CI_diff) )+
  geom_violin(alpha = 0.2)+
  geom_point(size =0.3, alpha = 0.3, position=position_jitter(height=0, width=0.2))+
     theme_bw()+facet_wrap(~sp, scales = "free_x", ncol= length(unique(toplot$sp)))+ 
  scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
    theme(axis.text.x = element_text(angle =90))+
     ggtitle("log2(higher predicted estimate) - log2(lower predicted estimate) !! ON logarithmic scale in addition")
  
print(gg)
#ggsave("deltas_of_CI_values_all_Quantile_bootstr.pdf", gg, width = 10, height =5)

```


## Uniformity of slopes (=coefficients) WITHIN settings for each regression_clustering setting 
## number of p-values <=0.05 ("significant") / total number of sets compared for each setting, (!!!) between different species (!!!) + CORRESPONDING PLOTS (1 set = same species predicted across X slides)
### Anova+Tukey for differences between coefficients between different species, WITHIN 1 clustering setting   

## How consistent is the coefficient of every setting depending on species+ CORRESPONDING PLOTS


```{r, echo=FALSE, message=FALSE, warning=FALSE}

subs <- toplot[!duplicated(toplot$bootstrap_setting), ]

subs$species_out <- str_replace(subs$bootstrap_setting, " .*", "")

coef_within <- anova_coef_within(subs)

coef_within$tuk_signif <- "p-value > 0.05"
coef_within$tuk_signif[coef_within$p.adj <= 0.05] <- "p-value <= 0.05"

ggplot(coef_within, aes(p.adj, comb, colour =comb, size = diff*10))+geom_point(alpha =0.5)+
  facet_wrap(~suns_for_clust)+theme_bw()+
  theme(axis.text.y = element_blank())+geom_vline(xintercept = 0.05, colour = "red")+
  ggtitle("p-values for Anova of quantile regression coefficients")


coeff_summary <- data.frame("metrics" = NA,
                               "suns_for_clust" = NA,
                               "sp" = NA,
                               "metrics_value" = NA)



  for (setting in unique(coef_within$suns_for_clust)) {
  df2 <- coef_within[coef_within$suns_for_clust == setting,]
    
  diff_big <- nrow(df2[df2$tuk_signif == "p-value <= 0.05",])/nrow(df2)
    
   tmp <-   data.frame("metrics" = "number of p-values <=0.05 / total number of sets compared (1 set = same species predicted across X slides)",
                       "suns_for_clust" = setting,
                        "sp" = "ALL",
                         "metrics_value" = diff_big )
   coeff_summary <- rbind(coeff_summary, tmp)
  }


datatable(coeff_summary)

```






## making a dataframe for summarising how good is every regression

```{r}

goodness_summary <- na.omit(rbind(Pred_bp_delta_summary, MEDIAN_summary, testpval_MEDIAN_summary, testpval_KS_summary, CI_summary, coeff_summary))

goodness_summary_R <- reshape(goodness_summary,
                   idvar = c("metrics", "suns_for_clust"),
                   timevar = "sp",
                   direction = "wide")
fwrite(goodness_summary_R, "goodness_summary_R.csv")


datatable(goodness_summary_R)
```




# THE EXTENT OF DIFFERENCES BETWEEN DIFFERENT METHODS

## KS and median tests comparing predicted counts for each slide BETWEEN different regressions

## "How different are different regression methods?"

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#
pval_summary <- ks_medians_estimator_between(toplot)

```

GENERATING INTERMEDIATE PLOTS - uncomment content of chunk below upon need

```{r,include=FALSE}
#ks_medians_intermediate_plots_within(pval_summary_within)
```




## Median test Between settings for each slide 
##number of p-values \<=0.05 ("significant") / total number of sets compared for each combination of settings  + CORRESPONDING PLOTS

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#########Actual summary plot
pval_summary$comp <- paste(pval_summary$X1, "//VS//", pval_summary$X2)

pval_summary$median_signif <- "p-value > 0.05"
pval_summary$median_signif[pval_summary$Median_test_p_value <= 0.05] <- "p-value <= 0.05"

median_summaries <- data.frame("comp"=unique(pval_summary$comp), "X1"=NA, "X2"=NA, "p_val_signif" = NA, "p_val_Insignif" = NA, "significance_norm" = NA )
median_summaries$X1 <- str_split(median_summaries$comp,  " //VS// ", simplify=TRUE)[,1]
median_summaries$X2 <- str_split(median_summaries$comp,  " //VS// ", simplify=TRUE)[,2]

for (i in unique(pval_summary$comp)) {
 print(i)
  df <- pval_summary[pval_summary$comp == i, ]  

  median_summaries$p_val_Insignif[median_summaries$comp ==i] <- nrow(df[df$median_signif == "p-value > 0.05",])
  median_summaries$p_val_signif[median_summaries$comp ==i] <- nrow(df[df$median_signif == "p-value <= 0.05",])
  median_summaries$significance_norm[median_summaries$comp ==i] <- nrow(df[df$median_signif == "p-value <= 0.05",])/nrow(df)
}
 

symmetry <- median_summaries
symmetry$X1 <- median_summaries$X2
symmetry$X2 <- median_summaries$X1

ggplot(rbind(median_summaries, symmetry), aes(X1, X2, fill = as.numeric(significance_norm )) ) +  geom_tile( alpha = 0.8)+ 
   geom_text( aes(label=round(significance_norm, digits = 3)), size =2 )+ 
  theme(axis.text.x = element_text(angle =90))+
    scale_fill_gradientn(colors = hcl.colors(20, "Cyan-Magenta"))+
  theme_classic()+ theme(axis.text.x = element_text(angle =90))+ggtitle(paste("Median test, p-value <=0.05/total number of slides+species compared"))
  


testpval_between_MEDIAN_summary <- data.frame("metrics" = "Median test, p-value <=0.05/total number of slides+species compared",
                       "suns_for_clust_1" = median_summaries$X1,
                        "suns_for_clust_2" = median_summaries$X2,
                         "metrics_value" = median_summaries$significance_norm )
 
datatable(testpval_between_MEDIAN_summary)
 
```

supporting slide-by-slide plots, uncomment if really interested

```{r, include=FALSE}

pval_summary$comp <- paste(pval_summary$X1, "//VS//", pval_summary$X2)
pval_summary$median_signif <- "p-value > 0.05"
pval_summary$median_signif[pval_summary$Median_test_p_value <= 0.05] <- "p-value <= 0.05"

for (i in unique(pval_summary$subset)) {
  
  df <- pval_summary[pval_summary$subset ==i, ]
  
  gg <-  ggplot(df, aes(X1, X2, fill = as.numeric(Median_D )) ) +  geom_tile( alpha = 0.8)+ 
   geom_text( aes(label=round(Median_test_p_value, digits = 3), colour = as.factor(median_signif)) )+ 
  theme(axis.text.x = element_text(angle =90))+
    scale_fill_gradientn(colors = hcl.colors(20, "Cyan-Magenta"))+
  theme_classic()+ theme(axis.text.x = element_text(angle =90))+ggtitle(paste("Median test", i))
  print(gg)
}  
########
```



##More general summary plots for KS test Between settings for each slide ##number of p-values \<=0.05 ("significant") / total number of sets compared for each combination of settings

```{r, echo=FALSE, message=FALSE, warning=FALSE}

########

#########Actual summary plot
pval_summary$comp <- paste(pval_summary$X1, "//VS//", pval_summary$X2)

pval_summary$ks_signif <- "p-value > 0.05"
pval_summary$ks_signif[pval_summary$p.value <= 0.05] <- "p-value <= 0.05"

ks_summaries <- data.frame("comp"=unique(pval_summary$comp), "X1"=NA, "X2"=NA, "p_val_signif" = NA, "p_val_Insignif" = NA, "significance_norm" = NA )
ks_summaries$X1 <- str_split(ks_summaries$comp,  " //VS// ", simplify=TRUE)[,1]
ks_summaries$X2 <- str_split(ks_summaries$comp,  " //VS// ", simplify=TRUE)[,2]

for (i in unique(pval_summary$comp)) {
 print(i)
  df <- pval_summary[pval_summary$comp == i, ]  

  ks_summaries$p_val_Insignif[ks_summaries$comp ==i] <- nrow(df[df$ks_signif == "p-value > 0.05",])
  ks_summaries$p_val_signif[ks_summaries$comp ==i] <- nrow(df[df$ks_signif == "p-value <= 0.05",])
  ks_summaries$significance_norm[ks_summaries$comp ==i] <- nrow(df[df$ks_signif == "p-value <= 0.05",])/nrow(df)
}
 

symmetry <- ks_summaries
symmetry$X1 <- ks_summaries$X2
symmetry$X2 <- ks_summaries$X1

ggplot(rbind(ks_summaries, symmetry), aes(X1, X2, fill = as.numeric(significance_norm )) ) +  geom_tile( alpha = 0.8)+ 
   geom_text( aes(label=round(significance_norm, digits = 3)), size =3 )+ 
  theme(axis.text.x = element_text(angle =90))+
    scale_fill_gradientn(colors = hcl.colors(20, "Cyan-Magenta"))+
  theme_classic()+ theme(axis.text.x = element_text(angle =90))+ggtitle(paste("KS test, p-value <=0.05/total number of slides+species compared"))
  
testpval_between_KS_summary <- data.frame("metrics" = "KS test, p-value <=0.05/total number of slides+species compared",
                       "suns_for_clust_1" = ks_summaries$X1,
                        "suns_for_clust_2" = ks_summaries$X2,
                         "metrics_value" = ks_summaries$significance_norm )
 
datatable(testpval_between_KS_summary)
 
```

supporting slide-by-slide plots, uncomment if really interested

```{r,include=FALSE}
pval_summary$comp <- paste(pval_summary$X1, "//VS//", pval_summary$X2)

pval_summary$ks_signif <- "p-value > 0.05"
pval_summary$ks_signif[pval_summary$p.value <= 0.05] <- "p-value <= 0.05"

for (i in unique(pval_summary$subset)) {
  
  df <- pval_summary[pval_summary$subset ==i, ]
  
  gg <-  ggplot(df, aes(X1, X2, fill = as.numeric(statistic)) ) +  geom_tile( alpha = 0.8)+ 
   geom_text( aes(label=round(p.value, digits = 3), colour = as.factor(ks_signif)) )+ 
  theme(axis.text.x = element_text(angle =90))+
    scale_fill_gradientn(colors = hcl.colors(20, "Cyan-Magenta"))+
  theme_classic()+ theme(axis.text.x = element_text(angle =90))+ggtitle(paste("KS test", i))
  print(gg)
}  
```





## Coefficients between regressions

```{r, echo=FALSE, message=FALSE, warning=FALSE}

coef_between <- anova_coef_between(subs)

coef_between$tuk_signif <- "p-value > 0.05"
coef_between$tuk_signif[coef_between$p.adj <= 0.05] <- "p-value <= 0.05"


#########Actual summary plot

anova_summaries <- data.frame("subset" = unique(coef_between$comb), "p_val_signif" = NA, "p_val_Insignif" = NA, "significance_norm" = NA )
subset_split <- str_split(anova_summaries$subset,  "-", simplify=TRUE)
anova_summaries$X1 <- subset_split[,1]
anova_summaries$X2 <- subset_split[,2]

for (i in unique(coef_between$comb)) {

  df <- coef_between[coef_between$comb == i, ]  

  anova_summaries$p_val_Insignif[anova_summaries$subset ==i] <- nrow(df[df$tuk_signif == "p-value > 0.05",])
  anova_summaries$p_val_signif[anova_summaries$subset ==i] <- nrow(df[df$tuk_signif == "p-value <= 0.05",])
  anova_summaries$significance_norm[anova_summaries$subset ==i] <- nrow(df[df$tuk_signif == "p-value <= 0.05",])/nrow(df)
}
 

symmetry <- anova_summaries

ggplot(rbind(anova_summaries, symmetry), mapping=aes(X1, X2, fill = as.numeric(significance_norm )) ) + 
  geom_tile( alpha = 0.8)+ 
   geom_text( aes(label=round(significance_norm, digits = 3)), size =3 )+ 
  theme(axis.text.x = element_text(angle =90))+
    scale_fill_gradientn(colors = hcl.colors(20, "Cyan-Magenta"))+
  theme_classic()+ theme(axis.text.x = element_text(angle =90))+ggtitle(paste("p-values for Anova of quantile regression coefficients"))

testpval_between_coef_summary <- data.frame("metrics" = "Anova on coefficients(slopes), p-value <=0.05/total number of slides+species compared",
                       "suns_for_clust_1" = anova_summaries$X1,
                        "suns_for_clust_2" = anova_summaries$X2,
                         "metrics_value" = anova_summaries$significance_norm )
 
datatable(testpval_between_coef_summary)
 

```



## making a dataframe for summarising how regressions differ between each other

```{r, echo=FALSE, message=FALSE, warning=FALSE}

difference_summary <- na.omit(rbind(testpval_between_MEDIAN_summary, testpval_between_KS_summary, testpval_between_coef_summary))

#difference_summary_R <- reshape(difference_summary, idvar = c("metrics", "suns_for_clust_1"),timevar = "suns_for_clust_2",  direction = "wide")
fwrite(difference_summary, "difference_summary.csv")

datatable(difference_summary)

```













# Testing if the difference between predicted medians for n and 1n ~2.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ploidy_ratios_MEAN_summary <- data.frame()

ploidy_ratios <- data_frame()

for(sp in unique(toplot$sp)){
  
  df <- toplot[toplot$sp == sp,]
  # actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  
  df1 <- df[df$if_test_sp=="TEST_bootstrap",]
  #df1$for_medians <- paste(df1$suns_for_clust, df1$slidedate)
  df1 <- df1[ which(is.na(df1$bp_exp_ploidy_corr) == FALSE), ]
  
  for (setting in unique(df1$suns_for_clust)) {
    df3 <- df1[df1$suns_for_clust == setting,]
    vect_med2nvs1n <- c()
  for (i in unique(df3$slidedate)) {
    
    #if (nrow(df3[df3$slidedate==i,]) == 0 ) next 
    df2 <- df3[df3$slidedate==i, ]
    
    n1_med <- median(df2$bp_exp_ploidy_corr[df2$ploidy_clust== "1n"])
    n2_med <- median(df2$bp_exp_ploidy_corr[df2$ploidy_clust== "2n"])
    
    if (length(n1_med) ==0 | length(n2_med) == 0) next
    
    med2nvs1n <- n2_med/n1_med
    
    vect_med2nvs1n <- c(vect_med2nvs1n, med2nvs1n)
    
    tmp1 <-   data.frame("suns_for_clust" = setting,
                         "slidedate" = i,
                        "sp" = sp,
                         "metrics_value" = med2nvs1n)
    
    ploidy_ratios <- rbind(ploidy_ratios, tmp1)
  }
   tmp <-   data.frame("metrics" = "mean(median of 1n cells/ median of 2n cells)",
                       "suns_for_clust" = setting,
                        "sp" = sp,
                         "metrics_value" = mean(vect_med2nvs1n))
                       
   ploidy_ratios_MEAN_summary <- rbind(MEDIAN_summary, tmp)
  } 
}


```



```{r, echo=FALSE, message=FALSE, warning=FALSE}

ploidy_ratios[ which(ploidy_ratios$metrics_value < -10 | ploidy_ratios$metrics_value > 10) , ]

gg <- ggplot(ploidy_ratios, aes( slidedate, metrics_value, colour=sp) )+geom_point()+
  theme_bw()+
  facet_wrap(~suns_for_clust, scales="free")+theme(axis.text.x = element_blank())+
  geom_hline(yintercept = 2, color = "red")+
  ggtitle("Median of 2n/Median of 1n cells for each slide, unfiltered")

print(gg)

# ploidy_ratios$exp_series_date <- str_split(ploidy_ratios$slidedate, " ", simplify = T )[ ,2]


```


```{r, echo=FALSE, message=FALSE, warning=FALSE}

ploidy_ratios_filt <- ploidy_ratios[ -which(ploidy_ratios$metrics_value < -0 | ploidy_ratios$metrics_value > 3) , ]

gg <- ggplot(ploidy_ratios_filt, aes( slidedate, metrics_value, colour=sp) )+geom_point()+
  theme_bw()+
  facet_wrap(~suns_for_clust, scales="free")+theme(axis.text.x = element_blank())+
  geom_hline(yintercept = 2, color = "red")+
  ggtitle("Median of 2n/Median of 1n cells for each slide, filtered")

print(gg)

# ploidy_ratios$exp_series_date <- str_split(ploidy_ratios$slidedate, " ", simplify = T )[ ,2]


```

### same as above, but for ratio of medians for sets of cells in exp_series_date and all spesies


```{r, echo=FALSE, message=FALSE, warning=FALSE}

ploidy_ratios_exp_series <- data_frame()

for(sp in unique(toplot$sp)){
  
  df <- toplot[toplot$sp == sp,]
  # actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  
  df1 <- df[df$if_test_sp=="TEST_bootstrap",]
  #df1$for_medians <- paste(df1$suns_for_clust, df1$slidedate)
  df1 <- df1[ which(is.na(df1$bp_exp_ploidy_corr) == FALSE), ]
  
  for (setting in unique(df1$suns_for_clust)) {
    df3 <- df1[df1$suns_for_clust == setting,]
    vect_med2nvs1n <- c()
  for (i in unique(df3$exp_series_date)) {
    
    #if (nrow(df3[df3$slidedate==i,]) == 0 ) next 
    df2 <- df3[df3$exp_series_date==i, ]
    
    n1_med <- median(df2$bp_exp_ploidy_corr[df2$ploidy_clust== "1n"])
    n2_med <- median(df2$bp_exp_ploidy_corr[df2$ploidy_clust== "2n"])
    
    if (length(n1_med) ==0 | length(n2_med) == 0) next
    
    med2nvs1n <- n2_med/n1_med
    
    tmp1 <-   data.frame("suns_for_clust" = setting,
                         "exp_series_date" = i,
                        "sp" = sp,
                         "metrics_value" = med2nvs1n)
    
    ploidy_ratios_exp_series <- rbind(ploidy_ratios_exp_series, tmp1)
  }

  } 
}

ploidy_ratios_exp_series[ which(ploidy_ratios_exp_series$metrics_value < -10 | ploidy_ratios_exp_series$metrics_value > 10) , ]

gg <- ggplot(ploidy_ratios_exp_series, aes( as.factor(exp_series_date) , metrics_value, colour=sp) )+geom_point()+
  theme_bw()+
  facet_wrap(~suns_for_clust, scales="free")+theme(axis.text.x = element_blank())+
  geom_hline(yintercept = 2, color = "red")+
  ggtitle("Median of 2n/Median of 1n cells for each experimental series date")

print(gg)



ploidy_ratios_est_sp <- data_frame()

for(sp in unique(toplot$sp)){
  
  df <- toplot[toplot$sp == sp,]
  # actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  
  df1 <- df[df$if_test_sp=="TEST_bootstrap",]
  #df1$for_medians <- paste(df1$suns_for_clust, df1$slidedate)
  df1 <- df1[ which(is.na(df1$bp_exp_ploidy_corr) == FALSE), ]
  
  for (setting in unique(df1$suns_for_clust)) {
    df3 <- df1[df1$suns_for_clust == setting,]
    vect_med2nvs1n <- c()

    #if (nrow(df3[df3$slidedate==i,]) == 0 ) next 
    df2 <- df3#[df3$exp_series_date==i, ]
    
    n1_med <- median(df2$bp_exp_ploidy_corr[df2$ploidy_clust== "1n"])
    n2_med <- median(df2$bp_exp_ploidy_corr[df2$ploidy_clust== "2n"])
    
    if (length(n1_med) ==0 | length(n2_med) == 0) next
    
    med2nvs1n <- n2_med/n1_med
    
    tmp1 <-   data.frame("suns_for_clust" = setting,
                         "exp_series_date" = i,
                        "sp" = sp,
                         "metrics_value" = med2nvs1n)
    
    ploidy_ratios_est_sp <- rbind(ploidy_ratios_est_sp, tmp1)
 
  } 
}


gg <- ggplot(ploidy_ratios_est_sp, aes( sp , metrics_value, colour=sp) )+geom_point()+
  theme_bw()+
  facet_wrap(~suns_for_clust, scales="free")+theme(axis.text.x = element_blank())+
  geom_hline(yintercept = 2, color = "red")+
  ggtitle("Median of 2n/Median of 1n cells for each species (all series, all slides)")

print(gg)


```














# Accuracy of predictions depending on the number of datapoints( = measured cells) used for estimating regression fit 
!!! Warning !!!(the size of the dataset for clusterisation is not considered here)

## For the test species, medians

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#toplot <- toplot[toplot$pvalue <= 0.05,]

test_cellcount_medians <- data.frame()

for(sp in unique(toplot$sp)){
  
  df <- toplot[toplot$sp == sp,]
  actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  
  df1 <- df[df$if_test_sp=="TEST_bootstrap",]
  #df1$for_medians <- paste(df1$suns_for_clust, df1$slidedate)
  df1 <- df1[ which(is.na(df1$bp_exp_ploidy_corr) == FALSE), ]
  
  for (setting in unique(df1$suns_for_clust)) {
  for (ploidy in unique(df1$ploidy_clust)) {
    df2 <- df1[df1$suns_for_clust == setting,]
    df2 <- df2[df2$ploidy_clust == ploidy,]
    
  for (i in unique(df2$slidedate)) {
    med <- median(df2$bp_exp_ploidy_corr[df2$slidedate==i])
  
    if (ploidy == "1n") { 
      diff <- min(actual_pos) - med 
      diff_norm <-(min(actual_pos) - med)/min(actual_pos) } else 
    if (ploidy == "2n") { 
      diff <- med - max(actual_pos) 
      diff_norm <- (med - max(actual_pos))/max(actual_pos) } else 
    if (ploidy == "Undefined") {
    if (med < min(actual_pos)) {
      diff <-  min(actual_pos) - med 
      diff_norm <-  (min(actual_pos) - med)/min(actual_pos) } 
    else if(med > max(actual_pos)){
      diff <- med - max(actual_pos)  
      diff_norm <- (med - max(actual_pos))/max(actual_pos) }
    else if(med > min(actual_pos) & med < max(actual_pos)) {
      diff <- mean( abs(min(actual_pos) - med), abs(med - max(actual_pos)) )
      diff_norm <- mean( abs(min(actual_pos) - med), abs(med - max(actual_pos)) )/mean((min(actual_pos)),(max(actual_pos)) ) } 
          }
    
    if ( str_split(unique(df2$suns_for_clust), " ", simplify = T )[3] == "slide" ) {
      test_cellcount <- length(unique(df2$cell_coord[df2$slidedate==i]))  }
    else if ( str_split(unique(df2$suns_for_clust), " ", simplify = T )[3] == "series" ) {
      series_date <- str_split(i, " ", simplify = T )[2] 
      test_cellcount <- length(unique(df2$cell_coord[df2$exp_series_date==series_date])) }
    #debug
    #print(i)
    #print(setting)
    #print(sp)
    #print(test_cellcount)
   PERSLIDEtest_cellcount <- length(unique(df2$cell_coord[df2$slidedate==i]))
   mean_CI <- as.numeric(df2$higher[df2$slidedate==i]) - as.numeric(df2$lower[df2$slidedate==i])
   mean_CI_norm <- as.numeric(df2$higher[df2$slidedate==i])/as.numeric(df2$bp_exp_ploidy_corr[df2$slidedate==i]) - as.numeric(df2$lower[df2$slidedate==i])/as.numeric(df2$bp_exp_ploidy_corr[df2$slidedate==i])
   
   tmp <-   data.frame("metrics" = "Mean(|Predicted medians - expected_bp|)/bp_exp",
                      "suns_for_clust" = setting,
                      "sp" = sp,
                      "slidedate" = i, 
                      "test_cellcount" = test_cellcount,
                      "PERSLIDEtest_cellcount" = PERSLIDEtest_cellcount,
                      "pred_med_vs_act" = diff, 
                      "pred_med_vs_act_norm" = diff_norm,
                      "mean_CI" = mean(mean_CI),
                      "mean_CI_norm" = mean(mean_CI_norm),
                      "bootstrap_setting" = unique(df2$bootstrap_setting) #paste(sp, setting, i)
                      )
   test_cellcount_medians <- rbind(test_cellcount_medians, tmp) 
  }
  } }
}



```


### corresponding plots.
Number of cells PER SINGLE SLIDE is plotted - as, so far, we estimate separate median for each slide measured for kinetoplastid. Can be changed to the total number of species within a given slide OR series - depending how regression was inferred - for this, change PERSLIDEtest_cellcount ==> test_cellcount

```{r, echo=FALSE, message=FALSE, warning=FALSE}
gg <- ggplot(na.omit(test_cellcount_medians), aes( mean_CI_norm, PERSLIDEtest_cellcount, colour=sp) )+geom_point()+
  theme_bw()+scale_x_continuous(trans = "log2")+
  facet_wrap(~suns_for_clust, scales="free")+theme(axis.text.x = element_text(angle =90))+
  ggtitle("CI (upper-lower)/bp_exp_ploidy_corr VS test species cells per 1 slide(!)")

print(gg)


gg <- ggplot(na.omit(test_cellcount_medians), aes( abs(pred_med_vs_act_norm), PERSLIDEtest_cellcount, colour=sp) )+
  geom_point()+
 #geom_line(aes(group = sp, colour=sp))+
  theme_bw()+#scale_x_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
  facet_wrap(~suns_for_clust, scales="free")+
  ggtitle("(Median pred - median exp)/median_exp VS test species cells - per 1 slide(!)")

print(gg)

```

### For the particular regression+clustering method (which was selected for further kinetoplastid estimates)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
gg <- ggplot(na.omit(test_cellcount_medians[test_cellcount_medians$suns_for_clust=="est_sp slidedate series",]), aes( mean_CI_norm, PERSLIDEtest_cellcount) )+
  geom_point(aes(colour=sp))+
  #geom_smooth(se = FALSE, method = lm)+
  theme_bw()+scale_x_continuous(trans = "log2")+
  facet_wrap(~suns_for_clust, scales="free")+
  theme(axis.text.x = element_text(angle =90))+
  ggtitle("CI (upper-lower)/bp_exp_ploidy_corr VS test species cells per 1 slide(!)")

print(gg)


gg <- ggplot(na.omit(test_cellcount_medians[test_cellcount_medians$suns_for_clust=="est_sp slidedate series",]), aes( abs(pred_med_vs_act_norm), PERSLIDEtest_cellcount) )+
  geom_point(aes(colour=sp))+
  #geom_smooth(se = FALSE, method = lm, aes(colour=sp))+
  theme_bw()+#scale_x_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
  facet_wrap(~suns_for_clust, scales="free")+
  ggtitle("(Median pred - median exp)/median_exp VS test species cells - per 1 slide(!)")

print(gg)

```



## For the control species count, how it affects predicted medians of test species

```{r, echo=FALSE, message=FALSE, warning=FALSE}

control_cellcount_medians <- data.frame()

for(BS in unique(toplot$bootstrap_setting)){
  
  df <- toplot[toplot$bootstrap_setting == BS,]
  df <- df[df$if_test_sp == "CONTROL_species",]
  #df1$for_medians <- paste(df1$suns_for_clust, df1$slidedate)
  df1 <- df[ which(is.na(df$bp_exp_ploidy_corr) == FALSE), ]
  
  for (setting in unique(df1$suns_for_clust)) {
    for (sp in unique(df1$sp)) {
      
    df2 <- df1[df1$suns_for_clust == setting,]
    df2 <- df2[df2$sp == sp,]
      
     for (i in unique(df2$slidedate)) {
       
       
    if ( str_split(unique(df2$suns_for_clust), " ", simplify = T )[3] == "slide" ) {
      control_cellcount <- length(unique(df2$cell_coord[df2$slidedate==i]))  }
    else if ( str_split(unique(df2$suns_for_clust), " ", simplify = T )[3] == "series" ) {
      series_date <- str_split(i, " ", simplify = T )[2] 
      control_cellcount <- length(unique(df2$cell_coord[df2$exp_series_date==series_date])) }
     }
    #debug
  #  print(i)
  #  print(setting)
  #  print(sp)
  #  print(control_cellcount)

    tmp <-   data.frame("suns_for_clust" = setting,
                      "control_sp" = sp,
                      "slidedate" = i, 
                      "control_cellcount" = control_cellcount, 
                      "bootstrap_setting"= BS #paste(sp, setting, i)
                        )

   control_cellcount_medians <- rbind(control_cellcount_medians, tmp) 

} } 
}



```


```{r, echo=FALSE, message=FALSE, warning=FALSE}

test_control_cellcount <- left_join(test_cellcount_medians, control_cellcount_medians, by = "bootstrap_setting")

#debug
#unique(test_cellcount_medians$bootstrap_setting) %in% unique(control_cellcount_medians$bootstrap_setting)
#unique(control_cellcount_medians$control_cellcount)

old_names <- names(test_control_cellcount)
new_names <- sub("\\..*", "", old_names)
will_be_duplicated <- duplicated(new_names) | duplicated(new_names, fromLast = TRUE)
final_names <- ifelse(will_be_duplicated, old_names, new_names)
names(test_control_cellcount) <- final_names


test_control_cellcount <- test_control_cellcount[, colSums(!is.na(test_control_cellcount)) > 0]

colnames(test_control_cellcount)

test_control_cellcount <- test_control_cellcount[!duplicated(test_control_cellcount),]

test_control_cellcount <- test_control_cellcount[test_control_cellcount$suns_for_clust.x == test_control_cellcount$suns_for_clust.y,]

```


### corresponding plots.


```{r, echo=FALSE, message=FALSE, warning=FALSE}
gg <- ggplot(na.omit(test_control_cellcount), aes( mean_CI_norm, control_cellcount, colour=suns_for_clust.x) )+
  geom_point(aes(size = test_cellcount), alpha = 0.2)+
  #geom_line()+
  theme_bw()+scale_x_continuous(trans = "log2")+
  facet_wrap(~control_sp+sp, scales="free")+theme(axis.text.x = element_text(angle =90))+
  ggtitle("CI (upper-lower)/bp_exp_ploidy_corr VS control species cells per regression")

#print(gg)

ggsave("cellcount_CI_values_control_species.pdf", gg, width = 15, height =15)

gg <- ggplot(na.omit(test_control_cellcount), aes( pred_med_vs_act_norm, control_cellcount, colour=suns_for_clust.x) )+
  geom_point(aes(size = test_cellcount), alpha = 0.2)+
  #geom_line()+
  theme_bw()+scale_x_continuous(trans = "log2")+
  facet_wrap(~control_sp+sp, scales="free")+theme(axis.text.x = element_text(angle =90))+
  ggtitle("(Median pred - median exp)/median_exp VS control species cells per regression")

#print(gg)

ggsave("cellcount_median_values_control_species.pdf", gg, width = 15, height =15)

```


