---
title: "Untitled"
author: "Zavadska Daryna"
date: "2025-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r, echo=FALSE , message=FALSE}
##upload packages
library(stringr)
library(openxlsx)
library(data.table)
library(magrittr)
#library(GGally)
#library(propr)
library(data.table)
library(ggplot2)
library(reshape2)
library(tidyverse)
library(dplyr)
library(plyr)
library(microshades)
library(scales)
library(ggbreak) 
library(patchwork)
library(ggpubr)
library(gplots)
library(PerformanceAnalytics)

library(patchwork)
library(microshades)
library(ggtext)
library(gridExtra)
library(ggpmisc)

```

```{r}
colors <-c(microshades_palette("micro_blue", lightest = FALSE), 
           microshades_palette("micro_purple", lightest = FALSE), microshades_palette("micro_cvd_purple", lightest = FALSE), 
           microshades_palette("micro_green", lightest = FALSE), 
           microshades_palette("micro_cvd_green", lightest = FALSE),
           microshades_palette("micro_orange", lightest = FALSE), 
           microshades_palette("micro_brown", lightest = FALSE), 
           microshades_palette("micro_cvd_turquoise", lightest = FALSE),
           microshades_palette("micro_cvd_gray", lightest = FALSE))

```





```{r}
#source tables here https://docs.google.com/spreadsheets/d/1t1BwSQRRvVuYbAhggKFb4Judfg_OvTrHUxUnBVghA5M/edit?gid=1209785171#gid=1209785171
list <- list.files(path = "./", pattern = "BUSCO_out.tsv")

```



```{r}

busco_db_sp <- fread("BUSCO_databases.tsv")




big_df <- data.frame()

for (i in list) {
  
  name <- str_replace(i, "_BUSCO_out.tsv","")
  
  BUSCO_eugl <- fread(i)
#check taxonomic affiliation, cause there was a typo in table 
  BUSCO_eugl$group[BUSCO_eugl$species == "Apiculatamorpha_spiralis"] <- "Prokinetoplastids"
  
  
BUSCO_eugl_m <- melt(BUSCO_eugl)

if (grepl("NUCL", i)==FALSE) { BUSCO_eugl_m <- BUSCO_eugl_m[-grep("Willaertia.*NON", BUSCO_eugl_m$file),] }

#subsetting euglenozoan BUSCO to only euglenozoa
if (grepl("EUGL", name)) {
  BUSCO_eugl_m <- BUSCO_eugl_m[BUSCO_eugl_m$group != "Other",]
  db_list <- busco_db_sp$species[busco_db_sp$busco_db=="EUGL"]  }else{
    db_list <- busco_db_sp$species[busco_db_sp$busco_db!="EUGL"]
  }


sp_to_label <- paste0( BUSCO_eugl_m$species[BUSCO_eugl_m$species %in% db_list ] , ".", BUSCO_eugl_m$group[BUSCO_eugl_m$species %in% db_list ] )

BUSCO_eugl_m$assembly <- NA
BUSCO_eugl_m$assembly[grep("denovo", BUSCO_eugl_m$file)] <- "de novo"
BUSCO_eugl_m$assembly[grep("REF", BUSCO_eugl_m$file)] <- "reference"
BUSCO_eugl_m$assembly[grep("NON", BUSCO_eugl_m$file)] <- "reference"



n <- unique(BUSCO_eugl_m$value[which(BUSCO_eugl_m$variable=="n")])

ggplot(BUSCO_eugl_m[!BUSCO_eugl_m$variable%in% c("C","n"),], aes(value/n, interaction(species,group), fill =  as.factor(variable)) )+
  geom_bar(stat = "identity", width = 0.8)+
  scale_y_discrete(labels = ~ if_else( .x %in% sp_to_label, paste0("<span style='color: red'><b>", .x, "</b></span>"),  .x  )) +
  #geom_text(aes(label = value), hjust=-0.5, size = 2) +
  theme_bw()+ theme(axis.text.y = element_markdown())+
  facet_wrap(~assembly,ncol =2) +
  ggtitle(name)


ggsave(paste0(name,".png"), width = 10, height = 10)

  
BUSCO_eugl_m$setting <- name
big_df <- rbind(BUSCO_eugl_m, big_df)
  
}


db_list_eugl <- busco_db_sp$species[busco_db_sp$busco_db=="EUGL"] 
db_list_euk <- busco_db_sp$species[busco_db_sp$busco_db!="EUGL"]

```


## Plots for the preprint

```{r}

big_df_plot <- big_df[!big_df$variable%in% c("C","n"),]

big_df_plot$redundancy <- str_split(big_df_plot$setting, "_", simplify = T)[,1]
big_df_plot$busco_db <- str_split(big_df_plot$setting, "_", simplify = T)[,2]

big_df_plot <- big_df_plot[big_df_plot$assembly == "reference" & big_df_plot$redundancy == "NR90",]


 db_list_euk <- busco_db_sp$species[busco_db_sp$busco_db=="EUK"] 
 db_list_eugl <- busco_db_sp$species[busco_db_sp$busco_db=="EUGL"] 

 
sp_to_label_eugl <- paste0( big_df_plot$species[big_df_plot$species %in% db_list_eugl ] , ".", big_df_plot$group[big_df_plot$species %in% db_list_eugl ] )
sp_to_label_euk <- paste0( big_df_plot$species[big_df_plot$species %in% db_list_euk ] , ".", big_df_plot$group[big_df_plot$species %in% db_list_euk ] )



p1 <- ggplot(big_df_plot[big_df_plot$busco_db=="EUGL",], aes(value/130, interaction(species,group), fill =  as.factor(variable)) )+
  geom_bar(stat = "identity", width = 0.8)+
  scale_y_discrete(labels = ~ if_else( .x %in% sp_to_label_eugl, paste0("<span style='color: red'><b>", .x, "</b></span>"),  .x  )) +
  #geom_text(aes(label = value), hjust=-0.5, size = 2) +
  theme_bw()+ theme(axis.text.y = element_markdown())+xlab("BUSCO completeness")+ylab("species") + labs(fill = "BUSCO category")+
   theme(legend.position = "bottom")+
   scale_fill_brewer(palette = "Set3")+
  ggtitle("Euglenozoa")


p2 <- ggplot(big_df_plot[big_df_plot$busco_db=="EUK",], aes(value/255, interaction(species,group), fill =  as.factor(variable)) )+
  geom_bar(stat = "identity", width = 0.8)+
  scale_y_discrete(labels = ~ if_else( .x %in% sp_to_label_euk, paste0("<span style='color: red'><b>", .x, "</b></span>"),  .x  )) +
  #geom_text(aes(label = value), hjust=-0.5, size = 2) +
  theme_bw()+ theme(axis.text.y = element_markdown() )+xlab("BUSCO completeness")+ylab("species") + labs(fill = "BUSCO category")+
   theme(legend.position = "bottom")+
  scale_fill_brewer(palette = "Set3")+
  ggtitle("Eukaryota")


gg <- (p1 + theme(legend.position="none")) | p2  

grid.arrange(p1,p2, nrow = 1)

ggsave("BAD_BUSCO.png", gg, width = 15, height = 10)



```




```{r}

big_df$redundancy <- str_split(big_df$setting, "_", simplify = T)[,1]
big_df$busco_db <- str_split(big_df$setting, "_", simplify = T)[,2]

big_df$value_percent <- NA
big_df$value_percent[big_df$busco_db=="EUGL"] <- big_df$value[big_df$busco_db=="EUGL"]/unique(big_df$value[big_df$busco_db=="EUGL"& big_df$variable=="n"])
big_df$value_percent[big_df$busco_db=="EUK"] <- big_df$value[big_df$busco_db=="EUK"]/unique(big_df$value[big_df$busco_db=="EUK"& big_df$variable=="n"])

p1 <- ggplot(big_df[big_df$variable=="D",], aes(redundancy, value_percent) )+
  geom_point(aes(colour=assembly), alpha=0.5)+
  geom_line(aes(group=interaction(species,assembly), colour=assembly), alpha=0.5)+
  scale_x_discrete(limits=c("NR90","ORIGINAL","NUCLEOTIDE"))+
  theme_bw()+ facet_wrap(~busco_db,ncol =2)+
    theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position = "none")+
  ggtitle("Duplicated BUSCO")

p2 <- ggplot(big_df[big_df$variable=="C",], aes(redundancy, value_percent) )+
  geom_point(aes(colour=assembly), alpha=0.5)+
  geom_line(aes(group=interaction(species,assembly), colour=assembly), alpha=0.5)+
  scale_x_discrete(limits=c("NR90","ORIGINAL","NUCLEOTIDE"))+
  theme_bw()+ facet_wrap(~busco_db,ncol =2)+
    theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position = "bottom")+
  ggtitle("Complete BUSCO")


gg <- (p1 + theme(legend.position="none")) | p2  

grid.arrange(p1,p2, nrow = 1)

ggsave("NR_BUSCO.png", gg)



```



## Comparing de novo vs reference assemblies BUSCO results

```{r}

#list of species for which de novo assembly was run
denov <- big_df[ big_df$species %in% unique(big_df$species[big_df$assembly=="de novo"]), ]

plots <- list()


for(i in unique(denov$setting)) {
  denov_sub <- denov[denov$setting == i, ]
 # n <- unique(denov_sub$value[denov_sub$variable == "n"])

  p <- ggplot(denov_sub[!denov_sub$variable %in% c("C","n"),],
              aes(x = value, y = interaction(species, group), fill = as.factor(variable))) +
        geom_bar(stat = "identity", width = 0.8) +
        theme_bw() + facet_wrap(~assembly, ncol = 2) +
        ggtitle(i)

  plots[[i]] <- p
  

}

combined <- wrap_plots(plots)

ggsave("all_denov_vs_ref_BUSCO.png", combined, width = 20, height = 10)




df_big <- data.frame()

for(i in unique(denov$setting)) {
  denov_sub <- denov[denov$setting == i, ] 
  
  n <- unique(denov_sub$value[denov_sub$variable == "n"]) 
  
  denov_sub <- denov_sub[denov_sub$variable=="C",] 
  
 df <- denov_sub %>% group_by(species) %>% 
    summarize(ref_value = value[assembly == "reference"], 
              denov_value = value[assembly == "de novo"] ) %>%  
    mutate(diff = (denov_value - ref_value)/n) 

   df$species <- unique(denov_sub$species)
   df$group <- denov_sub$group[seq(1, length(denov_sub$group), 2)]
  df$setting <- i
  
  df_big <- rbind(df_big, df)
  
}



ggplot(df_big, aes(diff*-1, interaction(species,group) ) )+
  geom_bar(stat="identity")+
  #geom_text(aes(label = value), hjust=-0.5, size = 2) +
  theme_bw()+ facet_wrap(~setting,ncol =2)

fwrite(df_big, "df_big_BUSCO_diff.tsv")

```


## Other assembly features (protein length)

```{r}

len_NR <- fread("NR90_assembly_lengths.tsv")
len_ORIGINAL <- fread("ORIGINAL_assembly_lengths.tsv")

len_NR$redundancy <- "NR90"
len_ORIGINAL$redundancy <- "ORIGINAL"
len <- rbind(len_NR, len_ORIGINAL)

len$species <- NA
for(sp in unique(big_df$species)) {
  len$species[grep(sp, len$file) ] <- sp }

len$species[grep("Trypanosoma_cruzi_CLBrenerEsmeraldo-like", len$file) ] <- "Trypanosoma_cruzi_CLBrenerEsmeraldo-like" 


len$assembly <- NA
len$assembly[grep("denovo", len$file)] <- "de novo"
len$assembly[grep("REF", len$file)] <- "reference"
len$assembly[grep("NON", len$file)] <- "reference"



denov_len <- denov[!(denov$busco_db == "EUGL"), ]
denov_len <- merge(denov_len, len, by = c("species", "redundancy", "assembly"))


df_big_len <- data.frame()

for(i in unique(denov_len$setting)) {
  denov_sub <- denov_len[denov_len$setting == i, ] 
  denov_sub <- denov_sub[denov_sub$variable=="C",] 
  
  for(sp in unique(denov_sub$species)) {
    denov_subsub <- denov_sub[denov_sub$species == sp,]
    diff <- denov_subsub$sum_len[denov_subsub$assembly == "reference"] - denov_subsub$sum_len[denov_subsub$assembly == "de novo"]
  print(diff)
  
   df <- data.frame(species = sp,
                    group = denov_subsub$group[1],
                    assembly = denov_subsub$assembly[1],
                    redundancy = denov_subsub$redundancy[1],
                    setting = i,
                    ref_vs_denov = diff, 
                    ref = denov_subsub$sum_len[denov_subsub$assembly == "reference"],
                    denov = denov_subsub$sum_len[denov_subsub$assembly == "de novo"])
   
   df_big_len <- rbind(df_big_len, df)
  }
}



ggplot(df_big_len, aes(ref_vs_denov*3/1000000, interaction(species,group) ) )+
  geom_bar(stat="identity")+
  #geom_text(aes(label = value), hjust=-0.5, size = 2) +
  theme_bw()+ facet_wrap(~ interaction(assembly, redundancy),ncol =2)+
  xlab("Reference - de novo assembled protein-coding part, Mbp")+
  ylab("")


ggplot(df_big_len, aes((ref_vs_denov/ref), interaction(species,group) ) )+
  geom_bar(stat="identity")+
  #geom_text(aes(label = value), hjust=-0.5, size = 2) +
  theme_bw()+ facet_wrap(~ interaction(assembly, redundancy),ncol =2)+
  xlab("(Reference - de novo assembled protein-coding part)/Reference protein-coding")+
  ylab("")


fwrite(df_big_len, "df_big_len.tsv")

```



## Plots for the preprint

```{r}


denov <- big_df[ big_df$species %in% unique(big_df$species[big_df$assembly=="de novo"]), ]
denov_sub <- denov[denov$redundancy=="NR90", ]
 # n <- unique(denov_sub$value[denov_sub$variable == "n"])

d1 <- ggplot(denov_sub[!denov_sub$variable %in% c("C","n"),],
        aes(x = value, y = interaction(species, group), fill = as.factor(variable))) +
        geom_bar(stat = "identity", width = 0.8) +
        theme_bw() + facet_wrap(~interaction(assembly, busco_db), scales = "free_x") +
   xlab("BUSCO completeness")+ylab("species") + labs(fill = "BUSCO category")+
      scale_fill_brewer(palette = "Set3")+
        ggtitle("A")


```


```{r}


df_big_len$redundancy <- str_split(df_big_len$setting, "_", simplify = T)[,1]
df_big_len$busco_db <- str_split(df_big_len$setting, "_", simplify = T)[,2]

dupl <- df_big_len
dupl$busco_db <- "EUGL"
dupl$setting <- paste0(dupl$redundancy, "_EUGL")

rbind(df_big_len, dupl) -> df_big_len_2

df_big_merged <- merge(df_big, df_big_len_2)

df_big_merged$redundancy <- str_split(df_big_merged$setting, "_", simplify = T)[,1]
df_big_merged$busco_db <- str_split(df_big_merged$setting, "_", simplify = T)[,2]



d2 <- ggplot(df_big_merged[df_big_merged$redundancy=="NR90",], aes(diff, ref_vs_denov/ref) )+
  geom_smooth(method = "lm")+
    stat_poly_eq( formula = y ~ x,  aes(label = paste( ..rr.label.., ..p.value.label.., sep = "~~~")),parse = TRUE, label.x = 0.05, label.y = 0.95 ) +
  geom_point()+
  geom_text(aes(label=species), size =2.5, hjust=1.1)+
  theme_bw()+ facet_wrap(~busco_db,ncol =2)+
    theme(legend.position = "bottom")+
  xlab("BUSCO de novo - BUSCO reference ")+ylab("(Length reference - Length de novo)/Length reference") +
  ggtitle("C")


df_big_merged$n <- NA
df_big_merged$n[df_big_merged$busco_db=="EUGL"] <- 130
df_big_merged$n[df_big_merged$busco_db=="EUK"] <- 255

d3 <- ggplot(df_big_merged[df_big_merged$redundancy=="NR90",], aes(denov_value/n, ref_vs_denov/ref) )+
  geom_point()+
 geom_text(aes(label=species), size =2.5, hjust=1.1)+
  theme_bw()+ facet_wrap(~busco_db,ncol =2)+
  geom_smooth(method = "lm")+
  stat_poly_eq( formula = y ~ x,  aes(label = paste( ..rr.label.., ..p.value.label.., sep = "~~~")),parse = TRUE, label.x = 0.05, label.y = 0.95 ) +
  xlab("BUSCO de novo ")+ylab("(Length reference - Length de novo)/Length reference") +
  ggtitle("D")




gg <- grid.arrange(d1,d2,d3,layout_matrix = rbind(c(1, 1), c(2, 3))  )


ggsave("denov_BUSCO.png", gg, width = 13, height = 13)


```





## Lenghts distributions

distributions of protein lengths in de novo vs reference assemblies

```{r}

pre_gen_len_or <- fread("BUSCO_ORIGINAL_length.tsv")
pre_gen_len_NR <- fread("BUSCO_NR90_length.tsv")

pre_gen_len_or$redundancy <- "ORIGINAL"
pre_gen_len_NR$redundancy <- "NR90"


pre_gen_len <- rbind(pre_gen_len_or, pre_gen_len_NR)


pre_gen_len$assembly <- NA
pre_gen_len$assembly[grep("denovo", pre_gen_len$V1)] <- "de novo"
pre_gen_len$assembly[grep("REF", pre_gen_len$V1)] <- "reference"
pre_gen_len$assembly[grep("NON", pre_gen_len$V1)] <- "reference"

pre_gen_len$species <- NA
for(sp in unique(big_df$species)) {
  pre_gen_len$species[grep(sp, pre_gen_len$V1) ] <- sp 
  }

pre_gen_len$species[grep("Trypanosoma_cruzi_CLBrenerEsmeraldo-like", pre_gen_len$V1) ] <- "Trypanosoma_cruzi_CLBrenerEsmeraldo-like" 


```

```{r}

gg <- ggplot(pre_gen_len , aes(V3 ) )+
  geom_histogram(aes(fill = interaction(assembly,redundancy)), alpha=0.3, position = "identity")+scale_x_log10()+
#  scale_color_manual(values = colors[c(38,48,23,4)] ) + scale_fill_manual(values = colors[c(38,48,23,4)] ) +
  theme_bw()+ facet_wrap(~species,scales = "free_y")

ggsave("all_denov_vs_ref_Length_PER_GENE.png", gg, width = 25, height = 10)


ggplot(pre_gen_len[pre_gen_len$species =="Bodo_saltans",] , aes(V3 ) )+
  geom_histogram(aes(fill = redundancy), alpha=0.25, position = "identity")+scale_x_log10()+
  #geom_text(aes(label = value), hjust=-0.5, size = 2) +
  theme_bw()+ facet_wrap(~ assembly,scales = "free_y")

```






```{r}
## sanity check, should return TRUE
sum(pre_gen_len$V3[pre_gen_len$species =="Bodo_saltans" & pre_gen_len$assembly == "reference"  & pre_gen_len$redundancy == "NR90"]) == df_big_len$ref[df_big_len$species =="Bodo_saltans" & df_big_len$redundancy == "NR90"]


sum(pre_gen_len$V3[pre_gen_len$species =="Bodo_saltans" & pre_gen_len$assembly == "reference"  & pre_gen_len$redundancy == "NR90"])
sum(pre_gen_len$V3[pre_gen_len$species =="Bodo_saltans" & pre_gen_len$assembly == "reference"  & pre_gen_len$redundancy == "ORIGINAL"])

sum(pre_gen_len$V3[pre_gen_len$species =="Bodo_saltans" & pre_gen_len$V3 <=100 & pre_gen_len$assembly == "reference"  & pre_gen_len$redundancy == "NR90"]) /
df_big_len$ref[df_big_len$species =="Bodo_saltans" & df_big_len$redundancy == "NR90"]


sum(pre_gen_len$V3[pre_gen_len$species =="Bodo_saltans" & pre_gen_len$V3 <=100 & pre_gen_len$assembly == "de novo" & pre_gen_len$redundancy == "NR90"]) /
df_big_len$ref[df_big_len$species =="Bodo_saltans" & df_big_len$redundancy == "NR90"]


```


## Contig(CDS or transcript)-level length distribution


```{r}
contig_len <- fread("Nucleotides_ORIGINAL_length.tsv")

contig_len$redundancy <- "ORIGINAL"

contig_len$assembly <- NA
contig_len$assembly[grep("denovo", contig_len$V1)] <- "de novo"
contig_len$assembly[grep("REF", contig_len$V1)] <- "reference"
contig_len$assembly[grep("NON", contig_len$V1)] <- "reference"

contig_len$species <- NA
for(sp in unique(big_df$species)) {
  contig_len$species[grep(sp, contig_len$V1) ] <- sp 
  }

contig_len$species[grep("Trypanosoma_cruzi_CLBrenerEsmeraldo-like", contig_len$V1) ] <- "Trypanosoma_cruzi_CLBrenerEsmeraldo-like" 


```

```{r}

gg <- ggplot(contig_len , aes(V3 ) )+
  geom_histogram(aes(fill = interaction(assembly)), alpha=0.5, position = "identity")+scale_x_log10()+
  geom_vline(xintercept=300, color = "red")+
  # scale_color_manual(values = colors[c(38,48,23,4)] ) + scale_fill_manual(values = colors[c(38,48,23,4)] ) +
  theme_bw()+ facet_wrap(~species,scales = "free_y")

ggsave("all_denov_vs_ref_Length_PER_GENE_NUCLEOTIDE.png", gg, width = 25, height = 10)


ggplot(pre_gen_len[pre_gen_len$species =="Bodo_saltans",] , aes(V3 ) )+
  geom_histogram(aes(fill = redundancy), alpha=0.25, position = "identity")+scale_x_log10()+
  #geom_text(aes(label = value), hjust=-0.5, size = 2) +
  theme_bw()+ facet_wrap(~ assembly,scales = "free_y")

```


## Length plots for the preprint
```{r}
gg1 <- ggplot(pre_gen_len[pre_gen_len$redundancy=="NR90",] , aes(V3 ) )+
  geom_histogram(aes(fill = interaction(assembly,redundancy)), alpha=0.5, position = "identity")+scale_x_log10()+
#  scale_color_manual(values = colors[c(38,48,23,4)] ) + scale_fill_manual(values = colors[c(38,48,23,4)] ) +
  theme_bw()+ facet_wrap(~species,scales = "free_y")+
  xlab("Protein length (aa), log10")+
  ggtitle("Distribution of protein lengths")


gg2 <- ggplot(contig_len , aes(V3 ) )+
  geom_histogram(aes(fill = interaction(assembly)), alpha=0.5, position = "identity")+scale_x_log10()+
  geom_vline(xintercept=300, color = "red")+
  # scale_color_manual(values = colors[c(38,48,23,4)] ) + scale_fill_manual(values = colors[c(38,48,23,4)] ) +
  theme_bw()+ facet_wrap(~species,scales = "free_y")+
  xlab("CDS/transcript length (bp), log10")+
  ggtitle("Distribution of contig (CDS or transcript) lengths")

gg <-(gg1 + theme(legend.position="none")) / gg2  

#gg <- grid.arrange(gg1, gg2, ncol=1)

ggsave("lengths_BUSCO_distr.png", gg, width = 26, height = 13)

```





## Different factors influeancing lengths and completeness of BUSCO assemblies

```{r}

factors <- fread("denovo_VS_ref_comp.csv")


ggplot(factors, aes(BUSCO_EUK_diff, ref_vs_denov) )+
     geom_text(aes(label=species), size =2.5)+
  geom_point(aes(colour=Less_proteins, shape = Less_contigs), size =3)+
  theme_bw()#+ facet_wrap(~ assembly,scales = "free_y")



ggplot(factors, aes(BUSCO_EUK_diff, ref_vs_denov) )+
     geom_text(aes(label=species), size =2.5)+
  geom_point(aes(colour=Short_proteins, shape = Short_contigs), size =3)+
  theme_bw()#+ facet_wrap(~ assembly,scales = "free_y")



ggplot(factors, aes(BUSCO_EUK_denov_value, ref_vs_denov) )+
     geom_text(aes(label=species), size =2.5)+
  geom_point(aes(colour=Short_proteins, shape = Short_contigs), size =3)+
  theme_bw()#+ facet_wrap(~ assembly,scales = "free_y")


```












