---
title: "Untitled"
author: "Zavadska Daryna"
date: "2025-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, echo=FALSE , message=FALSE}
##upload packages
library(stringr)
library(openxlsx)
library(data.table)
library(magrittr)
#library(GGally)
#library(propr)
library(data.table)
library(ggplot2)
library(reshape2)
library(tidyverse)
library(dplyr)
library(plyr)
library(microshades)
library(scales)
library(ggbreak) 
library(patchwork)
library(ggpubr)
library(gplots)
library(PerformanceAnalytics)
library(microshades)
library(gridExtra)
```

```{r}
colors <-c(microshades_palette("micro_blue", lightest = FALSE), 
           microshades_palette("micro_purple", lightest = FALSE), microshades_palette("micro_cvd_purple", lightest = FALSE), 
           microshades_palette("micro_green", lightest = FALSE), 
           microshades_palette("micro_cvd_green", lightest = FALSE),
           microshades_palette("micro_orange", lightest = FALSE), 
           microshades_palette("micro_brown", lightest = FALSE), 
           microshades_palette("micro_cvd_turquoise", lightest = FALSE),
           microshades_palette("micro_cvd_gray", lightest = FALSE))

```






```{r}
suffix <- "V3"

per_sp <- fread(paste0("/home/dzavadska/Data/QUANTITATIVE_DATASET/BUSCO_TITRATION_CLEAN_NOV2025/",suffix,"Statistics_PerSpecies.tsv"))

per_sp_molten <- melt(per_sp)

per_sp2 <- dcast(per_sp_molten, variable~V1)

groups <- fread("vocabulary_groups.csv", header = F)
colnames(groups)<- c("variable", "group")
per_sp_molten <- merge(per_sp_molten, groups, "variable")

per_sp_molten$variable <- factor(
    per_sp_molten$variable,
    levels = unique(per_sp_molten$variable[order(per_sp_molten$group)])
)



gg5 <- ggplot(per_sp_molten[grep("Number", per_sp_molten$V1),], aes( V1,variable, fill=value))+geom_tile()+
   facet_grid(group ~ ., scales="free_y", space="free_y") +
  theme_classic()+theme(axis.text.x = element_text(size=5,angle=90,hjust=0.99,vjust=0.1), axis.text.y = element_text(size=5))+scale_fill_gradientn(colors = hcl.colors(20, "RdYlGn"))+ggtitle(suffix)

ggsave(paste0("per_sp_num", suffix, ".png"), gg5)

gg6 <- ggplot(per_sp_molten[grep("Percentage", per_sp_molten$V1),], aes( V1,variable, fill=value))+geom_tile()+
     facet_grid(group ~ ., scales="free_y", space="free_y") +
  theme_classic()+theme(axis.text.x = element_text(size=6,angle=90,hjust=0.99,vjust=0.1), axis.text.y = element_text(size=5))+scale_fill_gradientn(colors = hcl.colors(20, "RdYlGn"))+ggtitle(suffix)

ggsave(paste0("per_sp_percent", suffix, ".png"),  gg6)

```




## All summary

```{r}

list <- c("V3")

OG_summary <- data.frame()
sg_summary <- data.frame()
sg_all <- data.frame()

for (suffix in list) {
  path <- paste0("/home/dzavadska/Data/QUANTITATIVE_DATASET/BUSCO_TITRATION_CLEAN_NOV2025/", suffix, "Orthogroups.GeneCount.tsv")
  df <- fread(path)

  df_molten <- melt(df)
  #######
  
  df_molten$value <- as.numeric(df_molten$value)
  df_molten$Orthogroup <- as.factor(df_molten$Orthogroup)

  OG_summary <- df_molten %>%
  group_by(Orthogroup) %>%
  dplyr::summarize(
    total = value[variable == "Total"], # Assumes "Total" exists once per Orthogroup
    sd = sd(value[variable != "Total"]),
    mean = mean(value[variable != "Total"]),
    sp_repr = length(which(value!=0))-1 ,
    setting = suffix
  )

write.csv(OG_summary, paste0("OG_summary_", suffix, ".csv"))

#########
subset <- OG_summary[which(OG_summary$mean<=1 & OG_summary$sd<=1 & OG_summary$sp_repr>25),]

subset2 <- df_molten[df_molten$Orthogroup %in% subset$Orthogroup,]
colnames(subset2) <-c("Orthogroup", "variable", "value" ) 


mergedsubset <- merge(subset,subset2,  by = "Orthogroup")
mergedsubset <- mergedsubset[mergedsubset$variable!="Total",]
mergedsubset$variable <- str_replace(mergedsubset$variable,".cdhit_pep97", " ")
#########
print(suffix)
print(length(unique(mergedsubset$Orthogroup)) )
print(length(unique(mergedsubset$Orthogroup[as.integer(mergedsubset$sp_repr)==mergedsubset$total])) )

sg <- mergedsubset[mergedsubset$Orthogroup %in% unique(mergedsubset$Orthogroup[as.integer(mergedsubset$sp_repr)==mergedsubset$total]),]
#########

groups <- fread("vocabulary_groups.csv", header = F)
colnames(groups)<- c("variable", "group")

sg_sp_occr <- sg %>% group_by(variable) %>% dplyr::summarise(occurence_in_sg = sum(value))



sg_sp_occr <- merge(sg_sp_occr, groups, "variable")
sg_sp_occr$settings <- suffix

sg_summary <- rbind(sg_summary, sg_sp_occr)


ogs <- sg %>% group_by(sp_repr) %>% dplyr::summarise(og_count = length(unique(Orthogroup)), settings = suffix)
sg_all <- rbind(sg_all, ogs)
}


sum_plot <- ggplot(sg_summary, aes(variable, occurence_in_sg))+geom_point(aes(shape = group, colour=settings), size=2, alpha=0.7)+theme(axis.text.x = element_text(angle=90,hjust=0.99,vjust=0.1, size = 7))
ggsave("sum_plot.png", sum_plot)

ggplot(sg_all, aes(sp_repr, og_count))+geom_point(aes(shape=settings,colour=settings), size=2)+theme_classic() + scale_shape_manual(values=seq(0,15))
sg_all
#ogs <- sg %>% group_by(sp_repr) %>% dplyr::summarise(og_count = length(unique(Orthogroup)), settings = suffix)

```



## Filtering by OG

```{r}

OG_local <- fread("OG_summary_V3.csv")

subset <- OG_local[which(OG_local$mean<=1 & OG_local$sd<=1 & OG_local$sp_repr>30),]

double_subset <- subset[which(subset$mean>=0.9 & subset$sd<=0.5 & subset$sp_repr>30),]

fwrite(double_subset, "double_subset_V3.csv")
#OG_local[which(OG_local$mean<=1 & OG_local$sd==0 & OG_local$sp_repr>25),]

ggplot(subset, aes(sd))+geom_histogram()+theme_classic()
ggplot(subset, aes(sp_repr))+geom_histogram()+theme_classic()
ggplot(subset, aes(mean))+geom_histogram()+theme_classic()

ggplot(subset, aes(mean, sd))+geom_point(aes(size = sp_repr))+theme_classic()+theme_classic()

```

##Plots for the preprint


```{r}
library(ggVennDiagram)
library(cowplot)

sd <- ggplot(OG_local, aes(sd))+geom_histogram()+theme_classic()+scale_x_log10()+geom_vline(xintercept = 0.5, linetype="dashed", color = "red")+xlab("Standard deviation of gene count/species in OG")
mean <- ggplot(OG_local, aes(mean))+geom_histogram()+theme_classic()+scale_x_log10()+
  geom_vline(xintercept = 1, linetype="dashed", color = "red")+
  geom_vline(xintercept = 0.9, linetype="dashed", color = "red")+
  xlab("Mean gene count/species in OG")
sp <- ggplot(OG_local, aes(sp_repr))+geom_histogram()+theme_classic()+scale_y_log10()+geom_vline(xintercept = 30, linetype="dashed", color = "red")+xlab("Number of species represented in the OG")



#subset[which(subset$mean>=0.9 & subset$sd<=0.5 & subset$sp_repr>35),]


x <- list("Mean" = OG_local$Orthogroup[which(OG_local$mean<=1 & OG_local$mean>=0.9)], 
          "SD" = OG_local$Orthogroup[which(OG_local$sd<=0.5)],
          ">30 species" = OG_local$Orthogroup[which(OG_local$sp_repr>30 )] )

# length(intersect(x$A, intersect(x$B, x$C)))

venn <- 
  ggVennDiagram(x, label_alpha = 0)+
  scale_fill_gradient(low="grey90", high = "grey50") + theme(legend.position = "none")+scale_alpha_continuous(range=c(0.2,0.4))+
  theme(plot.margin=margin(20, 60, 20, 60)) + coord_fixed(ratio = 0.7) 
#  +  annotate("text", label =c("Mean<=1&>=0.9","SD<=0.5",">35 species in OG"), x = c(-1, 0.8, 0.2),y = c(-1.1, -0.8, 0.2), size = 12) 


top_row <- plot_grid(sp, mean, sd,  labels = c("A", "B", "C"), ncol = 3)
bottom_row <- plot_grid(venn, labels = "D", ncol = 1)

# Combine rows
panel_OG_subset <- plot_grid(top_row, bottom_row, ncol = 1, rel_heights = c(3, 5))


ggsave("panel_OG_subset.pdf", panel_OG_subset, width=12, height=8)


```



```{r}

suffix <- "V3"


path <- paste0("/home/dzavadska/Data/QUANTITATIVE_DATASET/BUSCO_TITRATION_CLEAN_NOV2025/", suffix, "Orthogroups.GeneCount.tsv")
df <- fread(path)

df <- df[,!"Total"]
df <- df[df$Orthogroup %in% unique(double_subset$Orthogroup),]
  
df_molten <- melt(df)
  #######
  
pseudobusco <- df_molten %>% dplyr:::count(variable, value, name = "occurrence")

ggplot(pseudobusco, aes(occurrence, variable, fill =  as.factor(value)) )+geom_bar(stat = "identity", width = 0.8)+
  theme_classic()+ theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
     xlab("Occurence in OG")+ylab("species") + labs(fill = "Copy number per OG")+
      scale_fill_brewer(palette = "Set3", limits = c("1","2","3","0"))
  
ggsave("PSEUDOBUSCO.png")



df_molten$value <- as.numeric(df_molten$value)
df_molten$Orthogroup <- as.factor(df_molten$Orthogroup)


```



### genes <200 / all genes lengths
```{r}

len_df <- fread("Ortho_NR90_length.tsv")

len_df$V1 <- str_replace(len_df$V1, ".fasta", "")


ggplot(len_df, aes(V3))+geom_histogram()+
  facet_wrap(~V1, scales="free_x")+scale_x_log10() +theme_classic()
ggsave("protein_length.png", height=10, width=10)


g65 <- len_df[len_df$V1 == "G65133",]


sum(g65$V3[g65$V3>200])
sum(g65$V3[g65$V3<200])

sum(g65$V3[g65$V3>200])/sum(g65$V3[g65$V3<200])


```








### Unassigned genes length/genes <200
```{r}

unass <- fread("Orthogroups_UnassignedGenes.tsv")

unass_m <- pivot_longer(unass, cols = -Orthogroup)

len_df$V2 <- str_replace(len_df$V2, " .*", "")
len_df$V2 <- str_replace_all(len_df$V2, "::", "__")

un_len <- merge(len_df, unass_m, by.x=c("V1","V2"), by.y=c("name","value") )

sum( un_len$V3[un_len$V1 == "G65133" & un_len$V3<200] ) / sum(g65$V3[g65$V3<200])

ggplot(un_len , aes(V3))+geom_histogram()+scale_x_log10() +theme_classic()+facet_wrap(~V1, scales="free_x")
 

```




```{r}

path <- paste0("/home/dzavadska/Data/QUANTITATIVE_DATASET/BUSCO_TITRATION_CLEAN_NOV2025/", suffix, "Orthogroups.GeneCount.tsv")
dupl <- fread(path)

dupl <- dupl[,!"Total"]
#df <- df[df$Orthogroup %in% unique(double_subset$Orthogroup),]
  
dupl_m <- melt(dupl)
  #######
  
pseudobusco <- dupl_m %>% dplyr:::count(variable, value, name = "occurrence")

ggplot(dupl_m[dupl_m$value>0,], aes(value) )+geom_histogram()+scale_x_log10()+
  facet_wrap(~variable, scales="free_y")+theme_classic()
ggsave("all_dupl.png",height=10, width=15 )

  

```





```{r}

#unique(un_len$Orthogroup[un_len$V1 == "G65133" & un_len$V3<200])

#unique(dupl_m$Orthogroup[dupl_m$variable == "G65133"])

#dupl_m[dupl_m$value < 2,]

```



## Singletons
run the following chunk in bash to obtain a list of singleton OGs (Orthogroups.tsv are obtained directly from OrthoFinder output):

!!!!!!!!!!!!!! FOLLOWING CHUNK TAKES A LOT OF TIME uncomment if need to re-create the table from orthifinder output !!!!!!!!!!!!!!

```{r}

 og <- read.delim("Orthogroups.tsv", sep="\t", check.names=FALSE)

# species_cols <- colnames(og)[-1]

# species_specific <- og %>% rowwise() %>%
#  mutate(non_empty = sum(c_across(all_of(species_cols)) != "" & !is.na(c_across(all_of(species_cols))))) %>%
#  filter(non_empty == 1) %>% ungroup() %>%
#  pivot_longer(cols = all_of(species_cols), names_to="Species", values_to="Genes") %>%
#  filter(Genes != "" & !is.na(Genes)) %>%
#  select(Orthogroup, Species)



# write.table(species_specific, "species_specific_OGs.tsv", sep="\t", quote=FALSE, row.names=FALSE)

```


```{r}
species_specific <- fread("species_specific_OGs.tsv")

#sanity check
length(species_specific$Orthogroup[species_specific$Species == "G65133"])==per_sp_molten$value[per_sp_molten$V1=="Number of species-specific orthogroups" & per_sp_molten$variable == "G65133"]

og_subs <-og[og$Orthogroup %in% species_specific$Orthogroup[species_specific$Species == "G65133"], ]


gen_list <- str_replace_all(g65$V2[g65$V3<200], " .*", "" )
gen_list <- str_replace_all(gen_list, "::", "__" )


singletons_genes <- data.frame()

for(i in gen_list) {
  index <- grep(i, og_subs$G65133 )
  if(length(index)>0) {
   OG <- og_subs$Orthogroup[index]
   singletons_genes <- rbind(singletons_genes, data.frame(OG, i))
  }
}

g65$V2 <- str_replace_all(g65$V2, " .*", "" )
g65$V2 <- str_replace_all(g65$V2, "::", "__" )

g65_short <- g65[g65$V3<200]

sum(g65_short$V3[g65_short$V2 %in% singletons_genes$i]) / sum(g65_short$V3 )


```

##loop for all species

```{r}

species_specific <- fread("species_specific_OGs.tsv")

short_all <- data.frame()

for (sp in unique(len_df$V1)) {

g65 <- len_df[len_df$V1 == sp,]

g65_short <- g65[g65$V3<200]
#sum(g65$V3[g65$V3>200])
#sum(g65$V3[g65$V3<200])

short_len <- sum(g65_short$V3)
short_fract <- sum(g65_short$V3)/sum(g65$V3)
short_unass_len <- sum( un_len$V3[un_len$V1 == sp & un_len$V3<200] )

og_subs <-og[og$Orthogroup %in% species_specific$Orthogroup[species_specific$Species == sp], ]


gen_list <- str_replace_all(g65$V2[g65$V3<200], " .*", "" )
gen_list <- str_replace_all(gen_list, "::", "__" )


singletons_genes <- data.frame()
unassembled_genes <- un_len$V2[un_len$V1 == sp & un_len$V3<200]

for(i in gen_list) {
  index <- grep(i, og_subs[,sp] )
  if(length(index)>0) {
   OG <- og_subs$Orthogroup[index]
   singletons_genes <- rbind(singletons_genes, data.frame(OG, i))
  }
}

g65$V2 <- str_replace_all(g65$V2, " .*", "" )
g65$V2 <- str_replace_all(g65$V2, "::", "__" )

short_singleton_len <- sum( g65_short$V3 [ g65_short$V2 %in% singletons_genes$i] )
short_both_groups_len <- sum( g65_short$V3 [ g65_short$V2 %in% intersect(singletons_genes$i, unassembled_genes) ] )

short_all <- rbind(short_all, data.frame(sp, short_len, short_fract, short_unass_len, short_singleton_len, short_both_groups_len))

}

fwrite(short_all, "short_proteins_all_species.tsv")
```

##Plot for the preprint

```{r}

short_all <- fread("short_proteins_all_species.tsv")

short_all$diff_both <- (short_all$short_len  - short_all$short_unass_len- short_all$short_singleton_len)
short_all$diff_unass <- (short_all$short_len - short_all$short_unass_len)
short_all$diff_singl <- (short_all$short_len - short_all$short_singleton_len)

short_all_m <- melt(short_all)


d1 <- ggplot(short_all_m[short_all_m$variable %in% c("short_unass_len", "short_singleton_len", "diff_both"),],
        aes(x = value, y = sp, fill = as.factor(variable))) +
        geom_bar(stat = "identity", width = 0.8) +
        theme_bw() + #facet_wrap(~interaction(assembly, busco_db), scales = "free_x") +
   xlab("Total lengths across <200 aa")+ylab("species") + labs(fill = "Is in OG?")+
      scale_fill_brewer(palette = "Set3")+theme(legend.position = "bottom")+
        ggtitle("Total lengths of proteins in singletons or unassigned to OG, for all proteins <200 aa")



d2 <- ggplot(short_all_m[short_all_m$variable %in% c("short_fract"),],
        aes(x = value, y = sp, fill = as.factor(variable))) +
        geom_bar(stat = "identity", width = 0.8) +
        theme_bw() + #facet_wrap(~interaction(assembly, busco_db), scales = "free_x") +
   xlab("Total lengths across <200 aa")+ylab("species") +theme(legend.position = "none")+
        ggtitle("Total lengths of proteins <200 aa/Total proteome length")

        
        
gg <- grid.arrange(d1,d2,nrow=1 )


ggsave("short_prot_OGs.png", gg, width=15, height = 8)


                
```














### (single)Protein lengths per OG - to check for i.e. partials - especially, in KUCKIN
```{r}


og_subs2 <- og[og$Orthogroup %in% unique(df_molten$Orthogroup), ]

og_subs2_m <- melt(og_subs2, id.vars = "Orthogroup")

#merge(og_subs2_m, len_df)

#uncomment the part below, otherwise used saved table (takes time to run)

#pseudobusco_len_genes <- data.frame()

#for(i in unique(len_df$V2)) {
#  index <- grep(i, og_subs2_m$value)
  
#  if(length(index)>0) {
#   OG <- og_subs2_m$Orthogroup[index]
#   len <- len_df$V3[len_df$V2 == i]
#   pseudobusco_len_genes <- rbind(pseudobusco_len_genes, data.frame(OG,len, i))
#  }
#}



pseudobusco_len_genes$sp <- NA



#for (sp in unique(len_df$V1)) {
#  genes <- len_df$V2[len_df$V1 == sp]
#  pseudobusco_len_genes$sp[pseudobusco_len_genes$i %in% genes] <- sp
#  }


#length(unique(pseudobusco_len_genes$OG))

#fwrite(pseudobusco_len_genes, "pseudobusco_len_genes.tsv")

```

```{r}


pseudobusco_len_genes <-fread("pseudobusco_len_genes.tsv")


pseudobusco_len_genes %>%
  group_by(OG, sp) %>%
  dplyr::summarize(
    mean_len = mean(len),
    sd_len = sd(len),
    gene_count = length(len),
    .groups = "drop"
  ) -> pseudobusco_len_summary


unique(pseudobusco_len_summary$sp)

pseudobusco_len_summary <- pseudobusco_len_summary[pseudobusco_len_summary$OG != "",]

kuckin <- len_df[len_df$V1 == "KUCKIN",]

ggplot(pseudobusco_len_summary, aes(mean_len) )+geom_histogram()+scale_x_log10()+
  facet_wrap(~sp, scales="free_y")+theme_classic()

ggsave("pseudobusco_len_summary.png",height=10, width=15 )

 
ggplot(pseudobusco_len_summary[pseudobusco_len_summary$gene_count>1,], aes(mean_len) )+geom_histogram()+#scale_x_log10()+
  facet_wrap(~sp, scales="free_y")+theme_classic()
ggsave("pseudobusco_DUPLICATElen_summary.png",height=10, width=15 )


ggplot(pseudobusco_len_summary[pseudobusco_len_summary$gene_count>1,], aes(sd_len) )+geom_histogram()+#scale_x_log10()+
  facet_wrap(~sp, scales="free_y")+theme_classic()
ggsave("pseudobusco_DUPLICATElenSD_summary.png",height=10, width=15 )


```



```{r}



```











