---
title: "Untitled"
author: "Zavadska Daryna"
date: "2025-04-03"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = "ragg_png")
```

```{r}
library("tidyverse")
library("ggplot2")
library("data.table")
library("scales")
library("ggforce")
library("dgof")
library("remotes")
library("smmr")
#library("coin")

#library("rcompanion")

library("agricolae")
library("gridExtra")

library(cluster)    # clustering algorithms
```


# FUNCTIONS


```{r}

fluo_preprocess_adjustable <- function(all_meas, debugplot=FALSE, slideORseries = "slide", outlier_method = "IQR") {
all_meas_quart <- data.frame()

if(slideORseries == "slide")  { LIST <- unique(all_meas$file_prefix) } 
  else if(slideORseries == "series") { LIST <- unique(all_meas$exp_series) }
  else{print("slideORseries must be either slide or series - no other options are supported so far.")}
    
  for (UNIT in LIST) {
  
  if(slideORseries == "slide")  {
    one_sp1 <- all_meas[all_meas$file_prefix == UNIT,] } 
    else if(slideORseries == "series") {
    one_sp1 <- all_meas[all_meas$exp_series == UNIT,]
  }



for (sp in unique(one_sp1$est_sp)) {
  
  one_sp <- one_sp1[one_sp1$est_sp == sp, ]

  quant_filtered_df <- one_sp
  
  if (outlier_method == "QUART") {
    ##### filtering quartiles
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int >= summary(one_sp$nucl_sum_int)[2]), ]
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int <= quantile(one_sp$nucl_sum_int)[4]), ]
  }
  else if(outlier_method == "IQR") {
      ##### filtering outliers
       qnt <- quantile(quant_filtered_df$nucl_sum_int, probs=c(.25, .75))
   H <- 1.5*IQR(quant_filtered_df$nucl_sum_int)
   quant_filtered_df <- quant_filtered_df[intersect(which(quant_filtered_df$nucl_sum_int < (qnt[2] + H) ) , which(quant_filtered_df$nucl_sum_int > (qnt[1] - H) ) ), ]
 
  }
  else(print("outlier_method should be either IQR or QUART(default), no other values currently allowed!"))

  
  all_meas_quart <- rbind(all_meas_quart, quant_filtered_df)
  
    
  if(debugplot == TRUE){
  print( ggplot(one_sp, aes(est_sp, as.numeric(nucl_sum_int)))+geom_boxplot()+
     geom_jitter(position=position_jitter(height=0, width=0.2, seed = 1234), alpha = 0.7)+ #geom_point(aes(shape = sp, colour = file_prefix), size = 2, alpha = 0.7)+
    theme_classic()+ scale_y_continuous(labels = label_comma())+
    theme(axis.text.x = element_text(angle =90))+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+ggtitle(paste0("After filtering", UNIT)) )

  print( ggplot(quant_filtered_df, aes(est_sp, as.numeric(nucl_sum_int)))+geom_boxplot()+
     geom_jitter(position=position_jitter(height=0, width=0.2, seed = 1234), alpha = 0.7)+ #geom_point(aes(shape = sp, colour = file_prefix), size = 2, alpha = 0.7)+
    theme_classic()+ scale_y_continuous(labels = label_comma())+
    theme(axis.text.x = element_text(angle =90))+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+ggtitle(paste0("After filtering", UNIT)) )
  }
  
}
}
return(all_meas_quart)
}
```


```{r, include=FALSE}

fluo_clustering_combiner_and_filtrator <- function(sp_by_sp_quart, combination_factors = c("est_sp", "slidedate",  "exp_series_date", ""), outlier_method = "IQR" ) {
  
column_combinations <- combn(combination_factors, 2)
## !!! SUBSETTING K-MEAN CLUSTERING OU ONLY TO THAT WITHIN SPECIES (not permitting across-species clusterisation, as it makes no sense)!!!!
column_combinations <- column_combinations[, which(column_combinations[1,]=="est_sp")]
## !!! SUBSETTING K-MEAN CLUSTERING OU ONLY TO THAT WITHIN SPECIES!!!!

kmean_comb_out <- data.frame()


for (i in 1:ncol(column_combinations)) {
  #debug
 #print(paste(column_combinations[1, i], column_combinations[2, i]))
  
  for (est_sp in unique(sp_by_sp_quart[, column_combinations[1, i]])) {
    
    #if ("sp" %in% column_combinations[1, i]) { CLUSTER <- 2 }else{ CLUSTER <- 2 }
    
    dd_comb <- data.frame()
    
    if (column_combinations[2, i] %in% c("", NA)) {  # Handle missing column case properly
      local <- sp_by_sp_quart[sp_by_sp_quart[, column_combinations[1, i]] == est_sp,]
      
      
############## 
#  print(column_combinations[, i])
#  print(unique(local$est_sp))
#  print(nrow( local[local[, column_combinations[1, i]] == est_sp, ] ) > 2 )      
############## Pre-filtering      
  quant_filtered_df <- local
  
  if (outlier_method == "QUART") {
    ##### filtering quartiles
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int >= summary(local$nucl_sum_int)[2]), ]
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int <= quantile(local$nucl_sum_int)[4]), ]
  }
  else if(outlier_method == "IQR") {
      ##### filtering outliers
       qnt <- quantile(quant_filtered_df$nucl_sum_int, probs=c(.25, .75))
   H <- 1.5*IQR(quant_filtered_df$nucl_sum_int)
   quant_filtered_df <- quant_filtered_df[intersect(which(quant_filtered_df$nucl_sum_int <= (qnt[2] + H) ) , which(quant_filtered_df$nucl_sum_int >= (qnt[1] - H) ) ), ]
 
  }
  else(print("outlier_method should be either IQR or QUART(default), no other values currently allowed!"))
     
  local <- quant_filtered_df
   
############## 
############## Clustering      
      if (nrow(local[local[, column_combinations[1, i]] == est_sp, ]) > 2) {
        # Clustering
        km <- kmeans(local$nucl_sum_int[local[, column_combinations[1, i]] == est_sp], 2)
        dd <- cbind(local[local[, column_combinations[1, i]] == est_sp, ], cluster = km$cluster)
        
        # Determine cluster labels
        dipl <- dd$cluster[dd$nucl_sum_int == max(dd$nucl_sum_int)][1]
        dd$ploidy_clust <- ifelse(dd$cluster == dipl, "2n", "1n")
        
        
        dd$suns_for_clust <- paste(column_combinations[1, i], column_combinations[2, i])
############## Clustering 

############## Post-filtering 
        dd_backup <- dd
        
        for(ploidy in unique(dd_backup$ploidy_clust)){
           quant_filtered_df <- dd_backup[dd_backup$ploidy_clust == ploidy, ]
  
         if (outlier_method == "QUART") {
    ##### filtering quartiles
         quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int >= summary(local$nucl_sum_int)[2]), ]
         quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int <= quantile(local$nucl_sum_int)[4]), ]
          }
         else if(outlier_method == "IQR") {
      ##### filtering outliers
         qnt <- quantile(quant_filtered_df$nucl_sum_int, probs=c(.25, .75))
         H <- 1.5*IQR(quant_filtered_df$nucl_sum_int)
        quant_filtered_df <- quant_filtered_df[intersect(which(quant_filtered_df$nucl_sum_int <= (qnt[2] + H) ) , which(quant_filtered_df$nucl_sum_int >= (qnt[1] - H) ) ), ]
 
       }
          else(print("outlier_method should be either IQR or QUART(default), no other values currently allowed!"))
     
          #dd <- quant_filtered_df 
          dd_comb <- rbind(dd_comb, quant_filtered_df)
        }
 
   
############## 
        

    
      } else {
        next
      }
    } else {
      for (slide in unique(sp_by_sp_quart[, column_combinations[2, i]])) {
        local <- sp_by_sp_quart[sp_by_sp_quart[, column_combinations[1, i]] == est_sp,]
        local <- local[local[, column_combinations[2, i]] == slide,]
##############
# print(column_combinations[, i])
# print(unique(local$est_sp))
# print(nrow( local[local[, column_combinations[1, i]] == est_sp, ] ) > 2 )
############## Pre-filtering      
  quant_filtered_df <- local

  if (outlier_method == "QUART") {
    ##### filtering quartiles
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int >= summary(local$nucl_sum_int)[2]), ]
  quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int <= quantile(local$nucl_sum_int)[4]), ]
  }
  else if(outlier_method == "IQR") {
      ##### filtering outliers
       qnt <- quantile(quant_filtered_df$nucl_sum_int, probs=c(.25, .75))
   H <- 1.5*IQR(quant_filtered_df$nucl_sum_int)
    quant_filtered_df <- quant_filtered_df[intersect(which(quant_filtered_df$nucl_sum_int <= (qnt[2] + H) ) , which(quant_filtered_df$nucl_sum_int >= (qnt[1] - H) ) ), ]
 
  }
  else(print("outlier_method should be either IQR or QUART(default), no other values currently allowed!"))
     
  local <- quant_filtered_df
   
############## 
############## Clustering      
        if (nrow(local[local[, column_combinations[1, i]] == est_sp,]) > 2) {
          # Clustering
          km <- kmeans(local$nucl_sum_int[local[, column_combinations[1, i]] == est_sp], 2)
          dd <- cbind(local[local[, column_combinations[1, i]] == est_sp, ], cluster = km$cluster)
          
          # Determine cluster labels
          dipl <- dd$cluster[dd$nucl_sum_int == max(dd$nucl_sum_int)][1]
          dd$ploidy_clust <- ifelse(dd$cluster == dipl, "2n", "1n")
          
          dd$suns_for_clust <- paste(column_combinations[1, i], column_combinations[2, i])
############## Clustering            
############## Post-filtering      
#  print(unique(quant_filtered_df$est_sp))

        dd_backup <- dd
        
        for(ploidy in unique(dd_backup$ploidy_clust)){
           quant_filtered_df <- dd_backup[dd_backup$ploidy_clust == ploidy, ]
  
         if (outlier_method == "QUART") {
    ##### filtering quartiles
         quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int >= summary(local$nucl_sum_int)[2]), ]
         quant_filtered_df <- quant_filtered_df[which(quant_filtered_df$nucl_sum_int <= quantile(local$nucl_sum_int)[4]), ]
          }
         else if(outlier_method == "IQR") {
      ##### filtering outliers
         qnt <- quantile(quant_filtered_df$nucl_sum_int, probs=c(.25, .75))
         H <- 1.5*IQR(quant_filtered_df$nucl_sum_int)
        quant_filtered_df <- quant_filtered_df[intersect(which(quant_filtered_df$nucl_sum_int <= (qnt[2] + H) ) , which(quant_filtered_df$nucl_sum_int >= (qnt[1] - H) ) ), ]
 
       }
          else(print("outlier_method should be either IQR or QUART(default), no other values currently allowed!"))
     
          #dd <- quant_filtered_df 
          dd_comb <- rbind(dd_comb, quant_filtered_df)
        }
        
          
        } else {
          next
        }
      }
    }
    
    kmean_comb_out <- rbind(kmean_comb_out, dd_comb)
  }
}


kmean_comb_out$bp_exp_ploidy_corr <- NA

kmean_comb_out$bp_exp_ploidy_corr[kmean_comb_out$ploidy_clust =="1n"] <- kmean_comb_out$bp_exp[kmean_comb_out$ploidy_clust =="1n"] 
kmean_comb_out$bp_exp_ploidy_corr[kmean_comb_out$ploidy_clust =="2n"] <- kmean_comb_out$bp_exp[kmean_comb_out$ploidy_clust =="2n"]*2


return(kmean_comb_out)
}



```






# PREPROCESSING

### Expected ratios dataframe

```{r, include=FALSE}
expected_ratios <- data.frame("sp"= c("Diplo_1", "Diplo_2", "Capsa_1", "Capsa_2", "Cflexa", "Sros_1", "Sros_2", "Theca", "Yeast"),
"bp" =  c(280283133, 107915492, 27967784, 27398043, 56404751, 55440309, 55448948, 28680627, 11529746) )


species_1 <- c()
species_2 <- c()
median_ratio <- c()

for ( i in unique(expected_ratios$sp)) {
  for ( e in unique(expected_ratios$sp)) {
    species_1 <- c(species_1, i)
      species_2 <- c(species_2, e)
    
  median_1 <- median(as.numeric(expected_ratios$bp[which(expected_ratios$sp == i)]))
    median_2 <- median(as.numeric(expected_ratios$bp[which(expected_ratios$sp == e)]))
  median_ratio <- c(median_ratio, median_1/median_2)
}
}

exp_rat <- data.frame("sp_1" = species_1, "sp_2" = species_2, "median_ratio_observed" = median_ratio)
exp_rat_backup <- exp_rat

```

### Expected genome sizes dataframe

```{r}

expected_sizes_full <- data.frame("sp"= c("Diplo_1", "Diplo_2", "Capsa_1", "Capsa_2", "Cflexa", "Sros_1", "Sros_2", "Theca", "Yeast"),
"bp" =  c(280283133, 107915492, 27967784, 27398043, 56404751, 55440309, 55448948, 28680627, 11529746) )

expected_sizes_trunc <- data.frame("sp"= c("Diplo", "Capsa", "Cflexa", "Sros", "Theca", "Yeast"),
"bp" =  c(280283133, 27398043, 56404751, 55448948, 28680627, 11529746) )

```



set path to the folder containing subfolders with measurements in your filesystem
```{r}

MEAS_ANALYSIS <- "/home/darynazavadska/Data/kinetos_born_again/MEAS_ANALYSIS/"

```


```{r}
#list <- c("25022025","1022025", "14022025", "21122024" )
list <- c("20032025_thickstacks_noHU", "20032025_thinstacks_noHU")


all_meas <- data.frame()

for (suffix in list) {
  

KIN_nuclDNA_raw_meas <- list.files(paste0(MEAS_ANALYSIS, suffix), pattern = "KIN_nuclDNA")
KIN_nuclDNA_all_meas <- data.frame()

for (i in KIN_nuclDNA_raw_meas) {
  df <- fread(paste0(MEAS_ANALYSIS, suffix, "/", i))
  KIN_nuclDNA_all_meas <- rbind(KIN_nuclDNA_all_meas, df)
}
KIN_nuclDNA_all_meas$estimate <- "KIN_nuclDNA"
KIN_nuclDNA_all_meas <- KIN_nuclDNA_all_meas[!is.na(KIN_nuclDNA_all_meas$sp),]
KIN_nuclDNA_all_meas$exp_series <- suffix

############
kDNA_raw_meas <- list.files(paste0(MEAS_ANALYSIS, suffix), pattern = "kDNA_df")
kDNA_all_meas <- data.frame()

for (i in kDNA_raw_meas) {
  df <- fread(paste0(MEAS_ANALYSIS, suffix, "/", i))
  kDNA_all_meas <- rbind(kDNA_all_meas, df)
}
kDNA_all_meas$estimate <- "kDNA_df"
kDNA_all_meas <- kDNA_all_meas[!is.na(kDNA_all_meas$sp),]
kDNA_all_meas$exp_series <- suffix
##############
nucl_df_raw_meas <- list.files(paste0(MEAS_ANALYSIS, suffix), pattern = "nucl_df")
nucl_df_all_meas <- data.frame()

for (i in nucl_df_raw_meas) {
  df <- fread(paste0(MEAS_ANALYSIS, suffix, "/", i))
  nucl_df_all_meas <- rbind(nucl_df_all_meas, df)
}
nucl_df_all_meas$estimate <- "nucl_df"
nucl_df_all_meas <- nucl_df_all_meas[!is.na(nucl_df_all_meas$sp),]
nucl_df_all_meas$exp_series <- suffix
###

nucl_df_all_meas_to_bind <- nucl_df_all_meas[,c( "V1", "id", "cell", "sp", "cell_coord", "cyto_mean_int", "nucl_sum_int", "nucl_sum_int_corr", "estimate", "exp_series" )]
all_meas <- rbind(data.table(all_meas), nucl_df_all_meas_to_bind, KIN_nuclDNA_all_meas, kDNA_all_meas,  use.names=FALSE)


}
###

if (length(grep("ch1", all_meas$id))<nrow(all_meas)) {
  print("Warning! Some channels might have been incorrectly assigned!")
  print("check following images:")
  print(unique(all_meas$id[!str_detect(all_meas$id, "ch1")]))
}

all_meas$est_sp <- paste0(all_meas$sp, all_meas$estimate)

all_meas$expected_sizes <- NA 

for (sp in expected_sizes_trunc$sp) {
  all_meas$expected_sizes[all_meas$sp ==sp] <- expected_sizes_trunc$bp[expected_sizes_trunc$sp==sp]
}

all_meas_backup <- all_meas

```


### Adding staining parameters

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
all_meas <- data.frame(all_meas)

all_meas$exp_series_date <- NA
all_meas$file_prefix <- NA
all_meas$AT_stain <- NA
all_meas$AT_stain_conc <- NA
all_meas$quant_stain <- NA
all_meas$quant_stain_conc <- NA
all_meas$AT_stain_laser_int <- NA
all_meas$AT_stain_exp <- NA
all_meas$quant_stain_laser_int <- NA
all_meas$quant_stain_exp <- NA
all_meas$replicate <- NA
all_meas$other <- NA
all_meas$test_species <- NA
all_meas$images_taken <- NA
all_meas$images_processed <- NA

parameters <- fread("UPDATED_images_taken_data.csv")
# !!!!!!!!!!! SUBSET PARAMETERS IF NEEDED (IN CASE OF IDENTICAL NAMES IN DIFFERENT SERIES TROWS ERRORS - BECAUSE YOU SHOULDNT NAME THEM THAT WAY)



for(pref in unique(parameters$file_prefix)){
  
  all_meas[grep(paste0(pref), all_meas$id), 13:ncol(all_meas)] <- parameters[grep(paste0("\\b", pref, "\\b"), parameters$file_prefix)]
}

#all_meas[grep("Gemkin", all_meas$sp), ]
#all_meas[grep(paste0(pref), all_meas$id),]

```





### Subsetting dataset

```{r, results='hide', message=FALSE, warning=FALSE}
sp_by_sp <- all_meas
#sp_by_sp <- sp_by_sp[-grep("_B" , sp_by_sp$file_prefix ),]

sp_by_sp <- sp_by_sp[sp_by_sp$quant_stain_exp == 12.5, ]
sp_by_sp <- sp_by_sp[sp_by_sp$quant_stain == "PI",]

#by amount of slides processed
#tmp <- to_see[to_see$file_prefix %in% unique(sp_by_sp$file_prefix),]
#sp_by_sp$images_processed <- tmp$images_processed[match(sp_by_sp$file_prefix, tmp$file_prefix)]


sp_by_sp <- sp_by_sp[which(sp_by_sp$sp != "Theca"),]

sp_by_sp <- sp_by_sp[which(sp_by_sp$sp != "Cflexa"),]

sp_by_sp <- sp_by_sp[which(sp_by_sp$sp != "Cruz"),]

#sp_by_sp <- sp_by_sp[sp_by_sp$images_processed >= 5,]

sp_by_sp <- sp_by_sp[-grep("Gemkin", sp_by_sp$test_species ),]

```



### Matching with bp_exp and backing up

```{r, results='hide', message=FALSE, warning=FALSE}

sp_by_sp$bp_exp <- expected_sizes_trunc$bp[match(sp_by_sp$sp, expected_sizes_trunc$sp)]
sp_by_sp$slidedate <- paste(sp_by_sp$file_prefix, sp_by_sp$exp_series_date)

sp_by_sp_backup <- sp_by_sp

# unique(sp_by_sp_backup$exp_series[grep("Gemkin",sp_by_sp_backup$sp)] )
# sp_by_sp_backup[grep("Gemkin",sp_by_sp_backup$sp) ,] 
```


# CLUSTERISATION

### preprocessing
In this version, we apply (pre-prediction) clusterisation to outlier-filtered controls, as well as to kinetos - therefore, subsetting to only control species for clusterisation and assigning "Undefined" ploidy in this case makes few sense

```{r}

sp_by_sp_by_kmeans <- fluo_clustering_combiner_and_filtrator(sp_by_sp, combination_factors = c("est_sp", "slidedate",  "exp_series_date", ""), outlier_method = "IQR" ) 


# older version below, sp_by_sp_quart still used for pruducing unclustered dataset
 sp_by_sp_quart <- fluo_preprocess_adjustable(sp_by_sp, debugplot=FALSE, slideORseries = "series", outlier_method = "IQR")
# sp_by_sp_by_kmeans_unfilt <- fluo_clustering_combiner(sp_by_sp_quart)
# sp_by_sp_by_kmeans <- data.frame()
# for (ploidy in unique(sp_by_sp_by_kmeans_unfilt$ploidy_clust)) {
#  tmp <- sp_by_sp_by_kmeans_unfilt[sp_by_sp_by_kmeans_unfilt$ploidy_clust == ploidy, ]
#  tobind <- fluo_preprocess_adjustable(tmp, debugplot=FALSE, slideORseries = "series", outlier_method = "IQR")
#  sp_by_sp_by_kmeans <- rbind(sp_by_sp_by_kmeans, tobind) }



tmp <- sp_by_sp_quart
tmp$cluster <-  "Undefined" 
tmp$ploidy_clust <-  "Undefined" 
tmp$suns_for_clust <-  "NOT CLUSTERED" 
tmp$bp_exp_ploidy_corr <-  tmp$expected_sizes

sp_by_sp_by_kmeans_plus_uncl <- rbind(tmp, sp_by_sp_by_kmeans)

#in this version - must print both kinetos and control species
unique(sp_by_sp_by_kmeans$est_sp)

#debug for ploidy correction
#sp_by_sp_by_kmeans_plus_uncl[sp_by_sp_by_kmeans_plus_uncl$est_sp=="Gemkin_KINnucl_df",c("ploidy_clust", "suns_for_clust")]
```

### check how the outlier filtering and clusterisation went


```{r}

sp_by_sp_by_kmeans_plus_uncl$stack <- NA
sp_by_sp_by_kmeans_plus_uncl$stack[grep("thinstack" , sp_by_sp_by_kmeans_plus_uncl$file_prefix)] <- "37 slices"
sp_by_sp_by_kmeans_plus_uncl$stack[grep("thickstack" , sp_by_sp_by_kmeans_plus_uncl$file_prefix)] <- "8 slices"

sp_by_sp_by_kmeans_plus_uncl <- sp_by_sp_by_kmeans_plus_uncl[grep("KINnucl_df", sp_by_sp_by_kmeans_plus_uncl$est_sp, invert = TRUE),]

ggplot(sp_by_sp_by_kmeans_plus_uncl[sp_by_sp_by_kmeans_plus_uncl$suns_for_clust == "est_sp exp_series_date",], aes(interaction(stack,ploidy_clust), nucl_sum_int))+
  geom_violin()+geom_boxplot(alpha = 0)+
  geom_point(aes( colour=sp ), size =0.5, alpha = 0.5, position=position_jitter(height=0, width=0.2))+
  theme_bw()+ ggtitle("est_sp slidedate")+theme(axis.text.x = element_blank())+
  facet_wrap(~ interaction(sp,estimate))+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))


ggplot(sp_by_sp_by_kmeans_plus_uncl[sp_by_sp_by_kmeans_plus_uncl$suns_for_clust == "est_sp exp_series_date" & sp_by_sp_by_kmeans_plus_uncl$sp == "Capsa",], aes(interaction(stack,ploidy_clust), nucl_sum_int))+
  geom_violin()+geom_boxplot(alpha = 0)+
  geom_point(aes( colour=sp ), size =0.5, alpha = 0.5, position=position_jitter(height=0, width=0.2))+
  theme_bw()+ ggtitle("est_sp slidedate")+theme(axis.text.x = element_blank())+
  facet_wrap(~ interaction(sp,estimate))+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))


stack_validation <- ggplot(sp_by_sp_by_kmeans_plus_uncl[sp_by_sp_by_kmeans_plus_uncl$suns_for_clust == "est_sp exp_series_date",], aes(interaction(ploidy_clust,stack), nucl_sum_int))+
  geom_violin()+geom_boxplot(alpha = 0)+
  geom_point(size =0.5, alpha = 0.5, position=position_jitter(height=0, width=0.2))+
  theme_bw()+theme(axis.text.x = element_text(angle = 90))+
  facet_wrap(~ sp+estimate, nrow = 1,
             labeller = labeller( sp = c(Rhyncho_KIN = "Rhyncho", 
                                          Capsa = "Capsaspora", 
                                          Diplo = "Diplonema",
                                          Yeast = "Yeast"),
              estimate = c(kDNA_df = "kDNA", KIN_nuclDNA = "nuclear DNA", nucl_df = "")
            )   )+
  scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+ 
  labs( x = "Cell cycle stage, number of slices in stack", y = "Fluorescence intensity")
  

ggsave(stack_validation, filename = "stack_validation.pdf", width = 10, height = 5)


```




```{r}

ggplot(sp_by_sp_by_kmeans[sp_by_sp_by_kmeans$suns_for_clust == "est_sp slidedate",], aes(interaction(ploidy_clust, est_sp), nucl_sum_int))+geom_boxplot()+
  geom_point(aes( colour=sp ), size =1, alpha = 0.15, position=position_jitter(height=0, width=0.2))+
  theme_bw()+ ggtitle("est_sp slidedate")+theme(axis.text.x = element_blank())+
  facet_wrap(~suns_for_clust)+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))


ggplot(sp_by_sp_by_kmeans[sp_by_sp_by_kmeans$suns_for_clust == "est_sp exp_series_date",], aes(interaction(ploidy_clust, est_sp), nucl_sum_int))+geom_boxplot()+
  geom_point(aes( colour=sp ), size =1, alpha = 0.15, position=position_jitter(height=0, width=0.2))+
  theme_bw()+ ggtitle("est_sp slidedate")+theme(axis.text.x = element_blank())+
  facet_wrap(~suns_for_clust)+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))

ggplot(sp_by_sp_by_kmeans[sp_by_sp_by_kmeans$suns_for_clust == "est_sp ",], aes(interaction(ploidy_clust, est_sp), nucl_sum_int))+geom_boxplot()+
  geom_point(aes( colour=sp ), size =1, alpha = 0.15, position=position_jitter(height=0, width=0.2))+
  theme_bw()+ ggtitle("est_sp slidedate")+theme(axis.text.x = element_blank())+
  facet_wrap(~suns_for_clust)+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2), breaks = 2^(19:40))


```




# HYDROXYUREA

```{r}
#TOPLOT_bootstrap_out_clustered_HU.csv is produced initaillt by neat_knit_clustered_FUNCTION-BASED_parser.Rmd, as the default benchmarking set, TOPLOT_bootstrap_out_clustered.csv, and is manually renamed, but the corresponding folder with HU treatment measurements is included in the list in this case.

toplot <- fread("TOPLOT_bootstrap_out_clustered_HU.csv")

toplot$ploidy_clust <- str_replace(toplot$ploidy_clust, "1n", "G1")
toplot$ploidy_clust <- str_replace(toplot$ploidy_clust, "2n", "G2")

toplot$if_HU <- "no HU"
toplot$if_HU[intersect( grep("_HU", toplot$exp_series_date), which(toplot$sp == "Capsa")) ] <- "100mM HU"

toplot$ploidy_clust_if_HU <- paste(toplot$ploidy_clust, toplot$if_HU)
  
```




```{r}

  df <- toplot[toplot$sp == "Capsa",]
  df <- df[!grep("Undefined",df$ploidy_clust) ]
  
    actual_pos <- unique(df$bp_exp_ploidy_corr[df$validation == "ACTUAL"])
  
    df <- df[df$if_test_sp == "TEST_bootstrap",]
    
  #  df <- df[!is.na(df$sp) & !is.na(df$suns_for_clust),]

 ggplot(df, aes( bp_exp_ploidy_corr ) )+
   geom_histogram(aes(fill = ploidy_clust_if_HU), bins = 30, alpha = 0.25, position = "identity")+
     theme_bw()+ scale_x_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
   facet_wrap(~beatiful_suns)+
     geom_vline(xintercept = actual_pos[1] )+ 
     geom_vline(xintercept = actual_pos[2] )+
     xlab("predicted genome size, bp")+ylab("number of cells")
 
 df$if_HU[df$exp_series_date == "20032025_thickstacks_noHU"] <- "no HU control"
 df$if_HU[df$if_HU == "no HU"] <- NA
 
 
gg <-  ggplot(df, aes( ploidy_clust, bp_exp_ploidy_corr ) )+
    geom_violin()+
    geom_point(aes(colour=if_HU), alpha=0.5, size =0.25, position=position_jitter(width=0.2, height = 0))+
     theme_bw()+ scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
   facet_wrap(~beatiful_suns)+
     geom_hline(yintercept = actual_pos[1] )+ 
     geom_hline(yintercept = actual_pos[2] )+
     xlab("Predicted cell cycle stage")+ylab("Predicted genome size, bp")+
    scale_colour_manual(values = c("100mM HU" = "red", "no HU control" = "cyan"),  na.value = NA)

ggsave(gg, filename = "HU_treatment_BP.pdf", width = 10, height = 5)    
  
  
  
df$clusterisation <- str_split(df$beatiful_suns, ";", simplify = TRUE)[,1]    

df <- df[!duplicated(df[,c("clusterisation", "ploidy_clust", "nucl_sum_int","sp", "if_HU","id")]), ]
#length(unique(df$cell_coord))

 gg <-   ggplot(df, aes( ploidy_clust, nucl_sum_int ) )+
    geom_violin()+
      geom_boxplot(alpha=0)+
    geom_point(aes(colour=if_HU), alpha=0.5, size =0.25, position=position_jitter(width=0.2, height = 0))+
     theme_bw()+ scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
   facet_wrap(~clusterisation)+
     #geom_hline(yintercept = actual_pos[1] )+      geom_hline(yintercept = actual_pos[2] )+
     xlab("Predicted cell cycle stage")+ylab("Fluorescence intensity, A.U.")+
    scale_colour_manual(values = c("100mM HU" = "red", "no HU control" = "cyan"),  na.value = NA)

ggsave(gg, filename = "HU_treatment_INT.pdf", width = 5, height = 2.5)    

    
#df[df$if_HU=="100mM HU",] %>% group_by(clusterisation, ploidy_clust) %>% summarise(n_unique_cells = n_distinct(cell_coord)) 
#df[is.na(df$if_HU),] %>% group_by(clusterisation, ploidy_clust) %>% summarise(n_unique_cells = n_distinct(cell_coord))     
  
#df[df$suns_for_clust =="est_sp slidedate slide"]
 
```












## Summaries by median and expected bp sizes (NON-OUTLIER- FILTERED)

```{r}
########
median_summary <- data.frame()

for (species in unique(sp_by_sp$sp)) {
  for (date in unique(sp_by_sp$exp_series_date)) {
    for (slide in unique(sp_by_sp$file_prefix)) {
   tmp <- sp_by_sp[sp_by_sp$sp == species,]
   tmp <- tmp[tmp$exp_series_date == date,]
   tmp <- tmp[tmp$file_prefix == slide,]
   
   MS <- data.frame("sp" = species, "file_prefix" = slide, "exp_series_date" = date, "MEDIAN" = median(tmp$nucl_sum_int))
   median_summary <- rbind(median_summary, MS)
  }
  }
}


```   
   
   
   
   
## Ploidy check modification for plots above

```{r}

sp_by_sp_ploidy <- sp_by_sp
sp_by_sp_ploidy$stack <- NA
sp_by_sp_ploidy$stack[grep("thinstack" , sp_by_sp_ploidy$file_prefix)] <- "thinstack"
sp_by_sp_ploidy$stack[grep("thickstack" , sp_by_sp_ploidy$file_prefix)] <- "thickstack"

ggplot(sp_by_sp_ploidy[sp_by_sp_ploidy$sp == "Capsa",], aes(as.factor(file_prefix), as.numeric(nucl_sum_int) ))+geom_boxplot()+geom_violin(alpha = 0.5)+
     geom_jitter(aes(shape = as.factor(stack), colour=file_prefix),position=position_jitter(height=0, width=0.2), alpha = 0.3)+ #geom_point(aes(shape = sp, colour = file_prefix), size = 2, alpha = 0.7)+
    theme_classic()+ scale_y_continuous(labels = label_comma())+
    scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2)) + 
    #scale_shape_manual(values = 0:24)+
  theme(axis.text.x = element_blank())

median_summary_ploidy <- median_summary
median_summary_ploidy$stack <- NA
median_summary_ploidy$stack[grep("thinstack" , median_summary_ploidy$file_prefix)] <- "thinstack"
median_summary_ploidy$stack[grep("thickstack" , median_summary_ploidy$file_prefix)] <- "thickstack"

# ggplot(median_summary_ploidy, aes(bp_exp, MEDIAN))+geom_point(aes(colour=as.factor(file_prefix), shape = sp),size =3, alpha = 0.7)+theme_linedraw()+scale_x_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))


```   





## Confocal estimates



```{r}

confocal <- fread("confocal_meas/Confocal_meas.tsv")

confocal$expected_sizes <- NA 

for (sp in expected_sizes_trunc$sp) {
  confocal$expected_sizes[confocal$est_sp ==sp] <- expected_sizes_trunc$bp[expected_sizes_trunc$sp==sp]
}


 ggplot(confocal, aes(expected_sizes, RawIntDen))+geom_point(aes(colour=as.factor(slidedate), shape = est_sp),size =4)+
   geom_line(aes(group=id, colour=as.factor(slidedate) )  )+
  theme_linedraw()+
   facet_wrap(~deconvolved)+
   scale_x_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))+
   scale_y_continuous(trans = "log2", labels = scales::math_format(2^.x, format = log2))

```





